---
title: "forcats: Working with factor data in R"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: moon
    highlight: monochrome
    center: true
    reveal_options:
      slideNumber: true
      previewLinks: true
---

## Introduction

In this module, we will learn to work with categorical/qualitative data in R using [forcats](http://forcats.tidyverse.org). Let us begin by installing and loading forcats and a set of other pacakges we will be using.

```{r cat, eval=FALSE}
library(forcats)
library(tibble)
library(magrittr)
library(purrr)
library(dplyr)
library(ggplot2)
```

```{r cat1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(forcats)
library(tibble)
library(magrittr)
library(purrr)
library(dplyr)
library(ggplot2)
```

## Case Study

We will use a case study to explore the various features of the forcats package. You can download the data for the case study from [here](https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv) or directly import the data using the readr package. 

----

```{r import, eval=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
```

```{r show, echo=FALSE, eval=TRUE, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

## Data Dictionary

----

We will do the following in this case study:

- compute the frequency of different referrers
- plot average number of pages browsed for different referrers
- collapse referrers with low sample size into a single group
- club traffic from social media websites into a new category
- group referrers with traffic below a threshold into a single category
 
----

```{r cat2}
ecom %>%
  count(referrer)
```

----

```{r cat3}
refer_summary <- ecom %>%
  group_by(referrer) %>%
  summarise(
    page = mean(n_pages),
    tos = mean(duration),
    n = n()
  )
```

----

```{r cat10}
ggplot(refer_summary, aes(page, referrer)) + geom_point()
```

----

```{r cat11}
ggplot(refer_summary, aes(page, fct_reorder(referrer, page))) + geom_point()
```

----

```{r cat4}
ecom %>%
  mutate(ref = referrer %>% fct_infreq()) %>%
  ggplot(aes(ref)) +
  geom_bar()
```

----

```{r cat5}
ecom %>%
  mutate(ref = referrer %>% fct_infreq() %>% fct_rev()) %>%
  ggplot(aes(ref)) +
  geom_bar()
```

----

```{r import2, message=FALSE}
traffic <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web_traffic.csv')
traffic
```

----

```{r cat6}
traffic$traffics %>% fct_count()
```

----

```{r cat7}
traffic2 <- fct_collapse(traffic$traffics,
  social = c("facebook", "twitter", "instagram"),
  search = c("google", "bing", "yahoo")
)

traffic2 %>% fct_count() 
```

----

```{r cat8}
traffic$traffics %>% fct_lump() %>% table()
```

----

```{r cat9}
traffic$traffics %>% fct_lump(n = 3) %>% table()
```

----

```{r cat12}
traffic$traffics %>% fct_lump(prop = 0.1) %>% table()
```

----

```{r cat13}
traffic$traffics %>% fct_lump(prop = 0.15) %>% table()
```

----

```{r cat14}
traffic$traffics %>% fct_lump(n = -3) %>% table()
```

----

```{r cat15}
traffic$traffics %>% fct_lump(prop = -0.1) %>% table()
```

## Summary