---
output: _output.yaml
---

## {data-background="forcats.png"}

## Agenda

<hr>

- tabulate levels
- reorder levels
- reverse levels
- collapse levels
- recode levels
- recategorize levels
- shift levels

## Libraries

<hr>

```{r cat, eval=FALSE}
library(forcats)
library(magrittr)
library(dplyr)
library(readr)
```

```{r cat1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(forcats)
library(magrittr)
library(dplyr)
library(readr)
```


## Data 

<hr>

```{r import2, message=FALSE}
traffic <- read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web_traffic.csv',
  col_types = list(col_factor(levels = c("google", "facebook", "affiliates", 
    "bing", "yahoo", "twitter", "instagram", "unknown", "direct"))))

traffic
```

## Data

<hr>

```{r import2a}
traffics <- 
  traffic %>%
  use_series(traffics)

traffics
```


## Case Study

<hr>

In this case study, we will:

- compute the frequency of different referrers
- collapse referrers with low sample size into a single group
- club traffic from social media websites into a new category
- group referrers with traffic below a threshold into a single category

## {data-background="canva_fct_count.png"}
 
## Tabulate Referrers

<hr>

```{r cat2}
fct_count(traffics)
```

## Levels

<hr>

```{r cat22}
levels(traffics)
```

## {data-background="canva_fct_infreq.png"}

## Sort Referrer by Frequency

<hr>

```{r cat23}
traffics %>%
  fct_infreq() %>%
  levels()
```

## Sort Referrer by Order of Appearance 

<hr>

```{r inorder}
traffics %>%
  fct_inorder() %>%
  levels()
```

## {data-background="canva_fct_rev.png"}

## Reverse Levels

<hr>

```{r cat25}
traffics %>%
  fct_rev() %>%
  levels()
```

## Tabulate Referrer

<hr>

```{r cat6}
fct_count(traffics)
```

## {data-background="canva_fct_collapse.png"}

## Collapse Referrer Categories

<hr>

```{r cat7}
traffics %>% 
  fct_collapse(
  social = c("facebook", "twitter", "instagram"),
  search = c("google", "bing", "yahoo")) %>% 
  fct_count() 
```

## {data-background="canva_fct_lump_1.png"}

## Lump Infrequent Referrer Types

<hr>

```{r cat8, echo=FALSE}
fct_count(traffics)
```

```{r cat8b}
traffics %>% 
  fct_lump() %>% 
  table()
```

## Retain top 3 referrers

<hr>

```{r cat9, echo=FALSE}
traffics %>% 
  fct_count() %>% 
  arrange(desc(n))
```

```{r cat17}
traffics %>% 
  fct_lump(n = 3) %>% 
  table()
```

## {data-background="canva_fct_lump_2.png"}

## Lump Referrer Types with less than 10% traffic

<hr>

```{r cat12, echo=FALSE}
traffics %>% 
  fct_count() %>%
  mutate(
    percent = round((n / sum(n)) * 100, 2)
  )
```

```{r cat16}
traffics %>% 
  fct_lump(prop = 0.1) %>% 
  table()
```

## Lump Referrer Types with less than 15% traffic

<hr>

```{r cat13, echo=FALSE}
traffics %>% 
  fct_count() %>%
  mutate(
    percent = round((n / sum(n)) * 100, 2)
  )
```

```{r cat18}
traffics %>% 
  fct_lump(prop = 0.15) %>% 
  table()
```

## Retain 3 Referrer Types with lowest traffic

<hr>

```{r cat14, echo=FALSE}
traffics %>% 
  fct_count() %>% 
  arrange(n)
```

```{r cat19}
traffics %>% 
  fct_lump(n = -3) %>% 
  table()
```

## Retain 3 Referrer Types with less than 10% traffic

<hr>

```{r cat15, echo=FALSE}
traffics %>% 
  fct_count() %>%
  mutate(
    percent = round((n / sum(n)) * 100, 2)
  )
```

```{r cat20}
traffics %>% 
  fct_lump(prop = -0.1) %>% 
  table()
```

## {data-background="canva_fct_drop.png"}

## {data-background="canva_fct_others_1.png"}

## Replace Levels with Other

<hr>

```{r other_1}
fct_other(traffics, keep = c("google", "yahoo")) %>%
  levels()
```

## {data-background="canva_fct_others_2.png"}

## Replace Levels with Other

<hr>

```{r other_2}
fct_other(traffics, drop = c("instagram", "twitter")) %>%
  levels()
```

## {data-background="canva_fct_recode.png"}

## Recode Levels

<hr>

```{r recode}
fct_recode(traffics, search = "bing", search = "yahoo", search = "google",
  social = "facebook", social = "twitter", social = "instagram") %>%
  levels()
```

## {data-background="canva_fct_relevel_1.png"}

## Reorder Levels

<hr>

```{r relevel_1}
fct_relevel(traffics, "twitter") %>%
  levels()
```

## {data-background="canva_fct_relevel_2.png"}

## Reorder Levels

<hr>

```{r relevel_2}
fct_relevel(traffics, "google", after = 2) %>%
  levels()
```

## {data-background="canva_fct_relevel_3.png"}

## Reorder Levels

<hr>

```{r relevel_3}
fct_relevel(traffics, "facebook", after = Inf) %>%
  levels()
```

## Data 

<hr>

```{r import3, message=FALSE}
response <- read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/response.csv',
  col_types = list(col_factor(levels = c("like", "like somewhat", "neutral", 
    "dislike somewhat", "dislike"), ordered = TRUE)))

response
```

## Data

<hr>

```{r import2b}
responses <- 
  response %>%
  use_series(response)

responses
```


## {data-background="canva_fct_shift_1.png"}

## Shift Levels

<hr>

```{r shift_1}
fct_shift(responses, 2) %>%
  levels()
```

## {data-background="canva_fct_shift_2.png"}

## Shift Levels

<hr>

```{r shift_2}
fct_shift(responses, -2) %>%
  levels()
```

## {data-background="thankyou.png"}
  
<!-- To do: -->
<!--   - add agenda -->
<!--   - describe case study 2 -->
<!--   - modify ecommerce data -->