---
output: _output.yaml
---

## {data-background="forcats.png"}

## Agenda

<hr>

## Introduction

<hr>

In this module, we will learn to work with categorical/qualitative data in R using [forcats](http://forcats.tidyverse.org). Let us begin by installing and loading forcats and a set of other pacakges we will be using.

## Libraries

<hr>

```{r cat, eval=FALSE}
library(forcats)
library(tibble)
library(magrittr)
library(purrr)
library(dplyr)
library(ggplot2)
library(readr)
```

```{r cat1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(forcats)
library(tibble)
library(magrittr)
library(purrr)
library(dplyr)
library(ggplot2)
library(readr)
```

## Case Study

<hr>

We will use a case study to explore the various features of the forcats package. You can download the data for the case study from [here](https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv) or directly import the data using the readr package. 

## Data

<hr>

```{r import, eval=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
```

```{r show, echo=FALSE, eval=TRUE, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

## Data Dictionary

<hr>

- id: row id
- referrer: referrer website/search engine
- os: operating system
- browser: browser
- device: device used to visit the website
- n_pages: number of pages visited
- duration: time spent on the website (in seconds)
- repeat: frequency of visits
- country: country of origin
- purchase: whether visitor purchased
- order_value: order value of visitor (in dollars)


## Case Study

<hr>

We will do the following in this case study:

- compute the frequency of different referrers
- plot average number of pages browsed for different referrers
- collapse referrers with low sample size into a single group
- club traffic from social media websites into a new category
- group referrers with traffic below a threshold into a single category
 
## Tabulate Referrers

<hr>

```{r cat2}
ecom %>%
  count(referrer)
```

## Avg. page visits by referrer type

<hr>

```{r cat3}
refer_summary <- ecom %>%
  group_by(referrer) %>%
  summarise(
    page = mean(n_pages),
    tos = mean(duration),
    n = n()
  )

refer_summary
```

## Avg. page visits by referrer type

<hr>

```{r cat10, fig.align='center', fig.width=6, fig.height=4}
ggplot(refer_summary, aes(page, referrer)) + geom_point()
```

## Reorder Referrers

<hr>

```{r cat11, fig.align='center', fig.width=6, fig.height=4}
ggplot(refer_summary, aes(page, fct_reorder(referrer, page))) + geom_point()
```

## Plot Referrer Frequency (Descending Order)

<hr>

```{r cat21}
ecom %>%
  count(referrer, sort = TRUE)
```

## Plot Referrer Frequency (Descending Order)

<hr>

```{r cat22}
ecom %>%
  pull(referrer) %>%
  as_factor() %>%
  levels
```

## Plot Referrer Frequency (Descending Order)

<hr>

```{r cat23}
ecom %>%
  pull(referrer) %>%
  as_factor() %>%
  fct_infreq() %>%
  levels
```

## Plot Referrer Frequency (Descending Order)

<hr>

```{r cat4, fig.align='center', fig.width=6, fig.height=4}
ecom %>%
  mutate(ref = referrer %>% fct_infreq()) %>%
  ggplot(aes(ref)) +
  geom_bar()
```

## Plot Referrer Frequency (Ascending Order)

<hr>

```{r cat24}
ecom %>%
  pull(referrer) %>%
  as_factor() %>%
  levels
```

## Plot Referrer Frequency (Ascending Order)

<hr>

```{r cat25}
ecom %>%
  pull(referrer) %>%
  as_factor() %>%
  fct_rev() %>%
  levels 
```

## Plot Referrer Frequency (Ascending Order)

<hr>

```{r cat5, fig.align='center', fig.width=6, fig.height=4}
ecom %>%
  mutate(ref = referrer %>% fct_infreq() %>% fct_rev()) %>%
  ggplot(aes(ref)) +
  geom_bar()
```

## Case Study 2

<hr>

## Data 

<hr>

```{r import2, message=FALSE}
traffic <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web_traffic.csv')
traffic
```

## Tabulate Referrer

<hr>

```{r cat6}
traffic$traffics %>% fct_count()
```

## Collapse Referrer Categories

<hr>

```{r cat7}
traffic2 <- fct_collapse(traffic$traffics,
  social = c("facebook", "twitter", "instagram"),
  search = c("google", "bing", "yahoo")
)

traffic2 %>% fct_count() 
```

## Lump Infrequent Referrer Types

<hr>

```{r cat8}
traffic$traffics %>% fct_count()

traffic$traffics %>% fct_lump() %>% table()
```

## Retain top 3 referrers

<hr>

```{r cat9, echo=FALSE}
traffic$traffics %>% fct_count() %>% arrange(desc(n))
```

```{r cat17}
traffic$traffics %>% fct_lump(n = 3) %>% table()
```

## Lump Referrer Types with less than 10% traffic

<hr>

```{r cat12, echo=FALSE}
traffic$traffics %>% 
  fct_count() %>%
  mutate(
    percent = round((n / sum(n)) * 100, 2)
  )
```

```{r cat16}
traffic$traffics %>% fct_lump(prop = 0.1) %>% table()
```

## Lump Referrer Types with less than 15% traffic

<hr>

```{r cat13, echo=FALSE}
traffic$traffics %>% 
  fct_count() %>%
  mutate(
    percent = round((n / sum(n)) * 100, 2)
  )
```

```{r cat18}
traffic$traffics %>% fct_lump(prop = 0.15) %>% table()
```

## Retain 3 Referrer Types with lowest traffic

<hr>

```{r cat14, echo=FALSE}
traffic$traffics %>% fct_count() %>% arrange(n)
```

```{r cat19}
traffic$traffics %>% fct_lump(n = -3) %>% table()
```

## Retain 3 Referrer Types with less than 10% traffic

<hr>

```{r cat15, echo=FALSE}
traffic$traffics %>% 
  fct_count() %>%
  mutate(
    percent = round((n / sum(n)) * 100, 2)
  )
```

```{r cat20}
traffic$traffics %>% fct_lump(prop = -0.1) %>% table()
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add agenda -->
<!--   - describe case study 2 -->
<!--   - modify ecommerce data -->