---
output: _output.yaml
---

## {data-background="stringr.png"}

## Agenda

<hr>

- what are regular expressions?
- why use regular expression?
- what can you do with regular expressions?
	- find a word in a string
	- replace string
	- match a single character
	- match one of any of several letters
	- match series of range of characters
 
## Libraries

<hr>

```{r install, eval=FALSE}
library(stringr)
library(dplyr)
library(readr)
library(glue)
```

```{r str1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(stringr)
library(dplyr)
library(readr)
library(glue)
```

## Data

<hr>

```{r import, eval=FALSE}
mockstring <- read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/mock_strings.csv')
```

```{r show, echo=FALSE, eval=TRUE, message=FALSE}
mockstring <- read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/mock_strings.csv')
mockstring
```

## Sample Data

<hr>


```{r sample, echo = FALSE}
mock_data <-
  mockstring %>%
  slice(1:10) %>%
  select(email, address, full_name, currency)
```

```{r sample2}
mock_data
```

## cat

<hr>

## paste

<hr>

## paste0

<hr>

## format

<hr>

## glue

<hr>

## grep

<hr>

## grepl

<hr>

## sub

<hr>

## gsub

<hr>

## Meta Characters

<hr>

## Quantifiers

<hr>

## Sequences

<hr>

## Anchors

<hr>

## Character Classes

<hr>

## References

<hr>

- https://www.rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf
- https://stringr.tidyverse.org/articles/regular-expressions.html
- https://r4ds.had.co.nz/strings.html
- https://github.com/rstudio/cheatsheets/blob/master/strings.pdf
- https://www.garrickadenbuie.com/project/regexplain/

## {data-background="thankyou.png"}
