---
output: _output.yaml
---

## {data-background="img/mba_cover_image.png"}

## Connect With Us

<hr>

- Website     (https://www.rsquaredacademy.com/)
- Free Online R Courses   (https://rsquared-academy.thinkific.com/)
- R Packages  (https://pkgs.rsquaredacademy.com)
- Shiny Apps  (https://apps.rsquaredacademy.com)
- Blog        (https://blog.rsquaredacademy.com)
- GitHub      (https://github.com/rsquaredacademy)
- YouTube     (https://www.youtube.com/user/rsquaredin/)
- Twitter     (https://twitter.com/rsquaredacademy)
- Facebook    (https://www.facebook.com/rsquaredacademy/)
- Linkedin    (https://in.linkedin.com/company/rsquared-academy)

## Resources

<hr>

- <a href="https://slides.rsquaredacademy.com/mba/mba.html" target="_blank">Slides</a>
- <a href="https://github.com/rsquaredacademy-education/online-courses/tree/master/association-rule-mining-in-r" target="_blank">Code & Data</a>
- <a href="https://rstudio.cloud/project/335377" target="_blank">RStudio Cloud</a>

## {data-background="img/ws_agenda.png"}

<br>
<br>
<br>
<br>
<br>

>- What?
>- Why?
>- How?
>- Use Cases
>- Demo

## {data-background="img/ws_section_intro.png"}

## {data-background="img/mba_canva_intro.png"}

## {data-background="img/mba_canva_what.png"}

## {data-background="img/mba_canva_why.png"}

## {data-background="img/mba_canva_advantages.png"}

## {data-background="img/mba_canva_use_cases.png"}

## {data-background="img/mba_canva_simple_example.png"}

## {data-background="img/mba_canva_steps.png"}

## {data-background="img/mba_canva_itemset.png"}

## {data-background="img/mba_canva_antecedent_consequent.png"}

## {data-background="img/mba_canva_support.png"}

## {data-background="img/mba_canva_confidence.png"}

## {data-background="img/mba_canva_lift.png"}

## {data-background="img/ws_section_case_studies.png"}

## Data

<hr>

- <a href="http://archive.ics.uci.edu/ml/datasets/online+retail#" target="_blank">UCI</a>
- <a href="https://data.world/datasets/market-basket-analysis" target="_blank">data.world</a>

## Data Dictionary

<hr>

- invoice number
- stock code
- description
- quantity
- invoice date
- unit price
- customer id
- country

## Libraries

<hr>

```{r libs, message=FALSE, warning=FALSE}
library(readxl)
library(readr)
library(mbar)
library(arules)
library(arulesViz)
library(magrittr)
library(dplyr)
library(lubridate)
library(forcats)
library(ggplot2)
```

## Read Data

<hr>

```{r read_data, warning=FALSE, message=FALSE}
basket_data <- read.transactions("transaction_data.csv", format = "basket", 
  sep = ",")
basket_data
```

## Data Summary

<hr>

```{r summary}
summary(basket_data)
```

## {data-background="img/mba_canva_ifplot.png"}

## Generate Rules

<hr>

```{r rules_generate}
rules <- apriori(basket_data, parameter = list(supp=0.009, conf=0.8, 
	target = "rules", maxlen = 4))
```

## Rules Summary

<hr>

```{r rules_summary}
summary(rules)
```

## Inspect Rules

<hr>

```{r rules_inspect}
basket_rules <- sort(rules, by = 'confidence', decreasing = TRUE)
inspect(basket_rules[1:10])
```

## {data-background="img/mba_canva_redundant_rule_1.png"}

## {data-background="img/mba_canva_redundant_rule_2.png"}

## {data-background="img/mba_canva_redundant_rule_3.png"}

## Redundant Rules

<hr>

```{r rules_redundant_print}
inspect(rules[is.redundant(rules)])
```

## Non Redundant Rules

<hr>

```{r rules_unique}
inspect(rules[!is.redundant(rules)])
```

## What influenced purchase of product X?

<hr>

```{r rhs, message=FALSE}
sugar_rules <- apriori(basket_data, parameter = list(supp = 0.009, conf = 0.8), 
  appearance = list(default = "lhs", rhs = "SUGAR"))
```

## What influenced purchase of product X?

<hr>

```{r rhs2}
rules_sugar <- sort(sugar_rules, by = "confidence", decreasing = TRUE)
inspect(rules_sugar)
```

## What purchases did product X influence?

<hr>

```{r lhs, message=FALSE}
sugar_rules <- apriori(basket_data, parameter = list(supp = 0.009, conf = 0.8), 
  appearance = list(default = "rhs", lhs = "SUGAR")) 
```

## What purchases did product X influence?

<hr>

```{r lhs2, message=FALSE}
rules_sugar <- sort(sugar_rules, by = "confidence", decreasing = TRUE)
inspect(rules_sugar)
```

## Top Rules by Support

<hr>

```{r top_rules_supp}
supp_rules <- sort(rules, by = 'support', decreasing = TRUE)
top_rules <- supp_rules[1:10]
inspect(top_rules)
```

## Top Rules by Confidence

<hr>

```{r top_rules_conf}
conf_rules <- sort(rules, by = 'confidence', decreasing = TRUE)
top_rules <- conf_rules[1:10]
inspect(top_rules)
```

## Top Rules by Lift

<hr>

```{r top_rules_lift}
lift_rules <- sort(rules, by = 'lift', decreasing = TRUE)
top_rules <- lift_rules[1:10]
inspect(top_rules)
```

## {data-background="img/mba_canva_scatter_plot.png"}

## {data-background="img/mba_canva_network_plot.png"}

## {data-background="img/mba_canva_lift_warning.png"}

## {data-background="img/mba_canva_conf_warning.png"}

## {data-background="img/ws_section_summary.png"}

## {data-background="img/mba_summary_meme.png"}


## Summary

<hr>

- unsupervised data mining technique
- uncovers products frequently bought together
- creates if-then scenario rules
- cost-effective, insightful and actionable
- association rule mining has applications in several industries
- directionality of rule is lost while using lift
- confidence as a measure can be misleading

## {data-background="img/thankyou.png"}

