---
title: "Data Manipulation with dplyr - Part 3"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
---

<style type="text/css">
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: capitalize;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
-  }

  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

```{r libs, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(dplyr)
```


In this module, we will explore a set of helper functions in order to:

- extract unique rows 
- rename columns
- sample data
- extract columns
- slice rows
- arrange rows
- compare tables
- extract/mutate data using predicate functions
- count observations for different levels of a variable

## Case Study

<hr>

Let us look at a case study (e-commerce data) and see how we can use dplyr helper functions to answer questions we have about and to modify/transform the underlying data set. You can download the data from [here](https://github.com/rsquaredacademy/datasets/blob/master/web.csv) or import it directly using `read_csv()` from the [readr](readr.tidyverse.org) package.

## Data

<hr>

```{r import, eval=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
```



```{r show, echo=FALSE, eval=TRUE, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

## Data Dictionary

<hr>

- id: row id
- referrer: referrer website/search engine
- os: operating system
- browser: browser
- device: device used to visit the website
- n_pages: number of pages visited
- duration: time spent on the website (in seconds)
- repeat: frequency of visits
- country: country of origin
- purchase: whether visitor purchased
- order_value: order value of visitor (in dollars)


## Data Sanitization

<hr>

```{r dplyr1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(dplyr)
options(tibble.width = Inf)
```

## Traffic Sources

<hr>

```{r dplyr2}
ecom %>%
  distinct(referrer)
```

## Device Types

<hr>

```{r dplyr30}
ecom %>%
  distinct(device)
```

## Rename Columns

<hr>

Rename the `duration` column to `time_on_site`.

```{r dplyr3}
ecom %>%
  rename(time_on_site = duration)
```

## Sampling Data

<hr>

```{r dplyr4}
ecom %>%
  sample_n(700)
```

## Sampling Data

<hr>

```{r dplyr5}
ecom %>%
  group_by(referrer) %>%
  sample_n(100)
```

## Sampling Data

<hr>

```{r dplyr6}
ecom %>%
  sample_frac(size = 0.7)
```

## Sampling Data

<hr>

```{r dplyr7}
ecom %>%
  group_by(referrer) %>%
  sample_frac(0.3)
```


## Extract Columns

<hr>

Extract the `device` column.

```{r dplyr8}
ecom %>%
  pull(device)
```

## Extract Columns

<hr>

Extract the first column.

```{r dplyr9}
ecom %>%
  pull(1)
```

## Extract Columns

<hr>

Extract the last column

```{r dplyr10}
ecom %>%
  pull(-1)
```

## Slice Rows

<hr>

Extract the first 20 rows.

```{r dplyr11}
ecom %>%
  slice(1:20)
```

## Slice Rows

<hr>

Extract the last row.

```{r dplyr12}
ecom %>%
  slice(n())
```

## Tally 

<hr>

Total number of observations in the data.

```{r dplyr13}
ecom %>%
  tally()
```

## Tally

<hr>

Observations of different types of referrers.

```{r dplyr14}
ecom %>%
  group_by(referrer) %>%
  tally()
```

## Tally

<hr>

Observations of referrers and bouncers.

```{r dplyr15}
ecom %>%
  group_by(referrer, bouncers) %>%
  tally()
```

## Tally

<hr>

Observations of referrers and purchasers.

```{r dplyr16}
ecom %>%
  group_by(referrer, purchase) %>%
  tally()
```

## Tally

<hr>

```{r dplyr25}
ecom %>%
  group_by(referrer, purchase) %>%
  tally() %>%
  filter(purchase == 'true')
```

## Count

<hr>

Use `count()` instead of `tally()`.

```{r dplyr17}
ecom %>%
  count(referrer, purchase)
```

## Top Referrers

<hr>

Top 2 referrers that bring orders.

```{r dplyr18}
ecom %>%
  count(referrer, purchase) %>%
  filter(purchase == 'true') %>%
  arrange(desc(n)) %>%
  top_n(n = 2)
```

## Between

<hr>

```{r dplyr19}
ecom %>%
  pull(n_pages) %>%
  between(5, 15) 
```

## Case When

<hr>

```{r dplyr20}
mtcars %>%
  select(mpg, disp, cyl, gear, carb) %>%
  mutate(
    type = case_when(
      disp > 200 ~ 'dabba',
      cyl == 8 ~ 'kachda',
      TRUE ~ 'worst'
    )
  )
```

## Select Observation

<hr>

```{r dplyr21}
ecom %>%
  pull(referrer) %>%
  nth(1)
```

## Select Observation

<hr>

```{r dplyr22}
ecom %>%
  pull(referrer) %>%
  nth(1000)
```

## Select Observation

<hr>

```{r dplyr23}
ecom %>%
  pull(referrer) %>%
  last()
```

## Select Observation

<hr>

```{r dplyr24}
ecom %>%
  pull(referrer) %>%
  first()
```  

## Questions

<hr>

## Thank You

<hr>