---
output: _output.yaml
---

## {data-background="dplyr.png"}

## Agenda

<hr>

```{r libs, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(dplyr)
```

In the previous module, we learnt about dplyr verbs and used them to compute average order value for an e-commerce website data. In this module, we will learn to combine tables using different `*_join` functions provided in dplyr. Let us assume we have 2 tables **x** and **y**. 

We will explore the following type of `joins` 

- inner join
- left join
- right join
- semi join
- anti join
- full join

## {data-background="casestudy.png"}

## {data-background="join_case_study.png"}

```{r dplyr1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(dplyr)
options(tibble.width = Inf)
```

## Data: Orders

<hr>

```{r dplyr25, eval=FALSE}
order <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/order.csv')
```

```{r dplyr2, echo=FALSE, eval=TRUE, message=FALSE}
order <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/order.csv')
order
```

## Data: Customers

<hr>

```{r dplyr26, eval=FALSE}
customer <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/customer.csv')
```

```{r dplyr, echo=FALSE, eval=TRUE, message=FALSE}
customer <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/customer.csv')
customer
```

## {data-background="canva_join.png"}

## {data-background="innerjoin.png"}

## {data-background="canva_inner.png"}

## inner join

<hr>

```{r dplyr3}
customer %>%
  inner_join(order)
```

## {data-background="leftjoin.png"}

## {data-background="canva_left.png"}

## left join

<hr>

```{r dplyr4}
customer %>%
  left_join(order)
```

## {data-background="rightjoin.png"}

## {data-background="canva_right.png"}

## right join

<hr>

```{r dplyr5}
customer %>%
  right_join(order)
```

## {data-background="semijoin.png"}

## {data-background="canva_semi.png"}

## semi join

<hr>

```{r dplyr6}
customer %>%
  semi_join(order)
```

## {data-background="antijoin.png"}

## {data-background="canva_anti.png"}

## anti join

<hr>

```{r dplyr7}
customer %>%
  anti_join(order)
```

## {data-background="fulljoin.png"}

## {data-background="canva_full.png"}

## full join

<hr>

```{r dplyr8}
customer %>%
  full_join(order)
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - change agenda -->
<!--   - describe data sets in case study slide -->