---
output: _output.yaml
---

## {data-background="dplyr.png"}

## Agenda

<hr>

```{r libs, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(dplyr)
```


In this module, we will explore a set of helper functions in order to:

- extract unique rows 
- rename columns
- sample data
- extract columns
- slice rows
- arrange rows
- compare tables
- extract/mutate data using predicate functions
- count observations for different levels of a variable

## {data-background="casestudy.png"}

## Case Study

<hr>

Let us look at a case study (e-commerce data) and see how we can use dplyr helper functions to answer questions we have about and to modify/transform the underlying data set. You can download the data from [here](https://github.com/rsquaredacademy/datasets/blob/master/web.csv) or import it directly using `read_csv()` from the [readr](readr.tidyverse.org) package.

## Data

<hr>

```{r import, eval=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
```



```{r show, echo=FALSE, eval=TRUE, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

## Data Dictionary

<hr>

- id: row id
- referrer: referrer website/search engine
- os: operating system
- browser: browser
- device: device used to visit the website
- n_pages: number of pages visited
- duration: time spent on the website (in seconds)
- repeat: frequency of visits
- country: country of origin
- purchase: whether visitor purchased
- order_value: order value of visitor (in dollars)


## {data-background="sanitization.png"}

```{r dplyr1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(dplyr)
options(tibble.width = Inf)
```

## Data Sanitization

<hr>

- `distinct()`
- `rename()`

## Traffic Sources

<hr>

```{r dplyr2}
ecom %>%
  distinct(referrer)
```

## Device Types

<hr>

```{r dplyr30}
ecom %>%
  distinct(device)
```

## Rename Columns

<hr>

```{r dplyr3}
ecom %>%
  rename(time_on_site = duration)
```

## {data-background="sampling.png"}

## Sampling Data

<hr>

```{r dplyr4}
ecom %>%
  sample_n(700)
```

## Sampling Data

<hr>

```{r dplyr5}
ecom %>%
  group_by(referrer) %>%
  sample_n(100)
```

## Sampling Data

<hr>

```{r dplyr6}
ecom %>%
  sample_frac(size = 0.7)
```

## Sampling Data

<hr>

```{r dplyr7}
ecom %>%
  group_by(referrer) %>%
  sample_frac(0.3)
```


## {data-background="extract.png"}

## Extract Device Column

<hr>

```{r dplyr8}
ecom %>%
  pull(device)
```

## Extract First Column

<hr>

```{r dplyr9}
ecom %>%
  pull(1)
```

## Extract Last Column

<hr>

```{r dplyr10}
ecom %>%
  pull(-1)
```

## Extract First 20 Rows

<hr>

```{r dplyr11}
ecom %>%
  slice(1:20)
```

## Extract Last Row

<hr>

```{r dplyr12}
ecom %>%
  slice(n())
```

## {data-background="tabulate.png"}

## Total Observations in Data

<hr>

```{r dplyr13}
ecom %>%
  tally()
```

## Tabulate Referrers

<hr>

```{r dplyr14}
ecom %>%
  group_by(referrer) %>%
  tally()
```

## Tabulate referrers and bouncers

<hr>

```{r dplyr15}
ecom %>%
  group_by(referrer, bouncers) %>%
  tally()
```

## Tabulate referrers and purchasers

<hr>

```{r dplyr16}
ecom %>%
  group_by(referrer, purchase) %>%
  tally()
```

## Tabulate Referrers & Converts

<hr>

```{r dplyr25}
ecom %>%
  group_by(referrer, purchase) %>%
  tally() %>%
  filter(purchase == 'true')
```

## Count

<hr>

```{r dplyr17}
ecom %>%
  count(referrer, purchase)
```

## {data-background="sortfilter.png"}

## Top 2 referrers by orders

<hr>

```{r dplyr18}
ecom %>%
  count(referrer, purchase) %>%
  filter(purchase == 'true') %>%
  arrange(desc(n)) %>%
  top_n(n = 2)
```

## Between

<hr>

```{r dplyr19}
ecom %>%
  pull(n_pages) %>%
  between(5, 15) 
```

## Case When

<hr>

```{r dplyr20}
mtcars %>%
  select(mpg, disp, cyl, gear, carb) %>%
  mutate(
    type = case_when(
      disp > 200 ~ 'High',
      cyl == 8 ~ 'Eight',
      TRUE ~ 'True'
    )
  )
```

## Select First Observation

<hr>

```{r dplyr21}
ecom %>%
  pull(referrer) %>%
  nth(1)

ecom %>%
  pull(referrer) %>%
  first()
```

## Select 1000th Observation

<hr>

```{r dplyr22}
ecom %>%
  pull(referrer) %>%
  nth(1000)
```

## Select Last Observation

<hr>

```{r dplyr23}
ecom %>%
  pull(referrer) %>%
  last()
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - explain case study data -->
