---
output: _output.yaml
---

## {data-background="img/mba_cover_image.png"}

## Connect With Us

<hr>

- Website     (https://www.rsquaredacademy.com/)
- Free Online R Courses   (https://rsquared-academy.thinkific.com/)
- R Packages  (https://pkgs.rsquaredacademy.com)
- Shiny Apps  (https://apps.rsquaredacademy.com)
- Blog        (https://blog.rsquaredacademy.com)
- GitHub      (https://github.com/rsquaredacademy)
- YouTube     (https://www.youtube.com/user/rsquaredin/)
- Twitter     (https://twitter.com/rsquaredacademy)
- Facebook    (https://www.facebook.com/rsquaredacademy/)
- Linkedin    (https://in.linkedin.com/company/rsquared-academy)

## Resources

<hr>

- <a href="https://slides.rsquaredacademy.com/mba/mba.html" target="_blank">Slides</a>
- <a href="https://github.com/rsquaredacademy-education/online-courses/tree/master/association-rule-mining-in-r" target="_blank">Code & Data</a>
- <a href="https://rstudio.cloud/project/335377" target="_blank">RStudio Cloud</a>

## {data-background="img/ws_agenda.png"}

<br>
<br>
<br>
<br>
<br>
<br>

>- What?
>- Why?
>- How?
>- Demo

## {data-background="img/ws_section_intro.png"}

## Introduction

<hr>

```{r rfm_intro, fig.align='center', out.width="100%", echo=FALSE}

```

## What?

<hr>

```{r rfm_what, fig.align='center', out.width="100%", echo=FALSE}

```

## Why ?

<hr>

<br>
<br>

```{r rfm_why, fig.align='center', out.width="100%", echo=FALSE}

```

## {data-background="img/ws_section_case_studies.png"}

## Data

<hr>


## Data Dictionary

<hr>


## Libraries

<hr>

```{r libs, message=FALSE, warning=FALSE}
```

## {data-background="img/ws_section_summary.png"}

## Summary

<hr>


## {data-background="img/thankyou.png"}

