---
title: "ggplot2: Aesthetics"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

## Agenda

<style type="text/css">
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: capitalize;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
  }

  .reveal h2 {
    font-size: 22px;
    color: #0077aa;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
    font-size: 20px;
  }

  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>


<hr>

Explore aesthetics such as 

- color
- shape
- size
- fill
- alpha
- width

## Intro

<hr>

In this module, we will focus on the aesthetics i.e. color, shape, size, alpha, line type, line width etc. We can map these to variables or specify values for them. If we want to map the above to variables, we have to specify them within the `aes()` function. We will look at both methods.

## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Map Color to Variable

<hr>

```{r aes2, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg, color = factor(cyl))) +
  geom_point()
```

## Map Color to Variable

<hr>

```{r aes3, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(aes(color = factor(cyl)))
```

## Color: Specify Value

<hr>

```{r aes4, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(color = 'blue')
```

## Map Shape to Variable

<hr>

```{r aes5, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg, shape = factor(cyl))) +
  geom_point()
```

## Map Shape to Variable

<hr>

```{r aes6, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(aes(shape = factor(cyl)))
```

## Shape: Specify Value

<hr>

```{r aes7, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(shape = 5)
```

## Shape & Color

<hr>

```{r aes8, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(shape = 5, color = 'blue')
```

## Fill

<hr>

```{r aes9, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(shape = 5, fill = 'blue')
```

## Fill: Specify Value

<hr>

```{r aes10, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(shape = 22, fill = 'blue')
```

## Geom Border Color

<hr>

```{r aes11, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(shape = 22, color = 'blue')
```

## Map Size to Variable (Continuous)

<hr>

```{r aes12, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg, size = disp)) +
  geom_point()
```

## Specify Size Value

<hr>

```{r aes13, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(size = 4)
```

## Map Alpha to Variable

<hr>

```{r aes14, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(aes(alpha = factor(cyl)), color = 'blue')
```

## Line Color, Type & Width

<hr>

So far we have focussed on `geom_point()` to learn how to map aesthetics to variables. To explore line type and line width, we will use `geom_line()`. In the previous post, we used `geom_line()` to build line charts. Now we will modify the appearance of the line. In the section below, we will specify values for 

- line color
- line type 
- and line width

## Data

<hr>

```{r imports2, eval = FALSE}
gdp <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/gdp.csv')
```

```{r import2, echo=FALSE, eval=TRUE, message=FALSE}
gdp <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/gdp.csv')
gdp
```

## Data Dictionary

<hr>

It contains GDP (Gross Domestic Product) growth data for the BRICS (Brazil, Russia, India, China, South Africa) for the years 2000 to 2005.


## Line Color

<hr>

```{r aes15, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(color = 'blue')
```

## Line Type

<hr>

```{r aes16, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(linetype = 2)
```

## Line Type

<hr>

```{r aes17, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(linetype = 'dashed')
```

## Line Width

<hr>

```{r aes18, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(size = 2)
```

## Map Line Aesthetics to Variable

<hr>

Now let us map the aesthetics to the variables. The data used in the above example cannot be used as we need a variable with country names. We will use `gather()` function from the `tidyr` package to reshape the data.

## Data

<hr>

```{r aes19, fig.align='center'}
gdp2 <- gdp %>% 
  select(year, growth, india, china) %>% 
  gather(key = country, value = gdp, -year)

gdp2
```

## Grouped Line Chart

<hr>

```{r aes20, fig.align='center'}
ggplot(gdp2, aes(year, gdp, group = country)) +
  geom_line()
```

## Map Line Color to Country

<hr>

```{r aes21, fig.align='center'}
ggplot(gdp2, aes(year, gdp, group = country)) +
  geom_line(aes(color = country))
```

## Map Line Type to Country

<hr>

```{r aes22, fig.align='center'}
ggplot(gdp2, aes(year, gdp, group = country)) +
  geom_line(aes(linetype = country))
```

## Map Line Width to Country

<hr>

```{r aes23, fig.align='center'}
ggplot(gdp2, aes(year, gdp, group = country)) +
  geom_line(aes(size = country))
```

## Other Plots

<hr>

Before we wrap up, let us quickly see how we can map aesthetics to variables for 

- bar plots
- histograms
- box plots


## Data

<hr>

```{r import, eval=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
```

```{r use, echo=FALSE, eval=TRUE, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

## Data Dictionary

<hr>

- id: row id
- referrer: referrer website/search engine
- os: operating system
- browser: browser
- device: device used to visit the website
- n_pages: number of pages visited
- duration: time spent on the website (in seconds)
- repeat: frequency of visits
- country: country of origin
- purchase: whether visitor purchased
- order_value: order value of visitor (in dollars)

## Map Bar Color to Purchase

<hr>

```{r aes24, fig.align='center'}
ggplot(ecom, aes(device, fill = purchase)) +
  geom_bar()
```

## Map Histogram Color to Purchase

<hr>

```{r aes25, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(duration, fill = purchase), bins = 10)
```

## Map Box Plot Color to Purchase

<hr>

```{r aes26, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(device, duration, fill = purchase))
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - change line plot color -->
<!--   - add section slides -->