---
output: _output.yaml
---

## {data-background="gg_aesthetics.png"}

## Agenda

<hr>

Explore aesthetics such as 

- color
- shape
- size
- fill
- alpha
- width

## Intro

<hr>

In this module, we will focus on the aesthetics i.e. color, shape, size, alpha, line type, line width etc. We can map these to variables or specify values for them. If we want to map the above to variables, we have to specify them within the `aes()` function. We will look at both methods.

## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## {data-background="color.png"}

## Map Color to Variable

<hr>

```{r aes2, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg, color = factor(cyl))) +
  geom_point()
```

## Map Color to Variable

<hr>

```{r aes3, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(aes(color = factor(cyl)))
```

## Color: Specify Value

<hr>

```{r aes4, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(color = 'blue')
```

## {data-background="shape.png"}

## Map Shape to Variable

<hr>

```{r aes5, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg, shape = factor(cyl))) +
  geom_point()
```

## Map Shape to Variable

<hr>

```{r aes6, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(aes(shape = factor(cyl)))
```

## Shape: Specify Value

<hr>

```{r aes7, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(shape = 5)
```

## Shape & Color

<hr>

```{r aes8, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(shape = 5, color = 'blue')
```

## {data-background="fill.png"}

## Fill

<hr>

```{r aes9, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(shape = 5, fill = 'blue')
```

## Fill: Specify Value

<hr>

```{r aes10, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(shape = 22, fill = 'blue')
```

## Geom Border Color

<hr>

```{r aes11, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(shape = 22, color = 'blue')
```

## {data-background="size.png"}

## Map Size to Variable (Continuous)

<hr>

```{r aes12, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg, size = disp)) +
  geom_point()
```

## Specify Size Value

<hr>

```{r aes13, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(size = 4)
```

## Map Alpha to Variable

<hr>

```{r aes14, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(aes(alpha = factor(cyl)), color = 'blue')
```

## {data-background="other_plots.png"}

## Other Plots

<hr>

Before we wrap up, let us quickly see how we can map aesthetics to variables for 

- bar plots
- histograms
- box plots


## Data

<hr>

```{r import, eval=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
```

```{r use, echo=FALSE, eval=TRUE, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

## Data Dictionary

<hr>

- id: row id
- referrer: referrer website/search engine
- os: operating system
- browser: browser
- device: device used to visit the website
- n_pages: number of pages visited
- duration: time spent on the website (in seconds)
- repeat: frequency of visits
- country: country of origin
- purchase: whether visitor purchased
- order_value: order value of visitor (in dollars)

## Map Bar Color to Purchase

<hr>

```{r aes24, fig.align='center'}
ggplot(ecom, aes(device, fill = purchase)) +
  geom_bar()
```

## Map Histogram Color to Purchase

<hr>

```{r aes25, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(duration, fill = purchase), bins = 10)
```

## Map Box Plot Color to Purchase

<hr>

```{r aes26, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(device, duration, fill = purchase))
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - change line plot color -->