---
output: _output.yaml
---

## {data-background="gg_axis_labels.png"}

## Agenda

<hr>

- add title and subtitle to the plot
- modify axis labels
- modify axis range
- remove axis
- format axis

## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Basic Plot

<hr>

```{r axis2, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg))
```

## Title & Subtitle

<hr>

There are two ways to add title to a plot:

- `ggtitle()`
- `labs()`

## ggtitle()

<hr>

Let us explore the `ggtitle()` function first. It takes two arguments:

- label: title of the plot
- subtitle: subtitle of the plot

## Title & Subtitle

<hr>

```{r axis3, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  ggtitle(label = 'Displacement vs Mileage', subtitle = 'disp vs mpg')
```

## Axis Labels

<hr>

You can add labels to the axis using:

- `xlab()`
- `ylab()`
- `labs()`

## Axis Labels

<hr>

```{r axis4, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  xlab('Displacement') + ylab('Miles Per Gallon')
```

## Labs

<hr>

The `labs()` function can be used to add the following to a plot:

- title
- subtitle
- X axis label
- Y axis label

## Labs

```{r axis5, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  labs(title = 'Displacement vs Mileage', subtitle = 'disp vs mpg', 
    x = 'Displacement', y = 'Miles Per Gallon')
```

## Axis Range

<hr>

In certain scenarios, you may want to modify the range of the axis. In ggplot2, we can achieve this using:

- `xlim()`
- `ylim()`
- `expand_limits()`

## Axis Range

<hr>

- `xlim()` and `ylim()` take a numeric vector of length 2 as input
- `expand_limits()` takes two numeric vectors (each of length 2), one for each axis 
- in all of the above functions, the first element represents the lower limit and the second element represents the upper limit

## X Axis

<hr>

```{r axis6, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  xlim(c(0, 600))
```

## Y Axis

<hr>

```{r axis7, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  ylim(c(0, 40))
```

## Expand Limits

<hr>

```{r axis8, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  expand_limits(x = c(0, 600), y = c(0, 40))
```

## Remove Axis Labels

<hr>

If you want to remove the axis labels all together, use the `theme()` function. It allows us to modify every aspect of the theme of the plot. Within `theme()`, set the following to `element_blank()`.

- `axis.title.x`
- `axis.title.y`

## Remove Axis Labels using theme()

<hr>

```{r axis9, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank())
```


## Format Title & Axis Labels

<hr>

To format the title or the axis labels, we have to modify the theme of the plot using the `theme()` function.
We can modify: 

- color
- font family
- font face
- font size 
- horizontal alignment 
- and angle

## Color

<hr>

```{r axis10, fig.align='center', message=FALSE, warning=FALSE}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + ggtitle('Diaplacement vs Mileage') +
  theme(plot.title = element_text(color = 'blue'))
```

## Font Family

<hr>

```{r axis11, fig.align='center', message=FALSE, warning=FALSE}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + ggtitle('Diaplacement vs Mileage') +
  theme(plot.title = element_text(family = 'Arial'))
```

## Font Face

<hr>

The font face can be any of the following:

- `plain` 
- `bold`
- `italic` 
- `bold.italic`

## Font Face 

```{r axis12, fig.align='center', message=FALSE, warning=FALSE}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + ggtitle('Diaplacement vs Mileage') +
  theme(plot.title = element_text(face = 'bold'))
```

## Font Size

<hr>

```{r axis13, fig.align='center', message=FALSE, warning=FALSE}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + ggtitle('Diaplacement vs Mileage') +
  theme(plot.title = element_text(size = 8))
```

## Horizontal Alignment

<hr>

```{r axis14, fig.align='center', message=FALSE, warning=FALSE}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + ggtitle('Diaplacement vs Mileage') +
  theme(plot.title = element_text(hjust = 0.3))
```

## Title

<hr>

```{r axis15, fig.align='center', message=FALSE, warning=FALSE}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + ggtitle('Diaplacement vs Mileage') +
  theme(plot.title = element_text(color = 'blue', family = 'Arial',
    face = 'bold', size = 12, hjust = 0.5))
```

## Putting it all together.. X Axis

<hr>

```{r axis16, fig.align='center', message=FALSE, warning=FALSE}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + xlab('Diaplacement') +
  theme(axis.title.x = element_text(color = 'blue', family = 'Arial',
    face = 'bold', size = 8, hjust = 0.5, angle = 15))
```


## Putting it all together.. Y Axis

<hr>

```{r axis17, fig.align='center', message=FALSE, warning=FALSE}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + ylab('Miles Per Gallon') +
  theme(axis.title.y = element_text(color = 'blue', family = 'Arial',
    face = 'italic', size = 8, vjust = 0.3, angle = 90))
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - section for ggtitle, labs, axis labels, format and putting it all together -->