---
title: "ggplot2: Axis & Plot Labels"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

<style type="text/css">
    .reveal p img {
    align: middle;
  }
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: capitalize;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
    font-size: 25px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal p img {
    align: middle;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

In this module, we will learn to:

- add title and subtitle to the plot
- modify axis labels
- modify axis range
- remove axis
- format axis

## Intro

<hr>

## Installation

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Basic Plot

<hr>

Let us start by building a simple scatter plot. 

----

```{r axis2, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg))
```

## Title & Subtitle

<hr>

You can add a title and subtitle to the plot using the `ggtitle()` function.

----

```{r axis3, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  ggtitle(label = 'Displacement vs Mileage', subtitle = 'disp vs mpg')
```

## Axis Labels

<hr>

The axis labels can be modified using the `xlab()` and `ylab()` functions.

----

```{r axis4, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  xlab('Displacement') + ylab('Miles Per Gallon')
```

## Labs

<hr>

We can add all of the above using the `labs()` function as well.

----

```{r axis5, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  labs(title = 'Displacement vs Mileage',
    subtitle = 'disp vs mpg', x = 'Displacement',
    y = 'Miles Per Gallon')
```

## Axis Range

<hr>

Sometimes you may want to modify the range of the axis. 

Use the `xlim()` function to modify the range of the X axis.

----

```{r axis6, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  xlim(c(0, 600))
```

## Axis Range

<hr>

Use the `ylim()` function to modify the range of the Y axis.

----

```{r axis7, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  ylim(c(0, 40))
```

## Axis Range

<hr>

You can use the `expand_limits()` function as well.

----

```{r axis8, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  expand_limits(x = c(0, 600), y = c(0, 40))
```

## Remove Axis Labels

<hr>

If you want to remove the axis labels all together, use the `theme()` function. What we are doing above is to modify the theme of the plot and set the axis labels to blank using `element_blank()` function.

----

```{r axis9, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  theme(axis.title.x = element_blank(),
    axis.title.y = element_blank())
```

## Format Title & Axis Labels

<hr>

To format the title or the axis labels, we have to modify the theme of the plot using the `theme()` function.
We can modify the color, font family, font face, font size and horizontal position. Let us try them one by one.

The font family can be any of the valid font family types but it must be supported by OS.
The font face can be plain, bold, italic, bold.italic
hjust must be between 0 (extreme left) and 1 (extreme right). 

----

```{r axis10, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  ggtitle('Diaplacement vs Mileage') +
  theme(plot.title = element_text(color = 'blue',
                                  family = 'Arial',
                                  face = 'bold',
                                  size = 8,
                                  hjust = 0.3))
```

## X axis label

<hr>

```{r axis11, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  xlab('Diaplacement') +
  theme(axis.title.x = element_text(color = 'blue',
                                  family = 'Arial',
                                  face = 'bold',
                                  size = 8,
                                  hjust = 0.5,
                                  angle = 15))
```


## Y axis label

<hr>

```{r axis12, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  ylab('Miles Per Gallon') +
  theme(axis.title.y = element_text(color = 'blue',
                                  family = 'Arial',
                                  face = 'italic',
                                  size = 8,
                                  vjust = 0.3,
                                  angle = 90))
```

## Questions

<hr>

## Thank You

<hr>