---
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
    css: styles.css
---

## {data-background="thankyou.png"}

## Agenda

<hr>

- build 
    - simple bar plot
    - stacked bar plot
    - grouped bar plot
    - proportional bar plot
- map aesthetics to variables
- specify values for 
    - bar color
    - bar line color
    - bar line type
    - bar line size


## Intro

<hr>

## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Data

<hr>

```{r import, eval=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
```


```{r import2, echo=FALSE, eval=TRUE, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```


## Data Dictionary

<hr>

- id: row id
- referrer: referrer website/search engine
- os: operating system
- browser: browser
- device: device used to visit the website
- n_pages: number of pages visited
- duration: time spent on the website (in seconds)
- repeat: frequency of visits
- country: country of origin
- purchase: whether visitor purchased
- order_value: order value of visitor (in dollars)

## Aesthetics

<hr>

- `fill`
- `color`
- `linetype`
- `size`
- `position`

## Simple Bar Plot

<hr>

```{r bar2, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)))
```

## Bar Color

<hr>

```{r bar3, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)),
           fill = c('red', 'blue', 'green'))
```

## Stacked Bar Plot

<hr>

```{r bar7, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(device, fill = factor(referrer)))
```

## Grouped Bar Plot

<hr>

```{r bar8, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(device, fill = factor(referrer)), position = 'dodge')
```

## Proportional Bar Plot

<hr>

```{r bar10, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(device, fill = factor(referrer)), position = 'fill')
```

## Horizontal Bar Plot

<hr>

```{r bar9, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device), fill = factor(referrer))) +
  coord_flip()
```

## Bar Line Color

<hr>

```{r bar4, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)), fill = 'white',
           color = c('red', 'blue', 'green'))
```

## Bar Line Type

<hr>

```{r bar5, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)), fill = 'white', 
          color = 'black', linetype = 2)
```

## Bar Line Size

<hr>

```{r bar6, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)), fill = 'white', 
           color = 'black', size = 2)
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - add introduction -->