---
title: "ggplot2: Bar Plots"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

<style type="text/css">
    .reveal p img {
    align: middle;
  }
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: capitalize;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
    font-size: 25px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal p img {
    align: middle;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

## Intro

<hr>

## Installation

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```


## Case Study

<hr>

We will continue to use the ecommerce data. You can download it from [here](https://github.com/rsquaredacademy/datasets/blob/master/web.csv) or import directly using the readr package.

----

```{r import2, echo=FALSE, eval=FALSE, results='hide', message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
```

```{r import, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

## Data Dictionary

<hr>

## Aesthetics

<hr>

- `fill`
- `color`
- `linetype`
- `size`
- `position`

## Simple Bar Plot

<hr>

```{r bar2, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)))
```

## Bar Color

<hr>

```{r bar3, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)),
           fill = c('red', 'blue', 'green'))
```

## Bar Line Color

<hr>

```{r bar4, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)), fill = 'white',
           color = c('red', 'blue', 'green'))
```

## Bar Line Type

<hr>

```{r bar5, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)), fill = 'white', 
          color = 'black', linetype = 2)
```

## Bar Line Size

<hr>

```{r bar6, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)), fill = 'white', 
           color = 'black', size = 2)
```

## Stacked Bar Plot

<hr>

```{r bar7, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(device, fill = factor(referrer)))
```

## Grouped Bar Plot

<hr>

```{r bar8, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(device, fill = factor(referrer)), position = 'dodge')
```

## Horizontal Bar Plot

<hr>

```{r bar9, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device), fill = factor(referrer))) +
  coord_flip()
```

## Summary

<hr>

## Next Steps

<hr>
