---
output: _output.yaml
---

## {data-background="gg_bar.png"}

## Agenda

<hr>

- build 
    - simple bar plot
    - stacked bar plot
    - grouped bar plot
    - proportional bar plot
- map aesthetics to variables
- specify values for 
    - bar color
    - bar line color
    - bar line type
    - bar line size


## Intro

<hr>

- a bar plot represents data in rectangular bars 
- the length of the bars are proportional to the values they represent 
- bar plots can be either horizontal or vertical 
- the X axis of the plot represents the levels or the categories
- and the Y axis represents the frequency/count of the variable 

<!-- ## Univariate Bar Plot -->

<!-- <hr> -->

<!-- A univariate bar plot represents a single categorical variable. The categories in the variable are represented on the X axis and their frequencies on the Y axis. In the below example, the `cyl` variable from the `mtcars` data set is visualized using a bar plot. The categories or levels are `4`, `6` and `8` which represent the number of cylinders in the automobile and are represented on the X axis. The frequency for each type of cylinder is represented by the Y axis. -->


## Libraries

<hr>

```{r show1, eval=FALSE}
library(ggplot2)
library(readr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(readr)
```

## Data

<hr>

```{r import, eval=FALSE}
ecom <- read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/ecom.csv',
  col_types = list(col_factor(levels = c('Desktop', 'Mobile', 'Tablet')), 
  col_factor(levels = c(TRUE, FALSE)), col_factor(levels = c(TRUE, FALSE)), 
  col_factor(levels = c('Affiliates', 'Direct', 'Display', 'Organic', 'Paid', 'Referral', 'Social'))))
```

```{r show, echo=FALSE, eval=TRUE, message=FALSE}
ecom <- read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/ecom.csv',
  col_types = list(col_factor(levels = c('Desktop', 'Mobile', 'Tablet')), 
  col_factor(levels = c(TRUE, FALSE)), col_factor(levels = c(TRUE, FALSE)), 
  col_factor(levels = c('Affiliates', 'Direct', 'Display', 'Organic', 'Paid', 'Referral', 'Social'))))
ecom
```

## Data Dictionary

<hr>

Below is the description of the data set:

- device: device used to visit the website
- bouncers: whether visit was a bouncer (exit website from landing page)
- purchase: whether visitor purchased
- referrer: referrer website/search engine

## Aesthetics

<hr>

- `fill`
- `color`
- `linetype`
- `size`
- `position`

## Simple Bar Plot

<hr>

```{r bar2, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)))
```

## Bar Color

<hr>

```{r bar3, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)),
           fill = c('red', 'blue', 'green'))
```

## Stacked Bar Plot

<hr>

```{r bar7, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(device, fill = factor(referrer)))
```

## Grouped Bar Plot

<hr>

```{r bar8, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(device, fill = factor(referrer)), position = 'dodge')
```

## Proportional Bar Plot

<hr>

```{r bar10, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(device, fill = factor(referrer)), position = 'fill')
```

## Horizontal Bar Plot

<hr>

```{r bar9, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device), fill = factor(referrer))) +
  coord_flip()
```

## Bar Line Color

<hr>

```{r bar4, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)), fill = 'white',
           color = c('red', 'blue', 'green'))
```

## Bar Line Type

<hr>

```{r bar5, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)), fill = 'white', 
          color = 'black', linetype = 2)
```

## Bar Line Size

<hr>

```{r bar6, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)), fill = 'white', 
           color = 'black', size = 2)
```

## {data-background="thankyou.png"}
