---
output: _output.yaml
---

## {data-background="gg_box.png"}

## Agenda

<hr>

- build box plots
- modify box 
    - color
    - fill
    - alpha
    - line size
    - line type
- modify outlier 
    - color
    - shape 
    - size 
    - alpha

## Introduction

<hr>

- the box plot is a standardized way of displaying the distribution of data  
- box plots are useful for detecting outliers and for comparing distributions 
- it shows the shape, central tendancy and variability of the data

## Structure

<hr>

- the body of the boxplot consists of a "box" (hence, the name), which goes from the first quartile (Q1) to the third quartile (Q3)
- within the box, a vertical line is drawn at the Q2, the median of the data set
- two horizontal lines, called whiskers, extend from the front and back of the box 
- the front whisker goes from Q1 to the smallest non-outlier in the data set, and the back whisker goes from Q3 to the largest non-outlier
- if the data set includes one or more outliers, they are plotted separately as points on the chart

## Libraries

<hr>

```{r show3, eval=FALSE}
library(ggplot2)
library(readr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(readr)
```

## Data

<hr>

```{r import, eval=FALSE}
daily_returns <- read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/tickers.csv')
```

```{r show, echo=FALSE, eval=TRUE, message=FALSE}
daily_returns <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/tickers.csv')
daily_returns
```

## Univariate Box Plot

<hr>

```{r box2, fig.align='center'}
ggplot(daily_returns) +
  geom_boxplot(aes(x = factor(1), y = AAPL))
```

## Data

<hr>

```{r import2, eval=FALSE}
tidy_returns <- read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/tidy_tickers.csv')
```

```{r show2, echo=FALSE, eval=TRUE, message=FALSE}
tidy_returns <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/tidy_tickers.csv')
tidy_returns
```

## Box Plot

<hr>

```{r box3, fig.align='center'}
ggplot(tidy_returns) +
  geom_boxplot(aes(x = factor(stock), y = returns))
```

## Horizontal Box Plot

<hr>   

```{r box4, fig.align='center'}
ggplot(tidy_returns) +
  geom_boxplot(aes(x = factor(stock), y = returns)) +
  coord_flip()
```

## Notch

<hr>

```{r box5, fig.align='center'}
ggplot(tidy_returns) +
  geom_boxplot(aes(x = factor(stock), y = returns),
    notch = TRUE) 
```

## Jitter

<hr>

```{r box6, fig.align='center'}
ggplot(tidy_returns, aes(x = factor(stock), y = returns)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, color = 'blue')
```

## {data-background="outliers.png"}

## Outliers

<hr>

- color
- shape
- size
- alpha

## Outlier Color

<hr>

```{r box7, fig.align='center'}
ggplot(tidy_returns) +
  geom_boxplot(aes(x = factor(stock), y = returns),
               outlier.color = 'red')
```

## Outlier Shape

<hr>


```{r box9, fig.align='center'}
ggplot(tidy_returns) +
  geom_boxplot(aes(x = factor(stock), y = returns), outlier.shape = 23) 
```

## Outlier Size

<hr>

```{r box10, fig.align='center'}
ggplot(tidy_returns) +
  geom_boxplot(aes(x = factor(stock), y = returns), outlier.size = 3) 
```

## Outlier Alpha

<hr>

```{r box11, fig.align='center'}
ggplot(tidy_returns) +
  geom_boxplot(aes(x = factor(stock), y = returns),
               outlier.color = 'blue', outlier.alpha = 0.3) 
```

## {data-background="aesthetics.png"}

## Box Aesthetics

<hr>

- color
- fill
- alpha
- line type
- line width


## Specify Values for Fill

<hr>

```{r box12, fig.align='center'}
ggplot(tidy_returns) +
  geom_boxplot(aes(x = factor(stock), y = returns),
               fill = c('blue', 'red', 'green', 'yellow', 'brown')) 
```

## Map Fill to Variable

<hr>

```{r box13, fig.align='center'}
ggplot(tidy_returns) +
  geom_boxplot(aes(x = factor(stock), y = returns,
               fill = factor(stock))) 
```

## Specify Values for Alpha

```{r box14, fig.align='center'}
ggplot(tidy_returns) +
  geom_boxplot(aes(x = factor(stock), y = returns),
               fill = 'blue', alpha = 0.3) 
```

## Specify Values for Color

<hr>

```{r box15, fig.align='center'}
ggplot(tidy_returns) +
  geom_boxplot(aes(x = factor(stock), y = returns),
               color = c('blue', 'red', 'green', 'yellow', 'brown')) 
```

## Map Color to Variables

<hr>

```{r box16, fig.align='center'}
ggplot(tidy_returns) +
  geom_boxplot(aes(x = factor(stock), y = returns,
               color = factor(stock))) 
```


## Specify Values for Line Width

<hr>

```{r box17, fig.align='center'}
ggplot(tidy_returns) +
  geom_boxplot(aes(x = factor(stock), y = returns),
               size = 1.5) 
```

## Specify Values for Line Type

<hr>

```{r box18, fig.align='center'}
ggplot(tidy_returns) +
  geom_boxplot(aes(x = factor(stock), y = returns),
               linetype = 2) 
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - add intro -->
<!--   - outlier section -->
<!--   - aesthetics section -->