---
title: "ggplot2: Box Plots"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: moon
    highlight: monochrome
    center: true
    fig_width: 6
    fig_height: 5
    reveal_options:
      slideNumber: true
      previewLinks: true
---

## Intro

## Installation

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```


## Case Study

We will continue to use the ecommerce data. You can download it from [here](https://github.com/rsquaredacademy/datasets/blob/master/web.csv) or import directly using the readr package.

```{r import2, echo=FALSE, eval=FALSE, results='hide', message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
```

```{r import, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

## Data Dictionary

## Simple Box Plot

----

```{r box2}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(1), y = n_visit))
```

----

Now we need to mention something for X axis and hence we use the number 1. If we do not mention anything, ggplot2 will return an error. You can generally use any categorical variable for the X axis. Below is an example:

----

```{r box3}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = n_visit))
```

----

If you want horizontal boxplots, use `coord_flip()`. 

```{r box4}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = n_visit)) +
  coord_flip()
```

## Notch

----

Notches are used to compare groups; if the notches of two boxes do not overlap, this suggests that the medians are significantly different.

----

```{r box5}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = n_visit),
               notch = TRUE) 
```

## Jitter

----

Overlay points on boxplots.

```{r box6}
ggplot(ecom, aes(x = factor(device), y = n_visit)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, color = 'blue')
```

## Outliers

### Color

----

```{r box7}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               outlier.color = 'red')
```

## Shape

----

```{r box8}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               outlier.shape = 23)
```

----

Let us extend the Y axis and see if the ouliers appear more legible.

```{r box9}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               outlier.shape = 23) +
  expand_limits(y = c(0, 1100))
```

## Size

----

```{r box10}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               outlier.size = 3) +
  expand_limits(y = c(0, 1100))
```

Increasing the size is not very useful.

## Alpha

----

```{r box11}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               outlier.color = 'blue', outlier.alpha = 0.3) +
  expand_limits(y = c(0, 1100))
```

Decreasing the alpha increases the visibility of the outlier points.

## Box Color

----

```{r box12}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               fill = c('blue', 'red', 'green')) +
  expand_limits(y = c(0, 1100))
```

----

Map box color to variable.

```{r box13}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration,
               fill = factor(device))) +
  expand_limits(y = c(0, 1100))
```

----

```{r box14}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               fill = 'blue', alpha = 0.3) +
  expand_limits(y = c(0, 1100))
```

## Line Color

----

```{r box15}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               color = c('blue', 'red', 'green')) +
  expand_limits(y = c(0, 1100))
```

----

```{r box16}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration,
               color = factor(device))) +
  expand_limits(y = c(0, 1100))
```


## Line Width

----

```{r box17}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               size = 1.5) +
  expand_limits(y = c(0, 1100))
```

## Line Type

----

```{r box18}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               linetype = 2) +
  expand_limits(y = c(0, 1100))
```

## Summary

In this module, we learnt to:

- build box plots
- modify outlier color, shape, size etc.
- modify box color
- modify box line color, size and line type

## Next Steps

In the next module, we will learn to build histograms.
