---
title: "ggplot2: Box Plots"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

<style type="text/css">
    .reveal p img {
    align: middle;
  }
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: none;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
    font-size: 25px;
  }
  .reveal h1 {
    font-size: 24px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal p img {
    align: middle;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

- build box plots
- modify box 
    - color
    - fill
    - alpha
    - line size
    - line type
- modify outlier 
    - color
    - shape 
    - size 
    - alpha

## Intro

<hr>

## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Data

<hr>

```{r import2, eval=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
```

```{r import, echo=FALSE, eval=TRUE, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

## Data Dictionary

<hr>

- id: row id
- referrer: referrer website/search engine
- os: operating system
- browser: browser
- device: device used to visit the website
- n_pages: number of pages visited
- duration: time spent on the website (in seconds)
- repeat: frequency of visits
- country: country of origin
- purchase: whether visitor purchased
- order_value: order value of visitor (in dollars)


## Univariate Box Plot

<hr>

```{r box2, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(1), y = n_visit))
```

## Box Plot

<hr>

```{r box3, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = n_visit))
```

## Horizontal Box Plot

<hr>   

```{r box4, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = n_visit)) +
  coord_flip()
```

## Notch

<hr>

```{r box5, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = n_visit),
               notch = TRUE) 
```

## Jitter

<hr>

```{r box6, fig.align='center'}
ggplot(ecom, aes(x = factor(device), y = n_visit)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, color = 'blue')
```

## Outliers

<hr>

- color
- shape
- size
- alpha

## Outlier Color

<hr>

```{r box7, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               outlier.color = 'red')
```

## Outlier Shape

<hr>

```{r box8, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               outlier.shape = 23)
```

## Outlier Shape

<hr>


```{r box9, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration), outlier.shape = 23) +
  expand_limits(y = c(0, 1100))
```

## Outlier Size

<hr>

```{r box10, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration), outlier.size = 3) +
  expand_limits(y = c(0, 1100))
```

## Outlier Alpha

<hr>

```{r box11, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               outlier.color = 'blue', outlier.alpha = 0.3) +
  expand_limits(y = c(0, 1100))
```

## Box Aesthetics

<hr>

- color
- fill
- alpha
- line type
- line width


## Specify Values for Fill

<hr>

```{r box12, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               fill = c('blue', 'red', 'green')) +
  expand_limits(y = c(0, 1100))
```

## Map Fill to Variable

<hr>

```{r box13, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration,
               fill = factor(device))) +
  expand_limits(y = c(0, 1100))
```

## Specify Values for Alpha

```{r box14, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               fill = 'blue', alpha = 0.3) +
  expand_limits(y = c(0, 1100))
```

## Specify Values for Color

<hr>

```{r box15, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               color = c('blue', 'red', 'green')) +
  expand_limits(y = c(0, 1100))
```

## Map Color to Variables

<hr>

```{r box16, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration,
               color = factor(device))) +
  expand_limits(y = c(0, 1100))
```


## Specify Values for Line Width

<hr>

```{r box17, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               size = 1.5) +
  expand_limits(y = c(0, 1100))
```

## Specify Values for Line Type

<hr>

```{r box18, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               linetype = 2) +
  expand_limits(y = c(0, 1100))
```

## Questions

<hr>

## Thank You

<hr>
