---
output: _output.yaml
---

## {data-background="gg_facet.png"}

## Agenda

<hr>


## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
```

## {data-background="casestudy.png"}

## Facet - Intro

<hr>

Generate subplots that each display one subset of the data using

- `facet_grid()`
- `facet_wrap()`

## {data-background="grid.png"}

## Columns

<hr>

```{r facet1, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) + 
  geom_point() +
  facet_grid(. ~ cyl)
```

## Rows

<hr>

```{r facet2, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) + 
  geom_point() +
  facet_grid(cyl ~ .)
```

## Rows & Columns

<hr>

```{r facet3, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) + 
  geom_point() +
  facet_grid(gear ~ cyl)
```

## Rows & Columns

```{r facet4, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) + 
  geom_point() +
  facet_grid(cyl ~ gear)
```

## Scales

<hr>  

```{r facet5, fig.align='center'}
ggplot(mtcars, aes(disp, mpg, color = factor(cyl))) +
  geom_point() + 
  facet_grid(. ~ cyl, scales = "free")
```

## Switch Axis

<hr>

```{r facet6, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) + 
  geom_point() + 
  facet_grid(cyl ~ gear, switch = "both") 
```

## {data-background="wrap.png"}

## Wrap

<hr>

```{r facet7, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point() +
  facet_wrap(~cyl)
```

## Specify Rows

<hr>

```{r facet8, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point() +
  facet_wrap(~cyl, nrow = 2)
```

## Specify Rows

<hr>

```{r facet9, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point() +
  facet_wrap(~cyl, nrow = 3)
```

## Rows & Columns

<hr>

```{r facet10, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point() +
  facet_wrap(~cyl + gear)
```

## Rows & Columns

<hr>

```{r facet11, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point() +
  facet_wrap(c("cyl", "gear"))
```

## Scales

<hr>

```{r facet12, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point() +
  facet_wrap(~cyl, scales = "free")
```

## {data-background="thankyou.png"}

