---
title: "ggplot2: Facets"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
    css: styles.css
---

## Agenda

<hr>


## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Data

<hr>

```{r import2, eval=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
```

```{r import, echo=FALSE, eval=TRUE, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

## Data Dictionary

<hr>

- id: row id
- referrer: referrer website/search engine
- os: operating system
- browser: browser
- device: device used to visit the website
- n_pages: number of pages visited
- duration: time spent on the website (in seconds)
- repeat: frequency of visits
- country: country of origin
- purchase: whether visitor purchased
- order_value: order value of visitor (in dollars)

## Facet - Intro

<hr>

Generate subplots that each display one subset of the data using

- `facet_grid()`
- `facet_wrap()`

## Columns

<hr>

```{r facet1, fig.align='center'}
ggplot(ecom, aes(n_pages, duration)) + 
  geom_point() +
  facet_grid(. ~ device)
```

## Rows

<hr>

```{r facet2, fig.align='center'}
ggplot(ecom, aes(n_pages, duration)) + 
  geom_point() +
  facet_grid(device ~ .)
```

## Rows & Columns

<hr>

```{r facet3, fig.align='center'}
ggplot(ecom, aes(n_pages, duration)) + 
  geom_point() +
  facet_grid(purchase ~ device)
```

## Rows & Columns

```{r facet4, fig.align='center'}
ggplot(ecom, aes(n_pages, duration)) + 
  geom_point() +
  facet_grid(device ~ purchase)
```

## Scales

<hr>  

```{r facet5, fig.align='center'}
ggplot(ecom, aes(n_pages, duration, color = factor(device))) +
  geom_point() + 
  facet_grid(. ~ device, scales = "free")
```

## Switch Axis

<hr>

```{r facet6, fig.align='center'}
ggplot(ecom, aes(n_pages, duration)) + 
  geom_point() + 
  facet_grid(device ~ purchase, switch = "both") 
```

## Wrap

<hr>

```{r facet7, fig.align='center'}
ggplot(ecom, aes(n_pages, duration)) +
  geom_point() +
  facet_wrap(~device)
```

## Specify Rows

<hr>

```{r facet8, fig.align='center'}
ggplot(ecom, aes(n_pages, duration)) +
  geom_point() +
  facet_wrap(~device, nrow = 2)
```

## Specify Rows

<hr>

```{r facet9, fig.align='center'}
ggplot(ecom, aes(n_pages, duration)) +
  geom_point() +
  facet_wrap(~device, nrow = 3)
```

## Rows & Columns

<hr>

```{r facet10, fig.align='center'}
ggplot(ecom, aes(n_pages, duration)) +
  geom_point() +
  facet_wrap(~device + purchase)
```

## Rows & Columns

<hr>

```{r facet11, fig.align='center'}
ggplot(ecom, aes(n_pages, duration)) +
  geom_point() +
  facet_wrap(c("device", "purchase"))
```

## Scales

<hr>

```{r facet12, fig.align='center'}
ggplot(ecom, aes(n_pages, duration)) +
  geom_point() +
  facet_wrap(~device, scales = "free")
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - add agenda -->
<!--   - separate section for grid and wrap -->
