---
title: "ggplot2: Geoms"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

<style type="text/css">
    .reveal p img {
    align: middle;
  }
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: capitalize;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
    font-size: 25px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal p img {
    align: middle;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

- Scatter Plot
- Bar Plot
- Box Plot
- Histogram
- Line Chart
- Fit Line

## Intro

<hr>

In this module, we will create some of the most routinely used plots to explore data using the `geom_*` functions. 

## Installation

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
```

## Case Study

<hr>

You can download the data from [here](https://github.com/rsquaredacademy/datasets/blob/master/web.csv) or import it directly using `read_csv()` from the [readr](readr.tidyverse.org) package.

```{r import, eval=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
```

```{r use, echo=FALSE, eval=TRUE, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

## Data Dictionary

<hr>

## Scatter Plot

<hr>

## Point

<hr>

```{r geoms2, fig.align='center'}
ggplot(ecom, aes(x = n_pages, y = duration)) + 
  geom_point()
```

## Fit Line

<hr>

We can fit a regression line or add horizontal or vertical lines to the plots using 

- `geom_abline()`
- `geom_hline()`
- `geom_vline()`


## Vertical Line

<hr>

```{r geoms3, fig.align='center'}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  geom_vline(xintercept = 5) 
```

## Horizontal Line

<hr>

```{r geoms4, fig.align='center'}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  geom_hline(yintercept = 30) 
```

## Regression Line

<hr>

```{r geoms5, fig.align='center'}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() + 
  geom_abline(intercept = 10, slope = 1)
```

## Regression Line

<hr>

```{r geoms6, fig.align='center'}
ggplot(mtcars, aes(x = wt, y = mpg)) +
	geom_smooth(method = 'lm', se = TRUE)
```

## Regression Line

<hr>

```{r geoms7, fig.align='center'}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_smooth(method = 'loess', se = FALSE)
```

## Bar Plot

<hr>


## Bar

<hr>

```{r geoms8, fig.align='center'}
ggplot(ecom, aes(x = factor(device))) +
  geom_bar()
```

## Bar

<hr>

```{r geoms9, fig.align='center'}
ggplot(ecom, aes(x = factor(device))) +
  geom_bar(aes(weight = order_value))
```

## Bar

<hr>

```{r geoms10, fig.align='center'}
ggplot(ecom, aes(x = factor(device))) +
  geom_bar(aes(fill = purchase))
```

## Bar

<hr>

```{r geoms11, fig.align='center'}
ggplot(ecom, aes(x = factor(device))) +
  geom_bar(aes(fill = purchase)) +
  coord_flip()
```

## Bar

<hr>

```{r geoms12, fig.align='center'}
device <- c('laptop', 'mobile', 'tablet')
visits <- c(30000, 12000, 5000)
traffic <- tibble::tibble(device, visits)
ggplot(traffic, aes(x = device, y = visits)) +
  geom_col(fill = 'blue') 
```

## Boxplot

<hr>

- examine the distribution of a variable  
- detect outliers, boxplots are very handy

## Boxplot

<hr>

```{r geoms13, fig.align='center'}
ggplot(ecom, aes(x = factor(device), y = n_pages)) +
  geom_boxplot()
```

## Jitter

<hr>

```{r geoms14, fig.align='center'}
ggplot(ecom, aes(x = factor(device), y = n_pages)) +
  geom_boxplot() +
  geom_jitter()
```

## Histogram

<hr>

## Histogram

<hr>

```{r geoms15, fig.align='center'}
ggplot(ecom, aes(x = duration)) +
  geom_histogram()
```

## Bins

<hr>

```{r geoms16, fig.align='center'}
ggplot(ecom, aes(x = duration)) +
  geom_histogram(bins = 5)
```

## Line

<hr>


## Data

<hr>

```{r imports2, eval=FALSE}
gdp <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/gdp.csv')
```

```{r import2, echo=FALSE, eval=TRUE, message=FALSE}
gdp <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/gdp.csv')
gdp
```

## Line Chart

<hr>

```{r geomline1, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line()
```


## Line Color & Type 

<hr>

```{r geomline2, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(color = 'blue', linetype = 'dashed')
```

## Jitter

<hr>

```{r geoms17, fig.align='center'}
ggplot(ecom, aes(x = factor(device), y = duration)) +
  geom_jitter()
```

## Jitter

<hr>

```{r geoms18, fig.align='center'}
ggplot(ecom, aes(x = factor(device), y = duration)) +
  geom_jitter(width = 0.25, height = 0.5)
```

## Label

<hr>

```{r geoms20, fig.align='center'}
ggplot(mtcars, aes(disp, mpg, label = rownames(mtcars))) +
  geom_label()
```

## Text

<hr>

```{r geoms19, fig.align='center'}
ggplot(mtcars, aes(disp, mpg, label = rownames(mtcars))) +
  geom_text(check_overlap = TRUE, size = 2)
```

## Text

<hr>

```{r geoms21, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg, label = rownames(mtcars))) +
  geom_point() + 
  geom_text(aes(color = cyl), hjust = 0, nudge_x = 0.05, 
            size = 2, angle = 45)
```

## Questions

<hr>

## Thank You

<hr>