---
output: _output.yaml
---

## {data-background="gg_geoms.png"}

## Agenda

<hr>

Create some of the most routinely used plots to explore data using the `geom_*` functions:

- Scatter Plot
- Bar Plot
- Box Plot
- Histogram
- Line Chart
- Regression Line

## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
```

## Data

<hr>

```{r import, eval=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
```

```{r use, echo=FALSE, eval=TRUE, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

## Data Dictionary

<hr>

- id: row id
- referrer: referrer website/search engine
- os: operating system
- browser: browser
- device: device used to visit the website
- n_pages: number of pages visited
- duration: time spent on the website (in seconds)
- repeat: frequency of visits
- country: country of origin
- purchase: whether visitor purchased
- order_value: order value of visitor (in dollars)


## Point

<hr>

```{r geoms2, fig.align='center'}
ggplot(ecom, aes(x = n_pages, y = duration)) + 
  geom_point()
```

## Regression Line

<hr>

- `geom_abline()`
- `geom_smooth()`

## Regression Line

<hr>

```{r geoms5, fig.align='center'}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() + 
  geom_abline(intercept = 37.285, slope = -5.344)
```

## Regression Line

<hr>

```{r geoms6, fig.align='center'}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_smooth(method = 'lm', se = TRUE)
```

## Regression Line

<hr>

```{r geoms7, fig.align='center'}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_smooth(method = 'loess', se = FALSE)
```

## Horizontal/Vertical Lines

<hr>

- `geom_hline()`
- `geom_vline()`


## Horizontal Line

<hr>

```{r geoms4, fig.align='center'}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  geom_hline(yintercept = 30) 
```

## Vertical Line

<hr>

```{r geoms3, fig.align='center'}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  geom_vline(xintercept = 5) 
```

## Bar Plot

<hr>

## Frequency

<hr>

```{r geoms8, fig.align='center'}
ggplot(ecom, aes(x = factor(device))) +
  geom_bar()
```

## Weight

<hr>

```{r geoms9, fig.align='center'}
ggplot(ecom, aes(x = factor(device))) +
  geom_bar(aes(weight = order_value))
```

## Columns

<hr>

```{r geoms12, fig.align='center', fig.height=4}
device <- c('laptop', 'mobile', 'tablet')
visits <- c(30000, 12000, 5000)
traffic <- tibble::tibble(device, visits)
ggplot(traffic, aes(x = device, y = visits)) +
  geom_col(fill = 'blue') 
```

## Boxplot

<hr>

## Boxplot

<hr>

```{r geoms13, fig.align='center'}
ggplot(ecom, aes(x = factor(device), y = n_pages)) +
  geom_boxplot()
```

## Add Jitter

<hr>

```{r geoms14, fig.align='center'}
ggplot(ecom, aes(x = factor(device), y = n_pages)) +
  geom_boxplot() +
  geom_jitter()
```

## Histogram

<hr>

## Histogram

<hr>

```{r geoms15, fig.align='center'}
ggplot(ecom, aes(x = duration)) +
  geom_histogram()
```

## Bins

<hr>

```{r geoms16, fig.align='center'}
ggplot(ecom, aes(x = duration)) +
  geom_histogram(bins = 5)
```

## Line

<hr>

## Data

<hr>

```{r imports2, eval=FALSE}
gdp <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/gdp.csv')
```

```{r import2, echo=FALSE, eval=TRUE, message=FALSE}
gdp <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/gdp.csv')
gdp
```

## Line Chart

<hr>

```{r geomline1, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line()
```

## Jitter

<hr>

```{r geoms17, fig.align='center'}
ggplot(ecom, aes(x = factor(device), y = duration)) +
  geom_jitter()
```

## Jitter Width & Height

<hr>

```{r geoms18, fig.align='center'}
ggplot(ecom, aes(x = factor(device), y = duration)) +
  geom_jitter(width = 0.25, height = 0.5)
```

## Label

<hr>

```{r geoms20, fig.align='center'}
ggplot(mtcars, aes(disp, mpg, label = rownames(mtcars))) +
  geom_label()
```

## Text

<hr>

```{r geoms19, fig.align='center'}
ggplot(mtcars, aes(disp, mpg, label = rownames(mtcars))) +
  geom_text(check_overlap = TRUE, size = 2)
```

## Text

<hr>

```{r geoms21, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg, label = rownames(mtcars))) +
  geom_point() + 
  geom_text(aes(color = cyl), hjust = 0, nudge_x = 0.05, 
            size = 2, angle = 45)
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - change line chart data -->
<!--   - remove description of plots -->
<!--   - remove plot variations -->