---
title: "ggplot2: Legends - Part 1"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

<style type="text/css">
    .reveal p img {
    align: middle;
  }
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: none;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
    font-size: 25px;
  }
  .reveal h1 {
    font-size: 24px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal p img {
    align: middle;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

Modify legend when the aesthetics are mapped to variables. Specifically

- title
- breaks
- limits
- range
- labels
- values
- symbol type

## Intro

<hr>

Quite often the aesthetics are mapped to variables ggplot2 automatically creates legends wherever applicable. 
You may want to modify the appearance of legends. In this module, we will learn to modify the legends when
the following aesthetics are mapped to categorical/continuous variables

- continuous variables
    - alpha
    - size

- discrete variables
    - alpha
    - color
    - fill
    - shape
    - size


## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Continuous Variables

<hr>

If you have mapped aesthetics to a continuous variable, you can modify the legend

- title
- breaks
- limits 
- range
- and labels


## Map Size to Continuous Variable

<hr>

```{r leg36, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp))
```

## Legend Title

<hr>

```{r leg37, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp)) +
  scale_size_continuous(name = "Horsepower")
```

## Legend Range

<hr>

```{r leg38, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp)) +
  scale_size_continuous(range = c(3, 6))
```

## Legend Limits

<hr>

```{r leg39, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp)) +
  scale_size_continuous(limits = c(0, 400))
```

## Legend Breaks

<hr>

```{r leg40, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp)) +
  scale_size_continuous(breaks = c(100, 200, 300, 400))
```

## Legend Labels

<hr>

```{r leg41, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp)) +
  scale_size_continuous(breaks = c(100, 200, 300, 400),
    labels = c("Hundred", "2 Hundred", "3 Hundred", "4 Hundred"))
```

## Size - Continuous

<hr>

```{r leg42, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp)) +
  scale_size_continuous(name = "Horsepower", range = c(3, 6), 
    limits = c(0, 400), breaks = c(100, 200, 300, 400),
    labels = c("Hundred", "2 Hundred", "3 Hundred", "4 Hundred"))
```



## Map Alpha to Continuous Variable

<hr>

```{r leg2, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = hp), color = 'blue')
```

## Legend Title

<hr>

```{r leg3, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = hp), color = 'blue') +
  scale_alpha_continuous("Horsepower")
```

## Legend Breaks

<hr>

```{r leg4, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = hp), color = 'blue') +
  scale_alpha_continuous(breaks = c(0, 100, 200, 300, 400))
```

## Legend Limits

<hr>

```{r leg5, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = hp), color = 'blue') +
  scale_alpha_continuous(limits = c(0, 400))
```

## Legend Range

<hr>

```{r leg6, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = hp), color = 'blue') +
  scale_alpha_continuous(range = c(0.4, 0.8))
```

## Legend Labels

<hr>

```{r leg7, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = hp), color = 'blue') +
  scale_alpha_continuous(breaks = c(0, 100, 200, 300, 400),
    labels = c("Zero", "Hundred", "2 Hundred", 
      "3 Hundred", "4 Hundred"))
```

## Alpha - Continuous

<hr>

```{r leg8, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = hp), color = 'blue') +
  scale_alpha_continuous("Horsepower", breaks = c(0, 100, 200, 300, 400),
    limits = c(0, 400), range = c(0.4, 0.8),
    labels = c("Zero", "Hundred", "2 Hundred", 
      "3 Hundred", "4 Hundred"))
```




## Alpha Discrete

<hr>

If you have mapped aesthetics to a discrete variable, you can to modify legend:

- title
- breaks
- range
- limits
- labels
- and values

## Map Color to Discrete Variable

<hr>

```{r leg15, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl)))
```

## Legend Title

<hr>

```{r leg16, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(name = "Cylinders", 
    values = c("red", "blue", "green"))
```

## Legend Values

<hr>

```{r leg17, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"))
```

## Legend Labels

<hr>

```{r leg18, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    limits = c(4, 6, 8))
```

## Legend Limits

<hr>

```{r leg19, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    labels = c('Four', 'Six', 'Eight'))
```

## Legend Breaks

<hr>

```{r leg20, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    breaks = c(4, 6, 8))
```

## Color - Discrete Variable

<hr>

```{r leg21, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(name = "Cylinders", values = c("red", "blue", "green"),
    labels = c('Four', 'Six', 'Eight'), limits = c(4, 6, 8), breaks = c(4, 6, 8))
```

## Map Fill to Discrete Variable

<hr>

```{r leg22, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = factor(cyl)), shape = 22)
```

## Legend Title

<hr>

```{r leg23, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = factor(cyl)), shape = 22) +
  scale_fill_manual(name = "Cylinders", 
    values = c("red", "blue", "green"))
```

## Legend Values


<hr>

```{r leg24, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = factor(cyl)), shape = 22) +
  scale_fill_manual(values = c("red", "blue", "green"))
```

## Legend Labels

<hr>

```{r leg25, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = factor(cyl)), shape = 22) +
  scale_fill_manual(values = c("red", "blue", "green"),
    labels = c('Four', 'Six', 'Eight'))
```

## Legend Limits

<hr>

```{r leg26, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = factor(cyl)), shape = 22) +
  scale_fill_manual(values = c("red", "blue", "green"),
    limits = c(4, 6, 8))
```

## Legend Breaks

<hr>

```{r leg27, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = factor(cyl)), shape = 22) +
  scale_fill_manual(values = c("red", "blue", "green"),
    breaks = c(4, 6, 8))
```

## Legend Fill

<hr>

```{r leg28, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = factor(cyl)), shape = 22) +
  scale_fill_manual(name = "Cylinders", values = c("red", "blue", "green"),
    labels = c('Four', 'Six', 'Eight'), limits = c(4, 6, 8), breaks = c(4, 6, 8))
```

## Map Shape to Discrete Variable

<hr>

```{r leg29, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl)))
```

## Legend Title

<hr>

```{r leg30, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape_manual(name = "Cylinders",values = c(4, 12, 24))
```

## Legend Values

<hr>

```{r leg31, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape_manual(values = c(4, 12, 24))
```

## Legend Labels

<hr>

```{r leg32, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape_manual(values = c(4, 12, 24), labels = c('Four', 'Six', 'Eight'))
```

## Legend Limits

<hr>

```{r leg33, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape_manual(values = c(4, 12, 24), limits = c(4, 6, 8))
```

## Legend Breaks

<hr>

```{r leg34, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape_manual(values = c(4, 12, 24), breaks = c(4, 6, 8))
```

## Shape - Discrete Variable

<hr>

```{r leg35, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape_manual(name = "Cylinders", labels = c('Four', 'Six', 'Eight'),  
     values = c(4, 12, 24), limits = c(4, 6, 8), breaks = c(4, 6, 8))
```

## Map Alpha to Discrete Variable

<hr>

```{r leg9, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue')
```

## Legend Title

<hr>

```{r leg10, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue') +
  scale_alpha_discrete(name = "Cylinders")
```

## Legend Breaks

<hr>

```{r leg11, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue') +
  scale_alpha_discrete(breaks = c("4", "6", "8"))
```

## Legend Range

<hr>

```{r leg12, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue') +
  scale_alpha_discrete(range = c(0.4, 0.8))
```

## Legend Labels

<hr>

```{r leg13, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue') +
  scale_alpha_discrete(labels = c("Four", "Six", "Eight"))
```

## Alpha - Discrete Variable

<hr>

```{r leg14, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue') +
  scale_alpha_discrete(name = "Cylinders", range = c(0.4, 0.8),
    breaks = c("4", "6", "8"), labels = c("Four", "Six", "Eight"))
```


## Shape & Size

<hr>

If you have mapped shape/size to a discrete variable which has less than six categories, you can use the following functions:

- `scale_shape()`
- `scale_size()`

## Map Shape to Discrete Variable

<hr>

```{r leg50, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape()
```

## Legend Title

<hr>

```{r leg51, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape(name = 'Cylinders')
```

## Legend Labels

<hr>

```{r leg52, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape(labels = c('Four', 'Six', 'Eight'))
```

## Legend Limits

<hr>

```{r leg53, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape(limits = c(4, 6, 8))
```

## Legend Breaks

<hr>

```{r leg54, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape(breaks = c(4, 6, 8))
```

## Legend Types

<hr>

```{r leg55, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape(solid = FALSE)
```

## Shape - Discrete Variable

<hr>

```{r leg56, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape(solid = FALSE, name = 'Cylinders', limits = c(4, 6, 8),
    labels = c('Four', 'Six', 'Eight'), breaks = c(4, 6, 8))
```

## Questions

<hr>

## Thank You

<hr>