---
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

<style type="text/css">
    .reveal p img {
    align: middle;
  }
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: capitalize;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
    font-size: 25px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal p img {
    align: middle;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

## Intro

<hr>

## Installation

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## X Axis - Continuous

<hr>

`scale_x_continuous()`: modify x axis when x is continuous

The first function we will explore is `scale_x_continuous()`. Let us create a simple scatter plot and then use `scale_x_continuous()` 
to modify the X axis.It can be used for:

- axis title
- axis labels
- axis limits
- axis breaks
- and axis position

----

```{r guide2, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg))
```

----

```{r guide3, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  scale_x_continuous(name = "Displacement",
                     limits = c(0, 600),
                     breaks = c(0, 150, 300, 450, 600),
                     labels = c('0', '150', '300', '450', '600'),
                     position = 'top')
```


## Y Axis - Continuous

<hr>

`scale_y_continuous()`: modify y axis

Now let us modify the Y axis.

----

```{r guide4, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  scale_y_continuous(name = "Miles Per Gallon",
                     limits = c(0, 45),
                     breaks = c(0, 15, 30, 45),
                     labels = c('0', '15', '30', '45'),
                     position = 'right')
```

## X Axis - Discrete

<hr>

Now let us look at plots with discrete X axis.

```{r guide5, fig.align='center'}
ggplot(mtcars) +
  geom_bar(aes(factor(cyl)))
```

----

Use `scael_x_discrete()` to modify the X axis.

----

```{r guide6, fig.align='center'}
ggplot(mtcars) +
  geom_bar(aes(factor(cyl))) +
  scale_x_discrete(name = "Number of Cylinders",
                   labels = c("4" = "Four", "6" = "Six", "8" = "Eight"),
                   breaks = c("4", "6", "8"),
                   position = "bottom")
```

## Questions

<hr>

## Thank You

<hr>