---
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
    css: styles.css
---

## {data-background="thankyou.png"}

## Agenda

<hr>

Learn to modify legend

- title
- label
- bar

## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Title

<hr>

- position
    - top
    - bottom
    - left
    - right
- alignment
    - 0 (left)
    - 1 (right)


## Legend Title

<hr>

```{r leg1, fig.align='center'}
ggplot(mtcars) + geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    guide = guide_legend(title = "Cylinders"))
```

## Title Alignment

<hr>

```{r leg2, fig.align='center'}
ggplot(mtcars) + geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    guide = guide_legend(title = "Cylinders", title.hjust = 0.5))
```

## Title Position

<hr>

```{r leg3, fig.align='center'}
ggplot(mtcars) + geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    guide = guide_legend(title = "Cylinders", title.hjust = 0.5, 
      title.position = "top"))
```


## Label

<hr>

- position
    - top
    - bottom
    - left 
    - right
- alignment 
    - 0 (left)
    - 1 (right)
- direction
    - horizontal
    - vertical
- reverse
    - TRUE
    - FALSE
- rows

## Label Position

<hr>

```{r leg4, fig.align='center'}
ggplot(mtcars) + geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    guide = guide_legend(label.position = "right"))
```

## Label Alignment

<hr>

```{r leg5, fig.align='center'}
ggplot(mtcars) + geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    guide = guide_legend(label.hjust = 0.5))
```

## Direction

<hr>

```{r leg6, fig.align='center'}
ggplot(mtcars) + geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    guide = guide_legend(direction = "horizontal"))
```

## Rows

<hr>

```{r leg7, fig.align='center'}
ggplot(mtcars) + geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    guide = guide_legend(nrow = 2))
```

## Reverse

<hr>

```{r leg8, fig.align='center'}
ggplot(mtcars) + geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    guide = guide_legend(reverse = TRUE))
```

## Guides

<hr>

```{r leg20, fig.align='center'}
ggplot(mtcars) + geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    guide = guide_legend(title = "Cylinders", title.hjust = 0.5, 
      title.position = "top", label.position = "right", 
      direction = "horizontal", label.hjust = 0.5, nrow = 2, reverse = TRUE)
  )
```

## Title Alignment (Vertical)

<hr>

```{r leg22, fig.align='center'}
ggplot(mtcars) + geom_point(aes(disp, mpg, color = hp)) +
  scale_color_continuous(guide = guide_colorbar(
    title = "Horsepower", title.position = "top", title.vjust = 1))
```

## Labels Alignment (Vertical)

<hr>

```{r leg23, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = hp)) +
  scale_color_continuous(guide = guide_colorbar(
    label.vjust = 0.8))
```

## Legend Bar 

<hr>

- width
- height
- bins
- ticks
- limits


## Legend Bar Width

<hr>

```{r leg24, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = hp)) +
  scale_color_continuous(guide = guide_colorbar(
    barwidth = 10))
```

## Legend Bar Height

<hr>

```{r leg25, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = hp)) +
  scale_color_continuous(guide = guide_colorbar(
    barheight = 3))
```

## Legend Bar Bins

<hr>

```{r leg26, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = hp)) +
  scale_color_continuous(guide = guide_colorbar(
    nbin = 4))
```

## Legend Bar Ticks

<hr>

```{r leg28, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = hp)) +
  scale_color_continuous(guide = guide_colorbar(
    ticks = FALSE))
```

## Legend Bar Upper/Lower Limits

<hr>

```{r leg29, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = hp)) +
  scale_color_continuous(guide = guide_colorbar(
    draw.ulim = TRUE, draw.llim = FALSE))
```

## Guides

<hr>

```{r leg34, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = hp, 
    size = qsec, shape = factor(gear))) + 
  guides(color = "colorbar", shape = "legend", size = "legend")
```

## Guides: Color, Shape & Size

<hr>

```{r leg35, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = hp, size = wt, shape = factor(gear))) + 
  guides(color = guide_colorbar(title = "Horsepower"),
    shape = guide_legend(title = "Weight"), size = guide_legend(title = "Gear")
  )
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - section for title -->
<!--   - section for label -->
<!--   - section for legend bar -->
  