---
title: "ggplot2: Legends - Part 1"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

<style type="text/css">
    .reveal p img {
    align: middle;
  }
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: capitalize;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
    font-size: 25px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal p img {
    align: middle;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

Now that we have learnt to build different plots, let us look at different ways to modify the axis. Along the way, we will also explore the below functions:

- `scale_*()` family of functions 
- `guides()`
- `guide_legend()` 
- and `guide_colorbar()`


## Intro

<hr>

In this module, we will focus on modifying the legends of plots when the aesthetics are mapped to variables.

## Installation

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Alpha

<hr>

If you have mapped alpha to a continuous variable, use `scale_alpha_continuous()` to modify 

- title
- breaks
- limits 
- range
- and labels

## Scatter Plot

<hr>

```{r leg2, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = hp), color = 'blue')
```

## Title

<hr>

```{r leg3, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = hp), color = 'blue') +
  scale_alpha_continuous("Horsepower")
```

## Breaks

<hr>

```{r leg4, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = hp), color = 'blue') +
  scale_alpha_continuous(breaks = c(0, 100, 200, 300, 400))
```

## Limits

<hr>

```{r leg5, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = hp), color = 'blue') +
  scale_alpha_continuous(limits = c(0, 400))
```

## Range

<hr>

```{r leg6, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = hp), color = 'blue') +
  scale_alpha_continuous(range = c(0.4, 0.8))
```

## Labels

<hr>

```{r leg7, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = hp), color = 'blue') +
  scale_alpha_continuous(breaks = c(0, 100, 200, 300, 400),
    labels = c("Zero", "Hundred", "2 Hundred", 
      "3 Hundred", "4 Hundred"))
```

## Alpha - Continuous

<hr>

```{r leg8, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = hp), color = 'blue') +
  scale_alpha_continuous("Horsepower", breaks = c(0, 100, 200, 300, 400),
    limits = c(0, 400), range = c(0.4, 0.8),
    labels = c("Zero", "Hundred", "2 Hundred", 
      "3 Hundred", "4 Hundred"))
```

## Alpha Discrete

<hr>

If you have mapped alpha to a discrete variable, use `scale_alpha_discrete()` to modify

- title
- breaks
- limits 
- range
- and labels

## Scatter Plot

<hr>

```{r leg9, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue')
```

## Title

<hr>

```{r leg10, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue') +
  scale_alpha_discrete(name = "Cylinders")
```

## Breaks

<hr>

```{r leg11, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue') +
  scale_alpha_discrete(breaks = c("4", "6", "8"))
```

## Range

<hr>

```{r leg12, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue') +
  scale_alpha_discrete(range = c(0.4, 0.8))
```

## Labels

<hr>

```{r leg13, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue') +
  scale_alpha_discrete(labels = c("Four", "Six", "Eight"))
```

## Alpha Discrete

<hr>

```{r leg14, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue') +
  scale_alpha_discrete(name = "Cylinders", range = c(0.4, 0.8),
    breaks = c("4", "6", "8"), labels = c("Four", "Six", "Eight"))
```

## Color

<hr>

If you have mapped color to a discrete variable, use `scale_color_manual()` to modify 

- title
- values
- labels
- limits
- breaks

## Scatter Plot

<hr>

```{r leg15, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl)))
```

## Title

<hr>

```{r leg16, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(name = "Cylinders", 
    values = c("red", "blue", "green"))
```

## Values

<hr>

```{r leg17, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"))
```

## Labels

<hr>

```{r leg18, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    limits = c(4, 6, 8))
```

## Limits

<hr>

```{r leg19, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    labels = c('Four', 'Six', 'Eight'))
```

## Breaks

<hr>

```{r leg20, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    breaks = c(4, 6, 8))
```

## Color

<hr>

```{r leg21, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(name = "Cylinders", values = c("red", "blue", "green"),
    labels = c('Four', 'Six', 'Eight'), limits = c(4, 6, 8), breaks = c(4, 6, 8))
```

## Fill

<hr>

If you have mapped fill to a discrete variable, use `scale_fill_manual()` to modify 

- title
- values
- labels
- limits
- breaks

## Scatter Plot

<hr>

```{r leg22, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = factor(cyl)), shape = 22)
```

## Title

<hr>

```{r leg23, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = factor(cyl)), shape = 22) +
  scale_fill_manual(name = "Cylinders", 
    values = c("red", "blue", "green"))
```

## Values


<hr>

```{r leg24, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = factor(cyl)), shape = 22) +
  scale_fill_manual(values = c("red", "blue", "green"))
```

## Labels

<hr>

```{r leg25, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = factor(cyl)), shape = 22) +
  scale_fill_manual(values = c("red", "blue", "green"),
    labels = c('Four', 'Six', 'Eight'))
```

## Limits

<hr>

```{r leg26, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = factor(cyl)), shape = 22) +
  scale_fill_manual(values = c("red", "blue", "green"),
    limits = c(4, 6, 8))
```

## Breaks

<hr>

```{r leg27, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = factor(cyl)), shape = 22) +
  scale_fill_manual(values = c("red", "blue", "green"),
    breaks = c(4, 6, 8))
```

## Fill

<hr>

```{r leg28, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = factor(cyl)), shape = 22) +
  scale_fill_manual(name = "Cylinders", values = c("red", "blue", "green"),
    labels = c('Four', 'Six', 'Eight'), limits = c(4, 6, 8), breaks = c(4, 6, 8))
```

## Shape

<hr>

If you have mapped shape to a discrete variable, use `scale_shape_manual()` to modify 

- title
- values
- labels
- limits
- breaks

## Scatter Plot

<hr>

```{r leg29, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl)))
```

## Title

<hr>

```{r leg30, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape_manual(name = "Cylinders",values = c(4, 12, 24))
```

## Values

<hr>

```{r leg31, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape_manual(values = c(4, 12, 24))
```

## Labels

<hr>

```{r leg32, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape_manual(values = c(4, 12, 24), labels = c('Four', 'Six', 'Eight'))
```

## Limits

<hr>

```{r leg33, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape_manual(values = c(4, 12, 24), limits = c(4, 6, 8))
```

## Breaks

<hr>

```{r leg34, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape_manual(values = c(4, 12, 24), breaks = c(4, 6, 8))
```

## Shape

<hr>

```{r leg35, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape_manual(name = "Cylinders", labels = c('Four', 'Six', 'Eight'),  
     values = c(4, 12, 24), limits = c(4, 6, 8), breaks = c(4, 6, 8))
```

## Size

<hr>

If you have mapped size to a discrete variable, use `scale_size_manual()` to modify 

- title
- range
- limits
- breaks
- labels

## Scatter Plot

<hr>

```{r leg36, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp))
```

## Title

<hr>

```{r leg37, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp)) +
  scale_size_continuous(name = "Horsepower")
```

## Range

<hr>

```{r leg38, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp)) +
  scale_size_continuous(range = c(3, 6))
```

## Limits

<hr>

```{r leg39, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp)) +
  scale_size_continuous(limits = c(0, 400))
```

## Breaks

<hr>

```{r leg40, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp)) +
  scale_size_continuous(breaks = c(100, 200, 300, 400))
```

## Labels

<hr>

```{r leg41, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp)) +
  scale_size_continuous(breaks = c(100, 200, 300, 400),
    labels = c("Hundred", "2 Hundred", "3 Hundred", "4 Hundred"))
```

## Size

<hr>

```{r leg42, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp)) +
  scale_size_continuous(name = "Horsepower", range = c(3, 6), 
    limits = c(0, 400), breaks = c(100, 200, 300, 400),
    labels = c("Hundred", "2 Hundred", "3 Hundred", "4 Hundred"))
```

## Alpha

<hr>

If you have mapped alpha to a discrete variable, use `scale_alpha_manual()` to modify 

- title
- values
- labels
- limits
- breaks

## Scatter Plot

<hr>

```{r leg43, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue')
```

## Title

<hr>

```{r leg44, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue') +
  scale_alpha_manual(name = "Cylinders", values = c(0.4, 0.6, 0.8))
```

## Values

<hr>

```{r leg45, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue') +
  scale_alpha_manual(values = c(0.4, 0.6, 0.8))
```

## Labels

<hr>

```{r leg46, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue') +
  scale_alpha_manual(values = c(0.4, 0.6, 0.8), 
    labels = c('Four', 'Six', 'Eight'))
```

## Limits

<hr>

```{r leg47, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue') +
  scale_alpha_manual(values = c(0.4, 0.6, 0.8), limits = c(4, 6, 8))
```

## Breaks

<hr>

```{r leg48, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue') +
  scale_alpha_manual(values = c(0.4, 0.6, 0.8), breaks = c(4, 6, 8))
```

## Alpha

<hr>

```{r leg49, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, alpha = factor(cyl)), color = 'blue') +
  scale_alpha_manual(name = "Cylinders", labels = c('Four', 'Six', 'Eight'),
    values = c(0.4, 0.6, 0.8), limits = c(4, 6, 8), breaks = c(4, 6, 8))
```

## Shape

<hr>

`scale_shape()` maps discrete variable to six shapes. If you have mapped shape to a discrete variable, use `scale_shape()` to modify 

- title
- labels
- limits
- breaks
- type

## Scatter Plot

<hr>

```{r leg50, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape()
```

## Title

<hr>

```{r leg51, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape(name = 'Cylinders')
```

## Labels

<hr>

```{r leg52, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape(labels = c('Four', 'Six', 'Eight'))
```

## Limits

<hr>

```{r leg53, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape(limits = c(4, 6, 8))
```

## Breaks

<hr>

```{r leg54, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape(breaks = c(4, 6, 8))
```

## Types

<hr>

```{r leg55, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape(solid = FALSE)
```

## Shape

<hr>

```{r leg56, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl))) +
  scale_shape(solid = FALSE, name = 'Cylinders', limits = c(4, 6, 8),
    labels = c('Four', 'Six', 'Eight'), breaks = c(4, 6, 8))
```

## Size

<hr>

`scale_size()` maps discrete variable to six sizes. If you have mapped size to a discrete variable, use `scale_size()` to modify 

## Scatter Plot

<hr>

```{r leg57, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp))
```

## Scale Size

<hr>

```{r leg58, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp)) +
  scale_size(name = "Horsepower")
```

## Questions

<hr>

## Thank You

<hr>