---
title: "ggplot2: Histogram"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

<style type="text/css">
    .reveal p img {
    align: middle;
  }
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: capitalize;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
    font-size: 25px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal p img {
    align: middle;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

## Intro

<hr>

## Installation

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```


## Case Study

<hr>

We will continue to use the ecommerce data. You can download it from [here](https://github.com/rsquaredacademy/datasets/blob/master/web.csv) or import directly using the readr package.

## Data Dictionary

<hr>

```{r import2, echo=FALSE, eval=FALSE, results='hide', message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
```

```{r import, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

## Data Dictionary

<hr>

## Simple Histogram

<hr>

```{r hist2, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit))
```

----

Let us add some color to the histogram.

```{r hist3, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), fill = 'blue')
```

----

The default number of bins is 30. We can specify the number of bins using `bins`.

```{r hist4, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), bins = 7, fill = 'blue', color = 'black')
```

----

We can specify the width of the bins using `binwidth`.

```{r hist5, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), binwidth = 2, fill = 'blue', color = 'black')
```

----

Line type and size.

```{r hist6, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), binwidth = 2, 
                 fill = 'blue', color = 'black', alpha = 0.3,
                 linetype = 2, size = 1.25)
```

----

We can also map the color to a variable.


```{r hist7, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit, fill = device), bins = 7)
```

## Questions

<hr>

## Thank You

<hr>