---
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
    css: styles.css
---

## {data-background="thankyou.png"}

## Agenda

<hr>

- build histogram
- specify bins
- modify 
    - color
    - fill
    - alpha
    - bin width
    - line type
    - line size
- map aesthetics to variables

## Intro

<hr>

## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Data 

<hr>

```{r import2, eval=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
```

```{r import, echo=FALSE, eval=TRUE, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

## Data Dictionary

<hr>

- id: row id
- referrer: referrer website/search engine
- os: operating system
- browser: browser
- device: device used to visit the website
- n_pages: number of pages visited
- duration: time spent on the website (in seconds)
- repeat: frequency of visits
- country: country of origin
- purchase: whether visitor purchased
- order_value: order value of visitor (in dollars)


## Histogram

<hr>

```{r hist2, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit))
```

## Specify Bins

<hr>

```{r hist4, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), bins = 7)
```

## Fill

<hr>

```{r hist3, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), bins = 7, fill = 'blue')
```

## Alpha

<hr>

```{r hist11, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), bins = 7, fill = 'blue', alpha = 0.3)
```

## Color

<hr>

```{r hist8, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), bins = 7, fill = 'white', color = 'blue')
```

## Bins, Color & Fill

<hr>

```{r hist9, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), bins = 7, fill = 'blue', color = 'white')
```


## Bin Width

<hr>

```{r hist5, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), binwidth = 2, fill = 'blue', color = 'black')
```

## Line Type

<hr>

```{r hist6, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), bins = 5, fill = 'white', 
    color = 'blue', linetype = 3)
```

## Line Size

<hr>

```{r hist10, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), bins = 5, fill = 'white', 
    color = 'blue', size = 1.25)
```

## Map Fill to Variable

<hr>

```{r hist7, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit, fill = device), bins = 7)
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - add introduction -->
<!--   - list arguments with description -->