---
output: _output.yaml
---

## {data-background="gg_intro.png"}

## Agenda

<hr>

- understand the philosophy of Grammar of Graphics
- explore different aspects of ggplot2
- learn to build some of the basic plots regularly used for exploring data


## Introduction

<hr>

[ggplot2](http://ggplot2.tidyverse.org/index.html) is an alternative (and awesome) to base R for data visualization. It is based on
[The Grammar of Graphics](http://www.springer.com/in/book/9780387245447). In this post, we will understand the philosophy behind **ggplot2** and learn to build some of the most frequently used plots for visualizing data.


## Libraries

<hr>

```{r cat, eval=FALSE}
library(ggplot2)
```

```{r cat1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
```

## Grammar of Graphics

<hr>

Grammar of Graphics is a formal system for building plots. The core idea is that any plot can be uniquely described as a combination of

- a dataset
- a geom
- a set of mappings
- a statistic
- a position adjustment
- a coordinate system
- a faceting scheme

## Data

<hr>

```{r gg1, fig.align='center'}
ggplot(data = mtcars)
```

## Geom

<hr>

```{r gg2, fig.align='center'}
ggplot(data = mtcars) +
  geom_point(mapping = aes(x = disp, y = mpg))
```

## Mappings

<hr>

```{r gg3, fig.align='center'}
ggplot(data = mtcars) +
  geom_point(mapping = aes(x = disp, y = mpg))
```

## Stat

<hr>

```{r gg4, fig.align='center'}
ggplot(data = mtcars) +
  geom_point(mapping = aes(x = disp, y = mpg), stat = "identity")
```

## Position

<hr>

```{r gg5, fig.align='center'}
ggplot(data = mtcars) +
  geom_point(mapping = aes(x = disp, y = mpg), position = "identity")
```

## Coordinate System

<hr>

```{r gg6, fig.align='center'}
ggplot(data = mtcars) +
  geom_point(mapping = aes(x = disp, y = mpg), position = "identity") +
  coord_cartesian()
```

## Facet

<hr>

```{r gg7, fig.align='center'}
ggplot(data = mtcars) +
  geom_point(mapping = aes(x = disp, y = mpg), position = "identity") +
  coord_cartesian() + facet_grid(. ~ gear)
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - complete content -->
