---
title: "ggplot2: Legends - Part 1"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

<style type="text/css">
    .reveal p img {
    align: middle;
  }
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: capitalize;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
    font-size: 25px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal p img {
    align: middle;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

## Intro

<hr>

In this module, we will continue to explore different ways to modify/customize the legend of plots.

## Installation

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Guides

<hr>

----

```{r gui2}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    guide = guide_legend(title = "Cylinders", title.hjust = 0.5,
      title.position = "top", label.position = "right", nrow = 2,
      direction = "horizontal", label.hjust = 0.5, reverse = TRUE))
```

----

```{r gui3}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_discrete(guide = guide_legend( title = "Cylinders",
    title.hjust = 0.5, title.position = "top", label.position = "right",
    direction = "horizontal", label.hjust = 0.5, nrow = 2, reverse = TRUE))
```

----

```{r gui4}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = hp)) +
  scale_color_continuous(guide = guide_colorbar(
    title = "Horsepower",
    title.position = "top",
    title.vjust = 1,
    label.position = "bottom",
    label.vjust = 0.8,
    barwidth = 10,
    barheight = 3,
    nbins = 10,
    raster = FALSE,
    ticks = TRUE,
    draw.ulim = TRUE,
    draw.llim = FALSE,
    direction = "horizontal",
    reverse = TRUE
  ))
```

----

```{r gui5}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = hp), shape = 22) +
  scale_fill_continuous(guide = guide_colorbar(
    title = "Horsepower",
    title.position = "top",
    title.vjust = 1,
    label.position = "bottom",
    label.vjust = 0.8,
    barwidth = 10,
    barheight = 3,
    nbins = 10,
    raster = FALSE,
    ticks = TRUE,
    draw.ulim = TRUE,
    draw.llim = FALSE,
    direction = "horizontal",
    reverse = TRUE)
  )
```

----

```{r gui6}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = hp), shape = 22) +
  scale_fill_gradient2(guide = guide_colorbar(
    title = "Horsepower",
    title.position = "top",
    title.vjust = 1,
    label.position = "bottom",
    label.vjust = 0.8,
    barwidth = 10,
    barheight = 3,
    nbins = 10,
    raster = FALSE,
    ticks = TRUE,
    draw.ulim = TRUE,
    draw.llim = FALSE,
    direction = "horizontal",
    reverse = TRUE)
  ) 
```

----

```{r gui7}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = hp), shape = 22) +
  scale_fill_gradient(low = "white", high = "black", 
    guide = guide_colorbar(
      title = "Horsepower",
      title.position = "top",
      title.vjust = 1,
      label.position = "bottom",
      label.vjust = 0.8,
      barwidth = 10,
      barheight = 3,
      nbins = 10,
      raster = FALSE,
      ticks = TRUE,
      draw.ulim = TRUE,
      draw.llim = FALSE,
      direction = "horizontal",
      reverse = TRUE
    )
  ) 
```

----

```{r gui8}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, fill = hp), shape = 22) +
  scale_fill_gradientn(colors = terrain.colors(5), 
    guide = guide_colorbar(
      title = "Horsepower",
      title.position = "top",
      title.vjust = 1,
      label.position = "bottom",
      label.vjust = 0.8,
      barwidth = 10,
      barheight = 3,
      nbins = 10,
      raster = FALSE,
      ticks = TRUE,
      draw.ulim = TRUE,
      draw.llim = FALSE,
      direction = "horizontal",
      reverse = TRUE
    )
  ) 
```

----

```{r gui9}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = hp, size = qsec,  
    shape = factor(gear))) + 
  guides(color = "colorbar", shape = "legend", size = "legend")
```

----

```{r gui10}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = hp, size = wt, 
    shape = factor(gear))) + 
  guides(color = guide_colorbar(title = "Horsepower"),
    shape = guide_legend(title = "Weight"),
    size = guide_legend(title = "Gear"))
```

## Questions

<hr>

## Thank You

<hr>