---
title: "ggplot2: Line Graphs"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

<style type="text/css">
    .reveal p img {
    align: middle;
  }
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: none;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
    font-size: 25px;
  }
  .reveal h1 {
    font-size: 24px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal p img {
    align: middle;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

- build 
    - simple line chart
    - grouped line chart
- map aesthetics to variables
- modify line
    - color
    - type
    - size

## Intro

<hr>

Line charts are used to examine trend over time.

## Installation

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Case Study

<hr>

We will use a data set related to GDP growth rate. You can download it from [here](https://github.com/rsquaredacademy/datasets/blob/master/gdp.csv). It contains GDP (Gross Domestic Product) growth data for the BRICS (Brazil, Russia, India, China, South Africa) for the years 2000 to 2005.

## Data

<hr>

```{r imports2, eval=FALSE}
gdp <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/gdp.csv')
```

```{r import2, echo=FALSE, eval=TRUE, message=FALSE}
gdp <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/gdp.csv')
gdp
```

## Line Chart

<hr>

```{r line100, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line()
```

## Line Color

<hr>

```{r line1, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(color = 'blue')
```

## Line Type

<hr>

- 0 : blank
- 1 : solid
- 2 : dashed
- 3 : dotted
- 4 : dotdash
- 5 : longdash
- 6 : twodash

## Line Type

<hr>

```{r line2, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(linetype = 2)
```

## Line Type (Dashed)

<hr>

```{r line3, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(linetype = 'dashed')
```

## Line Size

<hr>

```{r line4, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(size = 2)
```

## Modify Data

<hr>

Now let us map the aesthetics to the variables. The data used in the above example cannot be used as we need a variable with country names. We will use `gather()` function from the `tidyr` package to reshape the data.

```{r line5, fig.align='center'}
gdp2 <- gdp %>% 
  select(year, growth, india, china) %>% 
  gather(key = country, value = gdp, -year)
gdp2
```

## Grouped Line Chart

<hr>

```{r line6, fig.align='center'}
ggplot(gdp2, aes(year, gdp, group = country)) +
  geom_line()
```

## Map Color to Country

<hr>

```{r line7, fig.align='center'}
ggplot(gdp2, aes(year, gdp, group = country)) +
  geom_line(aes(color = country))
```

## Map Line Type to Country

<hr>

```{r line8, fig.align='center'}
ggplot(gdp2, aes(year, gdp, group = country)) +
  geom_line(aes(linetype = country))
```

## Map Line Width to Country

<hr>

```{r line9, fig.align='center'}
ggplot(gdp2, aes(year, gdp, group = country)) +
  geom_line(aes(size = country))
```

## Questions

<hr>

## Thank You

<hr>

