---
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
    css: styles.css
---

## {data-background="thankyou.png"}

## Agenda

<hr>

- build 
    - simple line chart
    - grouped line chart
- map aesthetics to variables
- modify line
    - color
    - type
    - size

## Intro

<hr>

Line charts are used to examine trend over time.

## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Case Study

<hr>

We will use a data set related to GDP growth rate. You can download it from [here](https://github.com/rsquaredacademy/datasets/blob/master/gdp.csv). It contains GDP (Gross Domestic Product) growth data for the BRICS (Brazil, Russia, India, China, South Africa) for the years 2000 to 2005.

## Data

<hr>

```{r imports2, eval=FALSE}
gdp <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/gdp.csv')
```

```{r import2, echo=FALSE, eval=TRUE, message=FALSE}
gdp <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/gdp.csv')
gdp
```

## Line Chart

<hr>

```{r line100, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line()
```

## Line Color

<hr>

```{r line1, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(color = 'blue')
```

## Line Type

<hr>

- 0 : blank
- 1 : solid
- 2 : dashed
- 3 : dotted
- 4 : dotdash
- 5 : longdash
- 6 : twodash

## Line Type

<hr>

```{r line2, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(linetype = 2)
```

## Line Type (Dashed)

<hr>

```{r line3, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(linetype = 'dashed')
```

## Line Size

<hr>

```{r line4, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(size = 2)
```

## Modify Data

<hr>

Now let us map the aesthetics to the variables. The data used in the above example cannot be used as we need a variable with country names. We will use `gather()` function from the `tidyr` package to reshape the data.

```{r line5, fig.align='center'}
gdp2 <- gdp %>% 
  select(year, growth, india, china) %>% 
  gather(key = country, value = gdp, -year)
gdp2
```

## Grouped Line Chart

<hr>

```{r line6, fig.align='center'}
ggplot(gdp2, aes(year, gdp, group = country)) +
  geom_line()
```

## Map Color to Country

<hr>

```{r line7, fig.align='center'}
ggplot(gdp2, aes(year, gdp, group = country)) +
  geom_line(aes(color = country))
```

## Map Line Type to Country

<hr>

```{r line8, fig.align='center'}
ggplot(gdp2, aes(year, gdp, group = country)) +
  geom_line(aes(linetype = country))
```

## Map Line Width to Country

<hr>

```{r line9, fig.align='center'}
ggplot(gdp2, aes(year, gdp, group = country)) +
  geom_line(aes(size = country))
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - add introduction -->
<!--   - change case study data -->
<!--   - add section for grouped line chart -->
