---
output: _output.yaml
---

## {data-background="gg_line.png"}

## Agenda

<hr>

- build 
    - simple line chart
    - grouped line chart
- map aesthetics to variables
- modify line
    - color
    - type
    - size

## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(readr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(readr)
```

## Data

<hr>

```{r import, eval=FALSE}
users <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/users_device.csv',
  col_types = list(col_date(format = "%m/%d/%y"), col_integer(),
                   col_integer(), col_integer()))
```

```{r show2, echo=FALSE, eval=TRUE, message=FALSE}
library(readr)
users <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/users_device.csv',
  col_types = list(col_date(format = "%m/%d/%y"), col_integer(),
                   col_integer(), col_integer()))
users
```


## Data Dictionary

<hr>

Below is the description of the data set:

- Date: date 
- Desktop: visitors from a desktop device
- Mobile: visitors from a mobile device
- Tablet: visitors from a tablet device

## Line Chart

<hr>

```{r line100, fig.align='center'}
ggplot(users, aes(Date, Desktop)) +
  geom_line()
```

## Line Color

<hr>

```{r line1, fig.align='center'}
ggplot(users, aes(Date, Desktop)) +
  geom_line(color = 'blue')
```

## Line Type

<hr>

- 0 : blank
- 1 : solid
- 2 : dashed
- 3 : dotted
- 4 : dotdash
- 5 : longdash
- 6 : twodash

## Line Type

<hr>

```{r line2, fig.align='center'}
ggplot(users, aes(Date, Desktop)) +
  geom_line(linetype = 2)
```

## Line Type (Dashed)

<hr>

```{r line3, fig.align='center'}
ggplot(users, aes(Date, Desktop)) +
  geom_line(linetype = 'dashed')
```

## Line Size

<hr>

```{r line4, fig.align='center'}
ggplot(users, aes(Date, Desktop)) +
  geom_line(size = 2)
```

## Additional Lines

<hr>

```{r line10, fig.align='center'}
ggplot(users) +
  geom_line(aes(Date, Desktop)) +
  geom_line(aes(Date, Mobile)) +
  geom_line(aes(Date, Tablet)) 
```

## Data

<hr>

```{r import3, eval=FALSE}
tidy_users <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/tidy_users.csv',
  col_types = list(col_date(format = "%m/%d/%y"), col_character(), col_integer()))
```

```{r show3, echo=FALSE, eval=TRUE, message=FALSE}
library(readr)
tidy_users <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/tidy_users.csv',
  col_types = list(col_date(format = "%m/%d/%y"), col_character(), col_integer()))
tidy_users
```


## Data Dictionary

<hr>

Below is the description of the data set:

- Date: date 
- Device: device used by visitors
- Users: total daily visitors

## Grouped Line Chart

<hr>

```{r line6, fig.align='center'}
ggplot(tidy_users, aes(Date, Users, group = Device)) +
  geom_line()
```

## Map Color to Device

<hr>

```{r line7, fig.align='center'}
ggplot(tidy_users, aes(Date, Users, group = Device)) +
  geom_line(aes(color = Device))
```

## Map Line Type to Device

<hr>

```{r line8, fig.align='center'}
ggplot(tidy_users, aes(Date, Users, group = Device)) +
  geom_line(aes(linetype = Device))
```

## Map Line Width to Device

<hr>

```{r line9, fig.align='center'}
ggplot(tidy_users, aes(Date, Users, group = Device)) +
  geom_line(aes(size = Device))
```

## {data-background="thankyou.png"}

