---
title: "ggplot2: Positioning Geoms"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

<style type="text/css">
    .reveal p img {
    align: middle;
  }
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: capitalize;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
    font-size: 25px;
  }
  .reveal h1 {
    font-size: 24px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal p img {
    align: middle;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

## Intro

<hr>

In this module, we will learn to position the geoms to avoid overlapping.

## Installation

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Dodge

<hr>

Preserve the vertical position while adjusting the horizontal position.

Create grouped bar plots.

----

```{r pos2, fig.align='center'}
ggplot(data = mtcars, aes(factor(cyl), fill = factor(vs))) +
  geom_bar()
```

----

```{r pos3, fig.align='center'}
ggplot(data = mtcars, aes(factor(cyl), fill = factor(vs))) +
  geom_bar(position = position_dodge(width = 1))
```

## Jitter

<hr>

Add some noise to data points for better visualization.

----

```{r pos4, fig.align='center'}
ggplot(data = mtcars, aes(am, vs)) +
  geom_jitter()
```

----

```{r pos5, fig.align='center'}
ggplot(data = mtcars, aes(factor(am), factor(vs))) +
  geom_jitter()
```

----

```{r pos6, fig.align='center'}
ggplot(data = mtcars, aes(factor(am), factor(vs))) +
  geom_jitter(position = position_jitter(width = 0.1, height = 0.1))
```

----

```{r pos7, fig.align='center'}
ggplot(data = mtcars, aes(factor(am), factor(vs))) +
  geom_jitter(position = position_jitter(width = 0.05, height = 0.3))
```

----

```{r pos8, fig.align='center'}
ggplot(data = mtcars, aes(am, vs)) +
  geom_jitter(position = position_jitter(width = 0.1, height = 0.1))
```

## Nudge

<hr>

Use with geom_text. Useful for adjusting text around points.

----

```{r pos9}
df <- data.frame(
  x = c(1,3,2,5),
  y = c("a","c","d","c")
)
```

----

```{r pos12, fig.align='center'}
ggplot(df, aes(x, y)) +
  geom_point() +
  geom_text(aes(label = y),
            position = position_nudge(y = -0.1, x = -0.2))
```

## Fill

<hr>

Proportional bar plots.

----

```{r pos10, fig.align='center'}
ggplot(data = mtcars, aes(factor(cyl), fill = factor(gear))) +
  geom_bar(position = "fill")
```

## Stack

<hr>

```{r pos11}
series <- data.frame(
  time = c(rep(1, 4),rep(2, 4), rep(3, 4), rep(4, 4)),
  type = rep(c('a', 'b', 'c', 'd'), 4),
  value = rpois(16, 10)
)

series$type2 <- factor(series$type, levels = c('c', 'b', 'd', 'a'))
```

----

```{r pos13, fig.align='center'}
ggplot(series, aes(time, value, group = type)) +
  geom_line(aes(colour = type)) +
  geom_point(aes(colour = type))
```

----

```{r pos14, fig.align='center'}
ggplot(series, aes(time, value, group = type)) +
  geom_line(aes(colour = type), position = "stack") +
  geom_point(aes(colour = type), position = "stack")
```

## Questions

<hr>

## Thank You

<hr>