---
title: "ggplot2: Quick Tour"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

<style type="text/css">
    .reveal p img {
    align: middle;
  }
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: capitalize;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal p img {
    align: middle;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

## Intro

<hr>

In this module, we will learn to quickly build a set of plots that are routinely used to explore data:

- Scatter Plot
- Bar Plot
- Box Plot
- Line Chart
- Histogram

## Installation

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
```

## Scatter Plot

<hr>

```{r gg2, fig.align='center'}
qplot(disp, mpg, data = mtcars)
```

## Scatter Plot

<hr>

```{r gg3, fig.align='center'}
qplot(disp, mpg, data = mtcars, geom = c('point', 'line'))
```

## Scatter Plot

<hr>

```{r gg4, fig.align='center'}
qplot(disp, mpg, data = mtcars, color = factor(cyl))
```

## Scatter Plot

<hr>

```{r gg5, fig.align='center'}
qplot(disp, mpg, data = mtcars, shape = factor(cyl))
```

## Scatter Plot

<hr>

```{r gg6, fig.align='center'}
qplot(disp, mpg, data = mtcars, size = qsec)
```

## Bar Plot

<hr>

```{r gg7, fig.align='center'}
qplot(factor(cyl), data = mtcars, geom = c('bar'))
```

## Bar Plot

<hr>

```{r gg8, fig.align='center'}
qplot(factor(cyl), data = mtcars, geom = c('bar'), fill = factor(am))
```

## Box Plot

<hr>

```{r gg9, fig.align='center'}
qplot(factor(cyl), mpg, data = mtcars, geom = c('boxplot'))
```

## Box Plot

<hr>

```{r gg10, fig.align='center'}
qplot(factor(1), mpg, data = mtcars, geom = c('boxplot'))
```

## Box Plot

<hr>

```{r gg11, fig.align='center'}
qplot(factor(cyl), mpg, data = mtcars, geom = c('boxplot', 'jitter'))
```

## Line Chart

<hr>

```{r gg12, fig.align='center'}
qplot(x = date, y = unemploy, data = economics, geom = c('line'))
```

## Line Chart

<hr>

```{r gg13, fig.align='center'}
qplot(x = date, y = unemploy, data = economics, geom = c('line'),
      color = 'red')
```

## Histogram

<hr>

```{r gg14, fig.align='center'}
qplot(mpg, data = mtcars, bins = 5)
```

## Questions

<hr>

## Thank You

<hr>
