---
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
    css: styles.css
---

## {data-background="thankyou.png"}

## Agenda

<hr>

Learn to quickly build a set of plots that are routinely used to explore data:

- Scatter Plot
- Bar Plot
- Box Plot
- Line Chart
- Histogram

## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
```

## qplot

<hr>

Shortcut designed for those familiar with base plots. You can quickly produce a number of different
types of plots. Below are the key arguments:

- `x`    : data for X axis
- `y`    : data for Y axis
- `geom` : symbols to represent data
- `data` : a data frame or a tibble

## Scatter Plot

<hr>

```{r gg2, fig.align='center'}
qplot(disp, mpg, data = mtcars)
```

## Scatter Plot

<hr>

```{r gg3, fig.align='center'}
qplot(disp, mpg, data = mtcars, geom = c('point', 'line'))
```

## Scatter Plot

<hr>

```{r gg4, fig.align='center'}
qplot(disp, mpg, data = mtcars, color = factor(cyl))
```

## Scatter Plot

<hr>

```{r gg5, fig.align='center'}
qplot(disp, mpg, data = mtcars, shape = factor(cyl))
```

## Scatter Plot

<hr>

```{r gg6, fig.align='center'}
qplot(disp, mpg, data = mtcars, size = qsec)
```

## Bar Plot

<hr>

```{r gg7, fig.align='center'}
qplot(factor(cyl), data = mtcars, geom = c('bar'))
```

## Bar Plot

<hr>

```{r gg8, fig.align='center'}
qplot(factor(cyl), data = mtcars, geom = c('bar'), fill = factor(am))
```

## Box Plot

<hr>

```{r gg9, fig.align='center'}
qplot(factor(cyl), mpg, data = mtcars, geom = c('boxplot'))
```

## Box Plot

<hr>

```{r gg10, fig.align='center'}
qplot(factor(1), mpg, data = mtcars, geom = c('boxplot'))
```

## Box Plot

<hr>

```{r gg11, fig.align='center'}
qplot(factor(cyl), mpg, data = mtcars, geom = c('boxplot', 'jitter'))
```

## Line Chart

<hr>

```{r gg12, fig.align='center'}
qplot(x = date, y = unemploy, data = economics, geom = c('line'))
```

## Line Chart

<hr>

```{r gg13, fig.align='center'}
qplot(x = date, y = unemploy, data = economics, geom = c('line'),
      color = 'red')
```

## Histogram

<hr>

```{r gg14, fig.align='center'}
qplot(mpg, data = mtcars, bins = 5)
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->