---
title: "ggplot2: Quick Tour"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: moon
    highlight: monochrome
    center: true
    fig_width: 6
    fig_height: 5
    reveal_options:
      slideNumber: true
      previewLinks: true
---

## Intro

In this module, we will learn to quickly build a set of plots that are routinely used to explore data.

## Installation

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
```

## Scatter Plot

----

```{r gg2}
qplot(disp, mpg, data = mtcars)
```

----

```{r gg3}
qplot(disp, mpg, data = mtcars, geom = c('point', 'line'))
```

----

```{r gg4}
qplot(disp, mpg, data = mtcars, color = factor(cyl))
```

----

```{r gg5}
qplot(disp, mpg, data = mtcars, shape = factor(cyl))
```

----

```{r gg6}
qplot(disp, mpg, data = mtcars, size = qsec)
```

----

## Bar Plot

----

```{r gg7}
qplot(factor(cyl), data = mtcars, geom = c('bar'))
```

----

```{r gg8}
qplot(factor(cyl), data = mtcars, geom = c('bar'), fill = factor(am))
```

----

## Box Plot

----

```{r gg9}
qplot(factor(cyl), mpg, data = mtcars, geom = c('boxplot'))
```

----

```{r gg10}
qplot(factor(1), mpg, data = mtcars, geom = c('boxplot'))
```

----

```{r gg11}
qplot(factor(cyl), mpg, data = mtcars, geom = c('boxplot', 'jitter'))
```

## Line Chart

----

```{r gg12}
qplot(x = date, y = unemploy, data = economics, geom = c('line'))
```

----

```{r gg13}
qplot(x = date, y = unemploy, data = economics, geom = c('line'),
      color = 'red')
```

## Histogram

----

```{r gg14}
qplot(mpg, data = mtcars, bins = 5)
```

## Summary

In this module, we learnt to quickly create plots using the `qplot()` function. While useful, it has limitations and can be used only to quickly visualize data. 

## Next Steps..

In the next module, we will create the same set of plots but using **geoms**. 