---
output: _output.yaml
---

## {data-background="gg_scatter.png"}

## Agenda

<hr>

- build scatter plots
- modify point
    - color
    - fill
    - alpha
    - shape
    - size
- fit regression line

## Intro

<hr>

A scatter plot displays the relationship between two continuous variables. In ggplot2, we can build a scatter plot using `geom_point()`. Scatterplots can show you visually 

- the strength of the relationship between the variables 
- the direction of the relationship between the variables 
- and whether outliers exist


## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Basic Plot

<hr>

```{r scat2, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg))
```

## Jitter

<hr>

```{r scat22, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg), position = 'jitter')
```

## Jitter

<hr>

```{r scat3, fig.align='center'}
ggplot(mtcars) +
  geom_jitter(aes(disp, mpg))
```

## Aesthetics

<hr>

Now let us modify the aesthetics of the points. There are two ways:

- map them to variables using the `aes()` function
- specify values

## Map Color to Variable (Categorical)

<hr>

```{r scat4, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl)), 
             position = 'jitter')
```

## Map Color to Variable (Continuous)

<hr>

```{r scat5, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = hp), 
             position = 'jitter')
```

## Specify Value for Color

<hr>

```{r scat6, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg), color = 'blue', 
             position = 'jitter')
```

## Specify Value for Alpha

<hr>

```{r scat7, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg), color = 'blue', alpha = 0.4, 
             position = 'jitter')
```

## Map Shape to Variable

<hr>

```{r scat8, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl)),
             position = 'jitter')
```

## Specify Value for Shape

<hr>

```{r scat9, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg), shape = 3,
             position = 'jitter')
```

## Map Size to Variable

<hr>

```{r scat10, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp), color = 'blue',
             position = 'jitter')
```

## Specify Value for Size

<hr>

```{r scat11, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg), size = 3,
             position = 'jitter')
```

## Regression Line

<hr>

```{r aes12, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point(position = 'jitter') +
  geom_smooth(method = 'lm', se = FALSE)
```

## Regression Line - Conf. Interval

<hr>

```{r aes13, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point(position = 'jitter') +
  geom_smooth(method = 'lm', se = TRUE)
```

## Regression Line - Loess Method

<hr>

```{r aes14, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point(position = 'jitter') +
  geom_smooth(method = 'loess', se = FALSE)
```

## Fit Line - Intercept & Slope

<hr>

```{r aes17, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point(position = 'jitter') +
  geom_abline(intercept = 29.59985, slope = -0.04122)
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - add introduction -->
<!--   - section slide for aesthetics -->
<!--   - section for regression line -->
  