---
title: "ggplot2: Scatter Plots"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

<style type="text/css">
    .reveal p img {
    align: middle;
  }
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: capitalize;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
    font-size: 25px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal p img {
    align: middle;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

- build scatter plots
- map aesthetics to variables
- modify axis and legend
- fit regression line

## Intro

<hr>

## Installation

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Basic Plot

<hr>

```{r scat2, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg))
```

## Jitter

<hr>

```{r scat22, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg), position = 'jitter')
```

## Jitter

<hr>

```{r scat3, fig.align='center'}
ggplot(mtcars) +
  geom_jitter(aes(disp, mpg))
```

## Color

<hr>

Now let us modify the color of the points. There are two ways:

- map color to variable using the `aes()` function
- specify colors separately

## Color

<hr>

```{r scat4, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl)), 
             position = 'jitter')
```

## Color

<hr>

```{r scat5, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = hp), 
             position = 'jitter')
```

## Color

<hr>

```{r scat6, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg), color = 'blue', 
             position = 'jitter')
```

## Alpha

<hr>

```{r scat7, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg), color = 'blue', alpha = 0.4, 
             position = 'jitter')
```

## Shape

<hr>

```{r scat8, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl)),
             position = 'jitter')
```

## Shape

<hr>

```{r scat9, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg), shape = 3,
             position = 'jitter')
```

## Size

<hr>

```{r scat10, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp), color = 'blue',
             position = 'jitter')
```

## Size

<hr>

```{r scat11, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg), size = 3,
             position = 'jitter')
```

## Fit Line

<hr>

```{r aes12, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point(position = 'jitter') +
  geom_smooth(method = 'lm', se = FALSE)
```

## Confidence Interval

<hr>

```{r aes13, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point(position = 'jitter') +
  geom_smooth(method = 'lm', se = TRUE)
```

## Loess

<hr>

```{r aes14, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point(position = 'jitter') +
  geom_smooth(method = 'loess', se = FALSE)
```

## Fit Line

<hr>

If you use `geom_abline()`, specify the intercept and the slope.
Let us regress `duration` on `n_pages` and use the output.

```{r aes16}
lm(mpg ~ disp, data = mtcars)
```

----

```{r aes17, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point(position = 'jitter') +
  geom_abline(slope = 29.59985, intercept = -0.04122)
```

## Questions

<hr>

## Thank You

<hr>