---
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
    css: styles.css
---

## {data-background="thankyou.png"}

## Agenda

<hr>

Explore the following functions to build plots based on statistical transformation:

- `stat_count()`
- `stat_boxplot()`
- `stat_identity()`
- `stat_bin()`
- `stat_unique()`

## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Data

<hr>

```{r import2, eval=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
```

```{r import, echo=FALSE, eval=TRUE, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

## Data Dictionary

<hr>

- id: row id
- referrer: referrer website/search engine
- os: operating system
- browser: browser
- device: device used to visit the website
- n_pages: number of pages visited
- duration: time spent on the website (in seconds)
- repeat: frequency of visits
- country: country of origin
- purchase: whether visitor purchased
- order_value: order value of visitor (in dollars)


## Intro

<hr>

So far we have used `geom_*()` functions to build plots. Now we will use a new set of functions `stat_*()`: 

- `stat_count()`
- `stat_boxplot()`
- `stat_identity()`
- `stat_bin()`
- `stat_unique()`


## Bar Plots

<hr>

```{r stat2, fig.align='center'}
ggplot(ecom) +
  stat_count(aes(device))
```

## Box Plots

<hr>

```{r stat3, fig.align='center'}
ggplot(ecom) +
  stat_boxplot(aes(factor(1), n_visit))
```

## Scatter Plots

<hr>

```{r stat4, fig.align='center'}
ggplot(ecom) +
  stat_identity(aes(n_pages, duration), position = 'jitter')
```

## Histogram

<hr>

```{r stat5, fig.align='center'}
ggplot(ecom) +
  stat_bin(aes(n_visit), bins = 7, fill = 'blue', alpha = 0.3,
           color = 'black')
```

## Unique Values

<hr>

```{r stat6, fig.align='center'}
ggplot(ecom) +
  stat_unique(aes(n_pages, duration), 
              position = 'jitter')
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - change data for scatter plot -->