---
title: "ggplot2: Text Annotations"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

<style type="text/css">
    .reveal p img {
    align: middle;
  }
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: none;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
    font-size: 25px;
  }
  .reveal h1 {
    font-size: 24px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal p img {
    align: middle;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

- add custom text 
- modify color
- modify size
- modify fontface
- modify angle


## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
```

## Annotate

<hr>

Use `annotate()` to add custom text to a plot. It takes the following arguments:

- `geom`     : specify text
- `x`        : x axis location
- `y`        : y axis location
- `label`    : custom text 
- `color`    : color of text
- `size`     : size of text
- `fontface` : fontface of text
- `angle`    : angle of text

## Add Text

<hr>

```{r ann2, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text')
```

## Color

<hr>

```{r ann4, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text', color = 'red')
```

## Size

<hr>

```{r ann5, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text', size = 6)
```

## Font

<hr>

```{r ann6, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text', fontface = 'bold')
```

## Angle

<hr>

```{r ann7, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text', angle = 25)
```

## Annotate

<hr>

```{r ann3, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text',
           color = 'red', size = 6, fontface = 'bold', angle = 25)
```

## Questions

<hr>

## Thank You

<hr>