---
output: _output.yaml
---

## {data-background="gg_text_ann.png"}

## Agenda

<hr>

- add custom text 
- modify color
- modify size
- modify fontface
- modify angle


## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
```

## Annotate

<hr>

Use `annotate()` to add custom text to a plot. It takes the following arguments:

- `geom`     : specify text
- `x`        : x axis location
- `y`        : y axis location
- `label`    : custom text 
- `color`    : color of text
- `size`     : size of text
- `fontface` : fontface of text
- `angle`    : angle of text

## Add Text

<hr>

```{r ann2, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text')
```

## Color

<hr>

```{r ann4, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text', color = 'red')
```

## Size

<hr>

```{r ann5, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text', size = 6)
```

## Font

<hr>

```{r ann6, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text', fontface = 'bold')
```

## Angle

<hr>

```{r ann7, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text', angle = 25)
```

## Annotate

<hr>

```{r ann3, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text',
           color = 'red', size = 6, fontface = 'bold', angle = 25)
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - image for arguments -->