---
title: "ggplot2: Text Annotations"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: moon
    highlight: monochrome
    center: true
    fig_width: 6
    fig_height: 5
    reveal_options:
      slideNumber: true
      previewLinks: true
---

## Intro

## Installation

```{r show, eval=FALSE}
library(ggplot2)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
```

## Annotate

We will use the `annotate()` function to add custom text to the plots. You can use the `annotate()` function to add rectangles/segments/pointrange as well but our focus will be on adding text. Let us start with a simple scatter plot.

----

```{r ann2}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text')
```

----

We can format the text as well.

```{r ann3}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text',
           color = 'red', size = 6, fontface = 'bold', angle = 25)
```

## Summary

## Next Steps..

In the next module, we will learn how to work with guides/legends in ggplot2.
