---
output: _output.yaml
---

## {data-background="gg_themes.png"}

## Agenda

<hr>

In this module, we will learn to 

- modify the appearance of all non data components of the plot such as
    - axis
    - legend
    - panel
    - plot area
    - background
    - margin
    - facets
- use different themes 


## Libraries

<hr>

```{r show, eval=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Scatter Plot

<hr>

```{r theme1, fig.align='center'}
p <- ggplot(mtcars) +
  geom_point(aes(disp, mpg))
p
```

## Axis Title

<hr>

```{r theme2, fig.align='center'}
p + theme(axis.title.x = element_text(color = "red", size = 10))
```

## Axis Text

<hr>

```{r theme3, fig.align='center'}
p + theme(axis.text.x = element_text(color = "red", size = 10))
```

## Axis Font

<hr>


```{r theme4, fig.align='center'}
p + theme(axis.text = element_text(color = "red", size = 10, face = "italic"))
```

## Axis Ticks

<hr>

```{r theme5, fig.align='center'}
p + theme(axis.ticks = element_line(color = 'blue', size = 1.25, linetype = 2), 
          axis.ticks.length = unit(1, "cm"))
```

## Axis Line

<hr>

```{r theme6, fig.align='center'}
p + theme(axis.line = element_line(color = 'red', size = 1.5, linetype = 3))
```

## Legend

<hr>

```{r theme7, fig.align='center'}
p <- ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl), shape = factor(gear)))
p
```

## Background 

<hr>

```{r theme30, fig.align='center'}
p + theme(legend.background = element_rect(fill = 'gray', linetype = 3,  
          color = "black"))
```

## Margin 

<hr>

```{r theme31, fig.align='center'}
p + theme(legend.margin = margin(4, 8, 4, 8))
```

## Spacing 

<hr>

```{r theme32, fig.align='center'}
p + theme(legend.spacing = unit(1.5, "cm"))
```

## Key 

<hr>

```{r theme33, fig.align='center'}
p + theme(legend.key = element_rect(fill = 'red'),
          legend.key.size = unit(0.4, "cm"))
```

## Text 

<hr>

```{r theme34, fig.align='center'}
p + theme(legend.text = element_text(color = 'green', face = 'italic'))
```

## Title 

<hr>

```{r theme35, fig.align='center'}
p + theme(legend.title = element_text(color = 'blue', face = 'bold'),
          legend.title.align = 0.1)
```

## Position 

<hr>

```{r theme36, fig.align='center'}
p + theme(legend.position = "top")
```

## Direction 

<hr>

```{r theme37, fig.align='center'}
p + theme(legend.direction = "horizontal")
```

## Box 

<hr>

```{r theme38, fig.align='center'}
p + theme(legend.box = "horizontal",
  legend.box.background = element_rect(fill = "gray"),
  legend.box.spacing = unit(1, "cm"), legend.box.just = "left",
  legend.box.margin = margin(4, 8, 4, 8))
```

## Panel

<hr>

```{r theme9, fig.align='center'}
p <- ggplot(mtcars) +
  geom_point(aes(disp, mpg))
p
```

## Panel Background

<hr>

```{r theme10, fig.align='center'}
p + theme(panel.background = element_rect(fill = 'gray'))
```

## Panel Border

<hr>

```{r theme24, fig.align='center'}
p + theme(panel.border = element_rect(fill = NA, color = 'blue', 
          linetype = 1, size = 2))
```

## Panel Grid

<hr>

```{r theme25, fig.align='center'}
p + theme(panel.grid = element_line(color = 'blue', linetype = 2, size = 0.5))
```

## Panel Grid (Major)

<hr>

```{r theme12, fig.align='center'}
p + theme(panel.grid.major.x = element_line(color = 'blue', 
          linetype = 2, size = 0.5))
```


## Panel Grid (Minor)

<hr>

```{r theme26, fig.align='center'}
p + theme(panel.grid.minor.y = element_line(color = 'red', 
          linetype = 3, size = 0.2))
```


## Plot Area

<hr>

```{r theme13, fig.align='center'}
p <- ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + 
  ggtitle('Theme Modification')
p
```

## Plot Background

<hr>

```{r theme14, fig.align='center'}
p + theme(plot.background = element_rect(color = 'blue', 
          fill = '#4682B4'))
```

## Plot Title

<hr>

```{r theme27, fig.align='center'}
p + theme(plot.title = element_text(color = 'red'))
```

## Plot Margin

<hr>

```{r theme28, fig.align='center'}
p + theme(plot.margin = unit(c(1, 1, 1, 1), "cm"))
```

## Facets

<hr>

```{r theme15, fig.align='center'}
p <- ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + 
  facet_grid(~cyl)
p
```

## Strip Background

<hr>

```{r theme16, fig.align='center'}
p + theme(strip.background = element_rect(color = 'red', 
          fill = '#4682B4'))
```

## Strip Text

<hr>

```{r theme29, fig.align='center'}
p + theme(strip.text = element_text(face = 'italic', size = 8,
          color = 'red'))
```

## Themes

<hr>

## Classic Dark on Light

<hr>

```{r theme17, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  theme_bw()
```

## Default Gray

<hr>

```{r theme18, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  theme_gray()
```

## Light

<hr>

```{r theme19, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  theme_light()
```

## Minimal

<hr>

```{r theme20, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  theme_minimal()
```

## Dark

<hr>

```{r theme21, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  theme_dark()
```

## Classic

<hr>

```{r theme22, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  theme_classic()
```

## Void (Empty)

<hr>

```{r theme23, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  theme_void()
```

## {data-background="thankyou.png"}

To do:
  - add intro slide
  - section for axis, legend, panel, plot area, facets and themes