---
output: _output.yaml
---

## {data-background="github_pkginfo.png"}

## Agenda

<hr>

- Introduction
- GitHub
- Build Status
- Code Coverage
- Stack Overflow
- CRAN 

## 

<br>
<br>
<br>
<br>
<br>
<br>
<br>

`pkginfo` will allow users to retrieve information about a R package from [CRAN](https://cran.r-project.org/), [GitHub](https://github.com/), [Travis CI](https://travis-ci.org/), [Appveyor](https://ci.appveyor.com/), [Codecov](https://codecov.io/) & [Stack Overflow](https://stackoverflow.com/). While it was not developed keeping any specific end user in mind, package developers might find it more useful as they can monitor their R packages using a shiny app. `pkginfo` uses [cranlogs](https://cranlogs.r-pkg.org/),[crandb](https://github.com/r-hub/crandb) and [cchecksapi](https://github.com/ropenscilabs/cchecksapi) under the hood.

## Installation

<hr>

```{r install, eval=FALSE}
devtools::install_github("rsquaredacademy/pkginfo")
```

```{r load, echo=FALSE}
library(pkginfo)
```

## {data-background="section_github.png"}

## Username

<hr>

```{r username}
get_gh_username("dplyr")
```

## Stars, Forks & Issues

<hr>

```{r stats}
get_gh_stats("dplyr")
```

## Branches

<hr>

```{r branches}
get_gh_branches("dplyr")
```

## Issues

<hr>

```{r issues}
get_gh_issues("dplyr")
```

## Labels

<hr>

```{r labels}
get_gh_labels("dplyr")
```

## Milestones

<hr>

```{r milestones}
get_gh_milestones("dplyr")
```

## Pull Requests

<hr>

```{r pr}
get_gh_pr("dplyr")
```

## Releases

<hr>

```{r releases}
get_gh_releases("dplyr")
```

## {data-background="section_build.png"}

## Travis

<hr>

```{r travis}
get_status_travis("dplyr", "tidyverse")
```

## Appvetor

<hr>

```{r appveyor}
get_status_appveyor("dplyr", "tidyverse")
```

## Codecov

<hr>

```{r codecov}
get_code_coverage("dplyr", "tidyverse")
```

## {data-background="section_stack_overflow.png"}

## Questions

<hr>

```{r so}
get_so_questions("dplyr")
```

## {data-background="section_cran.png"}

## CRAN Check Results

<hr>

```{r cran_check}
get_pkg_cran_check_results("dplyr")
```

## Downloads

<hr>

```{r cran_downloads}
get_pkg_downloads("dplyr")
```

## {data-background="thankyou.png"}

