---
output: _output.yaml 
---

## {data-background="tibble_intro.png"}

## Agenda

<hr>

- what are tibbles?
- how are tibbles different from data frames?
- how to create tibbles?
- how to manipulate tibbles?

## {data-background="tibbles1.png"}

## What are tibbles?

<hr>

> A **tibble**, or `tbl_df`, is a modern reimagining of the data.frame, keeping what time has proven to be effective, and throwing out what is not. Tibbles are data.frames that are lazy and surly: they do less (i.e. they don’t change variable names or types, and don’t do partial matching) and complain more (e.g. when a variable does not exist). This forces you to confront problems earlier, typically leading to cleaner, more expressive code. Tibbles also have an enhanced `print method()` which makes them easier to use with large datasets containing complex objects.

> Source: http://tibble.tidyverse.org/

## {data-background="creatingtibbles.png"}

## Creating tibbles

<hr>

```{r tibble1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(tibble)
library(dplyr)
```

```{r tibble2}
tibble(x = letters,
       y = 1:26,
       z = sample(100, 26))
```

## {data-background="tibblefeatures.png"}

## tibble features

<hr>

- never changes input's types
- never adjusts variable names
- never prints all rows
- never recycles vector of length greater than 1

## Never changes input's types

<hr>

```{r tibble3}
tibble(x = letters,
       y = 1:26,
       z = sample(100, 26))
```

## Never changes input's types

<hr>

```{r tibble3a}
data <- data.frame(x = letters, y = 1:26, z = sample(100, 26))
str(data)
```

## Never adjusts variable names

<hr>

```{r tibble4}
names(data.frame(`order value` = 10))
names(tibble(`order value` = 10))
```

## Never prints all rows

<hr>

```{r tibble5}
x <- 1:100
y <- letters[1]
z <- sample(c(TRUE, FALSE), 100, replace = TRUE)
tibble(x, y, z)
```

## Never recycle vector of length greater than 1

<hr>

```{r tibble6, eval=FALSE}
x <- 1:100
y <- letters
z <- sample(c(TRUE, FALSE), 100, replace = TRUE)
tibble(x, y, z)
Error in overscope_eval_next(overscope, expr) : object 'y' not found
```

## {data-background="enframe.png"}

## Atomic Vectors

<hr>

```{r tibble12}
browsers <- c('chrome', 'safari', 'firefox', 'edge')
enframe(browsers)
```

## Atomic Vectors

<hr>

```{r tibble13}
browsers <- c(chrome = 40, firefox = 20, edge = 30, safari = 10)
enframe(browsers)
```

## {data-background="tribble.png"}

## Tribble

<hr>

Another way to create tibbles is using `tribble()`: 

- it is short for transposed tibbles
- it is customized for data entry in code 
- column names start with `~` 
- and values are separated by commas

## Tribble

<hr>

```{r tibble15}
tribble(
  ~x, ~y, ~z,
  #--|--|----
  1, TRUE, 'a',
  2, FALSE, 'b'
)
```

## Column Names

<hr>

Names of the columns in tibbles need not be valid R variable names. They can contain unusual characters like a space or a smiley but must be enclosed in ticks. 

```{r tibble16}
tibble(
  ` `  = 'space',
  `2`  = 'integer',
  `:)` = 'smiley'
)
```

## {data-background="manipulatetibble.png"}


## Add Rows

<hr>

```{r tibble21}
browsers <- enframe(c(chrome = 40, firefox = 20, edge = 30))
browsers
```

## Add Rows

<hr>

```{r tibble22}
add_row(browsers, name = 'safari', value = 10)
```

## Add Rows

<hr>

```{r tibble23}
add_row(browsers, name = 'safari', value = 10, .before = 2)
```

## Add Column

<hr>

```{r tibble24}
browsers <- enframe(c(chrome = 40, firefox = 20, edge = 30, safari = 10))
add_column(browsers, visits = c(4000, 2000, 3000, 1000))
```

## Remove Rownames

<hr>

```{r tibble26}
remove_rownames(mtcars)
```

## Rownames to Column

<hr>

```{r tibble27}
head(rownames_to_column(mtcars))
```

## Column to Rownames

<hr>

```{r tibble28}
mtcars_tbl <- rownames_to_column(mtcars)
column_to_rownames(mtcars_tbl)
```

## Glimpse

<hr>

```{r tibble 29}
glimpse(mtcars)
```

## Membership Testing

<hr>

```{r tibble14}
is_tibble(mtcars)
is_tibble(as_tibble(mtcars))
```

## Rownames

<hr>

```{r tibble25}
has_rownames(mtcars)
```

## Check Column 

<hr>

```{r tibble30}
has_name(mtcars, 'cyl')
has_name(mtcars, 'gears')
```

## {data-background="tibblesummary.png"}

## Summary

<hr>

- use `tibble()` to create tibbles
- use `as_tibble()` to coerce other objects to tibble
- use `enframe()` to coerce vector to tibble
- use `tribble()` to create tibble using data entry

## Summary

<hr>

- use `add_row()` to add a new row
- use `add_column()` to add a new column
- use `remove_rownames()` to remove rownames from data
- use `rownames_to_column()` to coerce rowname to first column
- use `column_to_rownames()` to coerce first column to rownames

## Summary

<hr>

- use `is_tibble()` to test if an object is a tibble
- use `has_rownames()` to check whether a data set has rownames
- use `has_name()` to check if tibble has a specific column 
- use `glimpse()` to get an overview of data

## {data-background="thankyou.png"}
