---
output: _output.yaml 
---

## {data-background="tibble_intro.png"}

## Agenda

<hr>

- What are tibbles?
- How are tibbles different from data frames?
- How to create tibbles?
- How to manipulate tibbles?

## {data-background="tibbles1.png"}

## What are tibbles?

<hr>

## {data-background="creatingtibbles.png"}

## Creating tibbles

<hr>

tibble can be created using any of the following:

- `tibble()`
- `as_tibble()`
- `tribble()`


## Creating tibbles

<hr>

```{r tibble1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(tibble)
library(dplyr)
```

```{r tibble2}
tibble(x = letters,
       y = 1:26,
       z = sample(100, 26))
```

## {data-background="tibblefeatures.png"}

## tibble features

<hr>

- never changes input's types
- never adjusts variable names
- never prints all rows
- never recycles vector of length greater than 1

## never changes input's types

<hr>

```{r tibble3}
tibble(x = letters,
       y = 1:26,
       z = sample(100, 26))
```

## never adjusts variable names

<hr>

```{r tibble4}
names(data.frame(`order value` = 10))
names(tibble(`order value` = 10))
```

## never prints all rows

<hr>

```{r tibble5}
x <- 1:100
y <- letters[1]
z <- sample(c(TRUE, FALSE), 100, replace = TRUE)
tibble(x, y, z)
```

## never recycle vector of length greater than 1

<hr>

```{r tibble6, eval=FALSE}
x <- 1:100
y <- letters
z <- sample(c(TRUE, FALSE), 100, replace = TRUE)
tibble(x, y, z)
Error in overscope_eval_next(overscope, expr) : object 'y' not found
```

## {data-background="creatingtibbles2.png"}

## Creating tibbles II

<hr>

In the previous section, we created tibbles using `tibble()` by directly entering data as arguments. In this section, we will convert the following objects to tibble using `as_tibble()`:

- data.frame
- list
- matrix
- table
- tbl_df
- atomic vectors

## DataFrames

<hr>

```{r tibble7}
as_tibble(iris)
```

## Lists

<hr>

```{r tibble8}
l <- list(x = 1:32, y = sample(100, 32), z = replicate(32, letters, simplify = FALSE))
as_tibble(l)
```

## Matrix

<hr>

```{r tibble9}
m <- matrix(sample(1000, 2600, replace = TRUE), ncol = 26)
colnames(m) <- letters
as_tibble(m)
```

## Table

<hr>

```{r tibble10}
k <- table(mtcars$cyl)
as_tibble(k)
```

## tbl_df

<hr>

```{r tibble11}
k <- tbl_df(iris)
as_tibble(k)
```

## {data-background="enframe.png"}

## atomic vectors

<hr>

```{r tibble12}
browsers <- c('chrome', 'safari', 'firefox', 'edge')
enframe(browsers)
```

## atomic vectors

<hr>

```{r tibble13}
browsers <- c(chrome = 40, firefox = 20, edge = 30, safari = 10)
enframe(browsers)
```

## Membership Testing

<hr>

We can test if an object is a tibble using `is_tibble()`.

```{r tibble14}
is_tibble(mtcars)
is_tibble(as_tibble(mtcars))
```

## {data-background="tribble.png"}

## Tribble

<hr>

Another way to create tibbles is using `tribble()`: 

- it is short for transposed tibbles
- it is customized for data entry in code 
- column names start with `~` 
- and values are separated by commas

## Tribble

<hr>

```{r tibble15}
tribble(
  ~x, ~y, ~z,
  #--|--|----
  1, TRUE, 'a',
  2, FALSE, 'b'
)
```

## Column Names

<hr>

Names of the columns in tibbles need not be valid R variable names. They can contain unusual characters like a space or a smiley but must be enclosed in ticks. 

```{r tibble16}
tibble(
  ` ` = 'space',
  `2` = 'integer',
  `:)` = 'smiley'
)
```

## {data-background="manipulatetibble.png"}


## Add Rows

<hr>

Let us add data related to **Safari** browser to the web traffic data using `add_row()`.

```{r tibble21}
browsers <- enframe(c(chrome = 40, firefox = 20, edge = 30))
browsers
```

```{r tibble22}
add_row(browsers, name = 'safari', value = 10)
```

## Add Rows

<hr>

If we want to add the data at a particular row, we can specify the row number using the `.before` argument. Let us add the data related to **Safari** browser in the second row instead of the last row.

```{r tibble23}
add_row(browsers, name = 'safari', value = 10, .before = 2)
```

## Add Column

<hr>

`add_column()` adds a new column to tibbles.

```{r tibble24}
browsers <- enframe(c(chrome = 40, firefox = 20, edge = 30, safari = 10))
add_column(browsers, visits = c(4000, 2000, 3000, 1000))
```

## Rownames

<hr>

The [tibble](tibble.tidyverse.org) package provides a set of functions to deal with rownames. Remember, `tibble` does not have `rownames` unlike `data.frame`. To check whether a data set has rownames, use `has_rownames()`.

```{r tibble25}
has_rownames(mtcars)
```

## Remove Rownames

<hr>

```{r tibble26}
remove_rownames(mtcars)
```

## Rownames to Column

<hr>

```{r tibble27}
head(rownames_to_column(mtcars))
```

## Column to Rownames

<hr>

To convert the first column in the data set to rownames, use `column_to_rownames()`:

```{r tibble28}
mtcars_tbl <- rownames_to_column(mtcars)
column_to_rownames(mtcars_tbl)
```

## Glimpse

<hr>

Use `glimpse()` to get an overview of the data.

```{r tibble 29}
glimpse(mtcars)
```

## Check Column 

<hr>

`has_name()` can be used to check if a tibble has a specific column.

```{r tibble30}
has_name(mtcars, 'cyl')
has_name(mtcars, 'gears')
```

## {data-background="tibblesummary.png"}

## Summary

<hr>

- use `tibble()` to create tibbles
- use `as_tibble()` to coerce other objects to tibble
- use `enframe()` to coerce vector to tibble
- use `tribble()` to create tibble using data entry

## Summary

<hr>

- use `add_row()` to add a new row
- use `add_column()` to add a new column
- use `remove_rownames()` to remove rownames from data
- use `rownames_to_colum()` to coerce rowname to first column
- use `column_to_rownames()` to coerce first column to rownames

## Summary

<hr>

- use `is_tibble()` to test if an object is a tibble
- use `has_rownames()` to check whether a data set has rownames
- use `has_name()` to check if tibble has a specific column 
- use `glimpse()` to get an overview of data

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - what are tibbles -->