## extract domain name from email ids

<hr>

```{r str15}
emails <- mockstring %>%
  pull(email) %>%
  head
emails
```

## extract domain name from email ids

<hr>

```{r str16}
emails %>%
  str_split(pattern = '@')
```

## extract domain name from email ids

<hr>

```{r str17}
emails %>%
  str_split(pattern = '@') %>%
  map_chr(2)
```

## extract domain name from email ids

<hr>

```{r str18}
emails %>%
  str_split(pattern = '@') %>%
  map_chr(2) %>%
  str_split(pattern = '\\.') 
```

## extract domain name from email ids

<hr>

```{r str19}
emails %>%
  str_split(pattern = '@') %>%
  map_chr(2) %>%
  str_split(pattern = '\\.') %>%
  map_chr(extract(1))
```

## extract domain name from email ids

<hr>

```{r str20}
emails %>%
  str_split(pattern = '@') %>%
  map_chr(2) %>%
  str_split(pattern = '\\.', simplify = TRUE) %>%
  extract(, 2)
```

## Extract image type from URL

<hr>

```{r str21}
img <- mockstring %>%
  pull(imageurl) %>%
  head
img
```

## Extract image type from URL

<hr>

```{r str22}
img %>%
  str_split(pattern = '\\.')
```

## Extract image type from URL

<hr>

```{r str23}
img %>%
  str_split(pattern = '\\.') %>%
  map_chr(extract(3))
```

## Extract image type from URL

<hr>

```{r str24}
img %>%
  str_split(pattern = '\\.') %>%
  map_chr(extract(3)) %>%
  str_sub(start = 1, end = 3)
``` 

## Extract image type from URL

<hr> 

```{r str25}
img %>%
  str_split(pattern = '\\.') %>%
  map_chr(extract(3)) %>%
  str_split(pattern = '/') %>%
  map_chr(extract(1))
```

## Extract Image Dimesion from URL

<hr>

```{r str26}
img %>%
  str_locate(pattern = "[0-9]") 
```

## Extract Image Dimesion from URL

<hr>

```{r str27}
img %>%
  str_sub(start = 23) 
```

## Extract Image Dimesion from URL

<hr>

```{r str28}
img %>%
  str_sub(start = 23) %>%
  str_split(pattern = '\\.') 
```

## Extract Image Dimesion from URL

<hr>

```{r str29}
img %>%
  str_sub(start = 23) %>%
  str_split(pattern = '\\.') %>%
  map_chr(extract(1))
```
  
## Extract HTTP Protocol from URL

<hr>

```{r str30}
url1 <- mockstring %>%
  pull(url) %>%
  first
url1
```

## Extract HTTP Protocol from URL

<hr>

```{r str31}
url1 %>%
  str_split(pattern = '://') 
```

## Extract HTTP Protocol from URL

<hr>

```{r str32}
url1 %>%
  str_split(pattern = '://') %>%
  map_chr(extract(1))
```

## Extract HTTP Protocol from URL

<hr>

```{r str33}
url1 %>%
  str_locate(pattern = '://') 
```

## Extract HTTP Protocol from URL

<hr>

```{r str34}
url1 %>%
  str_locate(pattern = '://') %>%
  data.frame() 
```

## Extract HTTP Protocol from URL

<hr>

```{r str35}
url1 %>%
  str_locate(pattern = '://') %>%
  data.frame() %>%
  extract2('start') 
```

## Extract HTTP Protocol from URL

<hr>

```{r str36}
url1 %>%
  str_locate(pattern = '://') %>%
  data.frame() %>%
  extract2('start') %>%
  subtract(1)
```

## Extract HTTP Protocol from URL

<hr>

```{r str37}
urls <- mockstring %>%
  pull(url) %>%
  head
```

## Extract HTTP Protocol from URL

<hr>

```{r str38}
k <- urls %>%
  str_locate(pattern = '://') %>%
  data.frame() %>%
  extract2('start') %>%
  subtract(1)
```

## Extract HTTP Protocol from URL

<hr>

```{r str39}
urls %>%
  str_sub(start = 1, end = k)
```

## Extract Domain Name

<hr> 

```{r str40}
url1 %>%
  str_locate_all(pattern = "/") 
```

## Extract Domain Name

<hr> 

```{r str41}
url1 %>%
  str_locate_all(pattern = "/") %>%
  map_int(extract(3))
```

## Extract Domain Name

<hr> 

```{r str42}
n <- url1 %>%
  str_locate_all(pattern = "/") %>%
  map_int(extract(3))
```

## Extract Domain Name

<hr> 

```{r str43}
urls %>%
  str_sub(end = n) 
```

## Extract Domain Name

<hr> 

```{r str44}
urls %>%
  str_sub(end = n) %>%
  str_split(pattern = '\\.') 
```

## Extract Domain Name

<hr> 

```{r str45}
urls %>%
  str_sub(end = n) %>%
  str_split(pattern = '\\.') %>%
  map_chr(extract(1)) 
```

## Extract Domain Name

<hr> 

```{r str46}
urls %>%
  str_sub(end = n) %>%
  str_split(pattern = '\\.') %>%
  map_chr(extract(1)) %>%
  str_sub(start = k + 4)
```

## extract domain extension

<hr>

```{r str47}
url1 %>%
  str_sub(end = n) 
```

## extract domain extension

<hr>

```{r str48}
url1 %>%
  str_sub(end = n) %>%
  str_split(pattern = '\\.') 
```

## extract domain extension

<hr>

```{r str49}
url1 %>%
  str_sub(end = n) %>%
  str_split(pattern = '\\.') %>%
  map_chr(extract(2)) 
```

## extract domain extension

<hr>

```{r str50}
url1 %>%
  str_sub(end = n) %>%
  str_split(pattern = '\\.') %>%
  map_chr(extract(2)) %>%
  str_split(pattern = '/') 
```

## extract domain extension

<hr>

```{r str51}
url1 %>%
  str_sub(end = n) %>%
  str_split(pattern = '\\.') %>%
  map_chr(extract(2)) %>%
  str_split(pattern = '/') %>%
  map_chr(extract(1))
```

## extract domain extension

<hr>

```{r str52}
urls %>%
  str_sub(end = n) %>%
  str_split(pattern = '\\.') %>%
  map_chr(extract(2)) %>%
  str_split(pattern = '/') %>%
  map_chr(extract(1))
```

## extract file type

<hr>

```{r str14}
mockstring$url[1:3]
```

## Steps

<hr>

- check if there are only 2 dots in the URL
- check if there is only 1 question mark in the URL
- detect the location of the second dot 
- detect the location of the first question mark
- use the locations to specify the index position for extracting file type

## step 1: check if there are only 2 dots in the URL

<hr>

```{r str9}
mockstring$url[1:3] %>%
  str_locate_all(pattern = '\\.') %>%
  map_int(nrow) %>%
  is_greater_than(2) %>%
  sum()
```

## step 2: check if there is only 1 question mark in the URL

<hr>

```{r str10}  
mockstring$url[1:3] %>%
  str_locate_all(pattern = "[?]") %>%
  map_int(nrow) %>%
  is_greater_than(1) %>%
  sum()
```

## step 3: detect the location of the second dot

<hr>

```{r str11}
d <- mockstring$url[1:3] %>%
  str_locate_all(pattern = '\\.') %>%
  map_int(extract(2)) %>%
  add(1)

d  
```

## step 4: detect the location of the first question mark

<hr>

```{r str12}
q <- mockstring$url[1:3] %>%
  str_locate_all(pattern = "[?]") %>%
  map_int(extract(1)) %>%
  subtract(1)

q
```

## step 5: specify the index position for extracting file type

<hr>

```{r str13}
mockstring$url[1:3] %>% 
  str_sub(start = d, end = q)
```