---
title: "Hacking strings with stringr"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
---

<style type="text/css">
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: capitalize;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
  }
  .reveal h1 {
    font-size: 24px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
-  }

  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

## Introduction

<hr>

In this module, we will learn to work with string data in R using [stringr](http://stringr.tidyverse.org). As we did in the earlier modules, we will use a case study to explore the various features of the stringr package. You can
download the data for the case study from [here](https://raw.githubusercontent.com/rsquaredacademy/datasets/master/mock_strings.csv) or directly import the data using the readr package.  Let us begin by installing and loading stringr and a set of other pacakges we will be using.
 
## Installation

<hr>

```{r install, eval=FALSE}
library(stringr)
library(tibble)
library(magrittr)
library(purrr)
library(dplyr)
```

```{r str1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(stringr)
library(tibble)
library(magrittr)
library(purrr)
library(dplyr)
```

## Data

<hr>

```{r import, eval=FALSE}
mockstring <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/mock_strings.csv')
```

```{r show, echo=FALSE, eval=TRUE, message=FALSE}
mockstring <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/mock_strings.csv')
mockstring
```

## Data Dictionary

<hr>

## Case Study

<hr>

- extract domain name from random email ids
- extract image type from url
- extract image dimension from url
- extract extension from domain name
- extract http protocol from url
- extract domain name from url
- extract extension from url
- extract file type from url

## extract domain name from email ids

<hr>

```{r str2}
mockstring$email[1:2] %>% 
  str_split(pattern = '@') %>%
  unlist() %>%
  str_subset(pattern = '\\.') %>%
  str_split(pattern = '\\.') %>%
  purrr::map_chr(extract(1))
```


## Extract image type from URL

<hr>

```{r str3}
mockstring$imageurl[1:5] %>%
  str_split(pattern = '\\.') %>%
  purrr::map_chr(extract(3)) %>%
  str_sub(start = 1, end = 3)

mockstring$imageurl[1:5] %>%
  str_split(pattern = '\\.') %>%
  purrr::map_chr(extract(3)) %>%
  str_split(pattern = '/') %>%
  purrr::map_chr(extract(1))
```

## Extract Image Dimesion from URL

<hr>

```{r str4}
pattern <- "[:digits:]"
mockstring$imageurl[1:5] %>%
  str_locate(pattern = "[0-9]") 
  
mockstring$imageurl[1:5] %>%
  str_sub(start = 23) %>%
  str_split(pattern = '\\.') %>%
  purrr::map_chr(extract(1))
```
  
## Extract HTTP Protocol from URL

<hr>

```{r str6}
mockstring$url[1:5] %>%
  str_split(pattern = '://') %>%
  map_chr(extract(1))
  
k <- mockstring$url[1:5] %>%
  str_locate(pattern = '://') %>%
  data.frame() %>%
  extract2('start') %>%
  subtract(1)

mockstring$url[1:5] %>%
  str_sub(start = 1, end = k)
```

## Extract Domain Name

<hr> 

```{r str7}
n <- mockstring$url[1:5] %>%
  str_locate_all(pattern = "/") %>%
  map_int(extract(3))

mockstring$url[1:5] %>%
  str_sub(end = n) %>%
  str_split(pattern = '\\.') %>%
  map_chr(extract(1)) %>%
  str_sub(start = k + 4)
```

## extract domain extension

<hr>

```{r str8}
mockstring$url[1:5] %>%
  str_sub(end = n) %>%
  str_split(pattern = '\\.') %>%
  map_chr(extract(2)) %>%
  str_split(pattern = '/') %>%
  map_chr(extract(1))
```

## extract file type

<hr>

```{r str14}
mockstring$url[1:3]
```

## Steps

<hr>

- check if there are only 2 dots in the URL
- check if there is only 1 question mark in the URL
- detect the location of the second dot 
- detect the location of the first question mark
- use the locations to specify the index position for extracting file type

## step 1: check if there are only 2 dots in the URL

<hr>

```{r str9}
mockstring$url[1:3] %>%
  str_locate_all(pattern = '\\.') %>%
  map_int(nrow) %>%
  is_greater_than(2) %>%
  sum()
```

## step 2: check if there is only 1 question mark in the URL

<hr>

```{r str10}  
mockstring$url[1:3] %>%
  str_locate_all(pattern = "[?]") %>%
  map_int(nrow) %>%
  is_greater_than(1) %>%
  sum()
```

## step 3: detect the location of the second dot

<hr>

```{r str11}
d <- mockstring$url[1:3] %>%
  str_locate_all(pattern = '\\.') %>%
  map_int(extract(2)) %>%
  add(1)

d  
```

## step 4: detect the location of the first question mark

<hr>

```{r str12}
q <- mockstring$url[1:3] %>%
  str_locate_all(pattern = "[?]") %>%
  map_int(extract(1)) %>%
  subtract(1)

q
```

## step 5: specify the index position for extracting file type

<hr>

```{r str13}
mockstring$url[1:3] %>% 
  str_sub(start = d, end = q)
```

## Question

<hr>

## Thank You

<hr>


