---
output: _output.yaml
---

## {data-background="stringr.png"}

## Agenda

<hr>

- detect patterns 
- count occurence of patterns 
- split strings
- replace strings
- extract patterns 
- locate patterns 
 
## Libraries

<hr>

```{r install, eval=FALSE}
library(stringr)
library(dplyr)
library(readr)
```

```{r str1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(stringr)
library(dplyr)
library(readr)
```

## Data

<hr>

```{r import, eval=FALSE}
mockstring <- read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/mock_strings.csv')
```

```{r show, echo=FALSE, eval=TRUE, message=FALSE}
mockstring <- read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/mock_strings.csv')
mockstring
```

## Case Study

<hr>

- extract domain name from random email ids
- extract image type from url
- extract image dimension from url
- extract extension from domain name
- extract http protocol from url
- extract domain name from url
- extract extension from url
- extract file type from url

## Sample Data

<hr>


```{r sample, echo = FALSE}
mock_data <-
  mockstring %>%
  slice(1:10) %>%
  select(email, address, full_name, currency, )
```

```{r sample2}
mock_data
```

## {data-background="str_canva_detect.png"}

## Detect @

<hr>

```{r detect}
str_detect(mock_data$email, pattern = "@")
```


## {data-background="str_canva_count.png"}

## Count @

<hr>

```{r count}
str_count(mock_data$email, pattern = "@")
```

## {data-background="str_canva_c.png"}

## Concatenate

<hr>

```{r concatenate}
str_c("email id:", mock_data$email)
```

## {data-background="str_canva_split.png"}

## Split

<hr>

```{r split}
str_split(mock_data$email, pattern = "@")
```

## {data-background="str_canva_sort.png"}

## Sort

<hr>

```{r sort}
str_sort(mock_data$email)
```

## {data-background="str_canva_sort_descending.png"}

## Sort

<hr>

```{r sort2}
str_sort(mock_data$email, descending = TRUE)
```

## {data-background="str_canva_to_upper.png"}

## Case

<hr>

```{r case}
str_to_upper(mock_data$full_name)
```


## {data-background="str_canva_replace.png"}

## Replace

<hr>

```{r replace}
str_replace(mock_data$address, "Street", "ST")
```


## {data-background="str_canva_extract.png"}

## Extract

<hr>

```{r extract}
str_extract(mock_data$email, pattern = "com")
```


## {data-background="str_canva_match.png"}

## Match

<hr>

```{r match}
str_match(mock_data$email, pattern = "com")
```


## {data-background="str_canva_which.png"}

## Index

<hr>

```{r which}
str_which(mock_data$email, pattern = "com")
```


## {data-background="str_canva_locate.png"}

## Locate

<hr>

```{r locate}
str_locate(mock_data$email, pattern = "com")
```

## {data-background="str_canva_length.png"}

## {data-background="str_canva_sub.png"}

## Extract

<hr>

```{r sub}
str_sub(mock_data$currency, start = 1, end = 1)
```

## {data-background="str_canva_word.png"}

## Word

<hr>

```{r word}
word(mock_data$full_name, 1)
```


## {data-background="thankyou.png"}
