---
title: "Hacking strings with stringr"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: moon
    highlight: monochrome
    center: true
    reveal_options:
      slideNumber: true
      previewLinks: true
---

## Introduction

In this module, we will learn to work with string data in R using [stringr](http://stringr.tidyverse.org). Let us begin by installing and loading stringr and a set of other pacakges we will be using.
 
## Installation

```{r install, eval=FALSE}
library(stringr)
library(tibble)
library(magrittr)
library(purrr)
library(dplyr)
```

```{r str1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(stringr)
library(tibble)
library(magrittr)
library(purrr)
library(dplyr)
```

## Introduction

As we did in the earlier modules, we will use a case study to explore the various features of the stringr package. You can
download the data for the case study from [here](https://raw.githubusercontent.com/rsquaredacademy/datasets/master/mock_strings.csv) or directly import the data using the readr package. 

## Case Study

```{r import, eval=FALSE}
mockstring <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/mock_strings.csv')
```

```{r show, echo=FALSE, eval=TRUE, message=FALSE}
mockstring <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/mock_strings.csv')
mockstring
```

## Data Dictionary

----

We will do the following in this case study:

- extract domain name from random email ids
- extract image type from url
- extract image dimension from url
- extract extension from domain name
- extract http protocol from url
- extract domain name from url
- extract extension from url
- extract file type from url

## extract domain name from random email ids

----

```{r str2}
mockstring$email[1:2] %>% 
  str_split(pattern = '@') %>%
  unlist() %>%
  str_subset(pattern = '\\.') %>%
  str_split(pattern = '\\.') %>%
  purrr::map_chr(extract(1))
```


## from image url, extract image type (jpg, png, bmp)

----

```{r str3}
mockstring$imageurl[1:5] %>%
  str_split(pattern = '\\.') %>%
  purrr::map_chr(extract(3)) %>%
  str_sub(start = 1, end = 3)

mockstring$imageurl[1:5] %>%
  str_split(pattern = '\\.') %>%
  purrr::map_chr(extract(3)) %>%
  str_split(pattern = '/') %>%
  purrr::map_chr(extract(1))
```

## from image url, extract dimension

----

```{r str4}
pattern <- "[:digits:]"
mockstring$imageurl[1:5] %>%
  str_locate(pattern = "[0-9]") 
  
mockstring$imageurl[1:5] %>%
  str_sub(start = 23) %>%
  str_split(pattern = '\\.') %>%
  purrr::map_chr(extract(1))
```
  
## extract protocol

----

```{r str6}
mockstring$url[1:5] %>%
  str_split(pattern = '://') %>%
  map_chr(extract(1))
  
k <- mockstring$url[1:5] %>%
  str_locate(pattern = '://') %>%
  data.frame() %>%
  extract2('start') %>%
  subtract(1)

mockstring$url[1:5] %>%
  str_sub(start = 1, end = k)
```

## extract domain name

----

```{r str7}
n <- mockstring$url[1:5] %>%
  str_locate_all(pattern = "/") %>%
  map_int(extract(3))

mockstring$url[1:5] %>%
  str_sub(end = n) %>%
  str_split(pattern = '\\.') %>%
  map_chr(extract(1)) %>%
  str_sub(start = k + 4)
```

## extract extension

----

```{r str8}
mockstring$url[1:5] %>%
  str_sub(end = n) %>%
  str_split(pattern = '\\.') %>%
  map_chr(extract(2)) %>%
  str_split(pattern = '/') %>%
  map_chr(extract(1))
```

## extract file type

----

only 2 dots in the url

```{r str9}
mockstring$url[1:3] %>%
  str_locate_all(pattern = '\\.') %>%
  map_int(nrow) %>%
  is_greater_than(2) %>%
  sum()
```

----  

only 1 question mark in the url

```{r str10}  
mockstring$url[1:3] %>%
  str_locate_all(pattern = "[?]") %>%
  map_int(nrow) %>%
  is_greater_than(1) %>%
  sum()
```

----

location of second dot and 1 question mark

```{r str11}
d <- mockstring$url[1:3] %>%
  str_locate_all(pattern = '\\.') %>%
  map_int(extract(2)) %>%
  add(1)
```

----

```{r str12}
q <- mockstring$url[1:3] %>%
  str_locate_all(pattern = "[?]") %>%
  map_int(extract(1)) %>%
  subtract(1)
```

----

```{r str13}
mockstring$url[1:3] %>% 
  str_sub(start = d, end = q)
```

## Summary
