---
output: _output.yaml
---

## {data-background="lubridate.png"}

## Agenda

<hr>
 
## Libraries

<hr>

```{r install, eval=FALSE}
library(lubridate)
library(dplyr)
library(magrittr)
library(readr)
```

```{r lub1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(lubridate)
library(dplyr)
library(magrittr)
library(readr)
```

## Origin

<hr>

```{r origin}
lubridate::origin
```

## Current Date/Time

<hr>

```{r today}
now()
today()
am(now())  
pm(now())
```

## Data

<hr>

```{r import, eval=FALSE}
transact <- read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/transact.csv')
```

```{r show, echo=FALSE, eval=TRUE, message=FALSE}
transact <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/transact.csv')
transact
```

## Data Dictionary

<hr>

The data set has 3 columns. All the dates are in the format (yyyy-mm-dd).

- Invoice: invoice date 
- Due: due date
- Payment: payment date

## Case Study

<hr>

We will use the functions in the lubridate package to answer a few 
questions we have about the transact data.

- extract date, month and year from Due
- compute the number of days to settle invoice
- compute days over due
- check if due year is a leap year
- check when due day in february is 29, whether it is a leap year
- how many invoices were settled within due date
- how many invoices are due in each quarter
- what is the average duration between invoice date and payment date

## {data-background="canva_day_week_month.png"}

## Extract Date, Month & Year from Due Date

<hr>

```{r lub18}
this_day <- as_date('2017-03-23')
day(this_day)
month(this_day)
year(this_day)
```

## Extract Date, Month & Year from Due Date

<hr>

```{r lub2}
transact %>%
  mutate(
    due_day   = day(Due),
    due_month = month(Due),
    due_year  = year(Due)
  )
```

## {data-background="canva_course_duration.png"}

## Course Duration

<hr>


```{r lub19}
course_start    <- as_date('2017-04-12')
course_end      <- as_date('2017-04-21')
course_duration <- course_end - course_start
course_duration
```


## Compute days to settle invoice

<hr>

```{r lub3}
transact %>%
  mutate(
    days_to_pay = Payment - Invoice
  )
```

## Compute days over due

<hr>

```{r lub4}
transact %>%
  mutate(
    delay = Payment - Due
  )
```

## Is due year a leap year?

<hr>

```{r lub5}
transact %>%
  mutate(
    due_year = year(Due),
    is_leap  = leap_year(due_year)
  )
```

## If due day is February 29, is it a leap year?

<hr>

```{r lub6}
transact %>%
  mutate(
    due_day   = day(Due),
    due_month = month(Due),
    due_year  = year(Due),
    is_leap   = leap_year(due_year)
  ) %>%
  select(-(Invoice), -(Payment)) %>%
  filter(due_month == 2 & due_day == 29)
```

## {data-background="canva_shift_dates.png"}

## Shift Date

<hr>

```{r lab40}
course_start + days(2)
course_start + weeks(1)
course_start + years(1)
```

## {data-background="canva_course_interval.png"}

## Interval

<hr>

```{r lub10}
interval(course_start, course_end)
```

## {data-background="canva_shift_interval.png"}

## Shift Interval

<hr>

```{r lab50}
course_interval <- interval(course_start, course_end)
course_interval %>%
  int_shift(by = days(1))
course_interval %>%
  int_shift(by = weeks(1))
course_interval %>%
  int_shift(by = years(1))
```

## {data-background="canva_interval_overlap.png"}

## Intervals Overlap

<hr>

```{r lub60}
vacation_start    <- as_date('2017-04-19')
vacation_end      <- as_date('2017-04-25')
vacation_interval <- interval(vacation_start, vacation_end)
int_overlaps(course_interval, vacation_interval)
```

## How many invoices were settled within due date?

<hr>

```{r lub7}
transact %>%
  mutate(
    inv_due_interval = interval(Invoice, Due),
    due_next         = Due + days(1),
    due_pay_interval = interval(due_next, Payment),
    overlaps         = int_overlaps(inv_due_interval, due_pay_interval)
  ) %>%
  select(Invoice, Due, Payment, overlaps)
```

## how many invoices were settled within due date?

<hr>

```{r lub12}
# using int_shift
transact %>%
  mutate(
    inv_due_interval = interval(Invoice, Due),
    due_pay_interval = interval(Due, Payment),  
    due_pay_next     = int_shift(due_pay_interval, by = days(1)),
    overlaps         = int_overlaps(inv_due_interval, due_pay_next)
  ) %>%
  select(Invoice, Due, Payment, overlaps)
```

## {data-background="canva_within.png"}

## Within

<hr>

```{r lub30}
conference <- as_date('2017-04-15')
conference %within% interval(course_start, course_end)
```

## how many invoices were settled within due date?

<hr>

```{r lub13}
transact %>%
  mutate(
    inv_due_interval = interval(Invoice, Due),
    overlaps = Payment %within% inv_due_interval
  ) %>%
  select(Due, Payment, overlaps)
```

## {data-background="canva_quarter_semester.png"}

## Quarter

<hr>

```{r lub20}
course_start
course_start %>%
  quarter()
course_start %>%
  quarter(with_year = TRUE)
```


## Quarter

<hr>

```{r lub8}
transact %>%
  mutate(
    Quarter = quarter(Due)
  )
```

## Case Study

<hr>

```{r lub14}
transact %>%
  mutate(
    Quarter = quarter(Due, with_year = TRUE)
  )
```

## Duration

<hr>

```{r lub15}
course_duration %>%
  as.numeric() %>%
  duration(units = 'days')
```

## {data-background="canva_convert.png"}

## Case Study

<hr>

```{r lub11}
interval(course_start, course_end) / dseconds()
interval(course_start, course_end) / dminutes()
interval(course_start, course_end) / dhours()
interval(course_start, course_end) / dweeks()
interval(course_start, course_end) / dyears()
```

## {data-background="canva_time_length.png"}

## Case Study

<hr>

```{r lub16}
interval(course_start, course_end) %>%
  time_length(unit = "seconds")
interval(course_start, course_end) %>%
  time_length(unit = "minutes")
interval(course_start, course_end) %>%
  time_length(unit = "hours")
```

## {data-background="canva_as_period.png"}

## Case Study

<hr>

```{r lub17}
interval(course_start, course_end) %>%
  as.period(unit = "seconds")
interval(course_start, course_end) %>%
  as.period(unit = "minutes")
interval(course_start, course_end) %>%
  as.period(unit = "hours")
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add agenda -->
