---
output: _output.yaml
---

## {data-background="forcats.png"}

## Agenda

<hr>

In this post, we will learn to:

- create matrices
- combining matrices
- index/subset matrices
- dissolve matrices

## Introduction

<hr>

In this post, we will learn to create matrices. A matrix is a regular array of data elements, arranged in rows and columns. Matrices in R are homogeneous i.e. they can hold a single type of data. Matrix elements are indexed by specifying the row and column index and the elements of a matrix can filled either by row or column. 

## Creating Matrix

<hr>

The easiest way to create a matrix in R is to use the `matrix()` function. Let us look at its syntax:

```{r mat1}
args(matrix)
```

Now that we have understood the syntax of the `matrix()` function, let us create a simple numeric matrix.

```{r mat2, collapse=TRUE}
# matrix of 3 rows filled by columns
mat <- matrix(data = 1:9, nrow = 3, byrow = FALSE)
mat
```

## Creating Matrix

<hr>

In the above example, we created a matrix of 3 rows where the data elements are filled by columns. We need to specify either the 
number of `rows` or `columns` and R will automatically compute the other. The number of data elements should be equal to the product of the rows and columns, else R will return a warning.

```{r mat3, collapse=TRUE}
matrix(data = 1:9, nrow = 2, byrow = FALSE)
matrix(data = 1:10, nrow = 3, byrow = FALSE)
```

## Creating Matrix

<hr>

We can follow some general rules to avoid the mistakes made in the previous examples:

- if the number of elements is odd, both the number of rows and columns must be odd and their product should equal
the number of data elements

- if the number of elements is even, either the number of rows or columns must be even. In some cases, both the rows and
columns must be even

Let us continue to explore the syntax of the `matrix()` function. 

## Fill Data by Row

<hr>

```{r mat4}
matrix(data = 1:9, nrow = 3, byrow = TRUE)
```

## Fill Data by Column

<hr>

```{r mat5}
matrix(data = 1:9, nrow = 3, byrow = FALSE)
```

## Specify Rows

<hr>

```{r mat6}
matrix(data = 1:10, nrow = 2)
```

## Specify Columns

<hr>

```{r mat7}
matrix(data = 1:10, ncol = 5)
```

## Row & Column Names

<hr>

You can specify names for the rows and columns of a matrix. To do so, we need to use `list`. Lists can contain other data structures such as vectors, matrices and even other lists. They are heterogeneous i.e. they can contain different data types.	We will learn more about lists in a future post, for the time being let us learn how to create a basic list using the `list()` function:

## Row & Column Names

<hr>

```{r mat8, collapse=TRUE}
# character vector
first_name <- 	c("John", "Jill", "Jack")

# numeric vector
age <- c(20, 24, 32)

# list 
details <- list(first_name, age)
details
```

## Row & Column Names

<hr>

Now that we know how to create a list, let us go ahead and create a matrix and name its rows and columns.

```{r mat9, collapse=TRUE}
# row names
row_names <- c('row_1', 'row_2', 'row_3')

# column names
col_names <- c('col_1', 'col_2', 'col_3')

# matrix with row and column names
matrix(data = 1:9, nrow = 3, dimnames = list(row_names, col_names))
```

## Matrix Dimension

<hr>

Another useful function is `dim()`. It can be used to:

- check the dimension (rows and columns) of a matrix
- modify the dimension of a matrix
- coerce a vector to a matrix

## Check dimension of a matrix

<hr>

```{r mat10}
mat <- matrix(data = 1:9, nrow = 3, byrow = TRUE)
dim(mat)
```

## Modify dimension of a matrix

<hr>

```{r mat11, collapse=TRUE}
# sample matrix
mat <- matrix(data = 1:12, nrow = 3, byrow = TRUE)
dim(mat)

# change the dimension to 4 x 3
dim(mat) <- c(4, 3)
dim(mat)
```

## Coerce vector to matrix

<hr>

```{r mat12, collapse=TRUE}
# numeric vector
vect1 <- 1:12
vect1

# coerce vect1 to a 4 x 3 matrix
dim(vect1) <- c(4, 3)
vect1
```

## Coerce vector to matrix

<hr>

Another way to coerce an R data structure to `matrix` is to use the `as.matrix()` function. Since the only other data structure we have learnt so far is the vector, we will coerce a vector into a matric using `as.matrix()`. We will deal with the other data structures as and when we learn them.

```{r mat13, collapse=TRUE}
# numeric vector
vect1 <- 1:12
vect1

# coerce vect1 to a matrix
as.matrix(vect1)
```

Regardless of the data type of the vector, all of them will be coerced to a matrix of dimension `n x 1` i.e. they will all have only one column.

## Append Data

<hr>

In this section, we will learn how to append data to a matrix. There are two functions that can be used for this purpose:

- `rbind()`
- `cbind()`

`cbind` will append a new column to the matrix while `rbind` will append a new row. 

## Append Row/Column

<hr>

```{r mat20, collapse=TRUE}
# 3 x 3 matrix
mat <- matrix(data = 1:9, nrow = 3)
mat

# numeric vector
v <- c(10, 11, 12)
v

# append row
rbind(mat, v)

# append column
cbind(mat, v)
```

## Combine Matrices

<hr>

When you use `rbind` to combine two matrices, the number of columns must match and in case of `cbind`, the number of rows must match.

## Append Row/Column

<hr>

```{r mat21, collapse=TRUE}
# 3 x 3 matrix
mat1 <- matrix(data = 1:9, nrow = 3)
mat2 <- matrix(data = sample(9), nrow = 3)

# append rows
rbind(mat1, mat2)

# append columns
cbind(mat1, mat2)
```

## Subset Matrices

<hr>

In this section, we will learn to subset matrices. The `[]` operator can be used to subset matrices just like vectors but since matrices are two dimensional, we need to specify both the row number and the column number. Below are a few examples:

```{r mat22, collapse=TRUE}
# 3 x 4 matrix
mat <- matrix(data = 1:12, nrow =3)
mat

# extract element from first row, first column
mat[1, 1]

# extract all rows of first column
mat[, 1]

# extract all columns of first row
mat[1,]

# extract 2nd and 3rd row of first column
mat[c(2, 3), 1]

# extract 2nd and 3rd column of first row
mat[1, c(2, 3)]

# extract 2nd and 3rd row of first and third column
mat[c(2, 3), c(1, 3)]
```

## Using Row & Column Names

<hr>

In an earlier section, we learnt how to name the rows and columns of a matrix. Let us see how these names can be used to subset matrices.

```{r mat23, collapse=TRUE}
# row names
row_names <- c('row_1', 'row_2', 'row_3')

# column names
col_names <- c('col_1', 'col_2', 'col_3')

# matrix with row and column names
mat <- matrix(data = 1:9, nrow = 3, dimnames = list(row_names, col_names))

# extract elements from first row/columns
mat['row_1', 'col_1']

# extract all rows of first column
mat[, 'col_1']

# extract all columns of first row
mat['row_1',]
```

## Using Logical Expressions

<hr>

We can use logical expressions to subset matrices.

```{r mat24, collapse=TRUE}
# 3 x 4 matrix
mat <- matrix(data = 1:12, nrow =3)
mat

# elements greater than 4
mat > 4

# extract elements greater than 4
mat[mat > 4]
```

## Dissolve Matrices

<hr>

Till now we have learnt how to coerce a vector into matrix. Now let us learn how to coerce a matrix into a vector using:

- `c()`
- `as.vector()`

```{r mat25, collapse=TRUE}
# 3 x 3 matrix
mat <- matrix(data = 1:9, nrow =3)
mat

# using c()
c(mat)

# using as.vector()
as.vector(mat)
```

## {data-background="thankyou.png"}
  