---
output: _output.yaml
---

## {data-background="pipes.png"}

## Agenda

<hr>

- measures of location
- measures of variation
- measures of symmetry
- tabulation
- grouped summaries
- data visualization

## Introduction

<hr>

- what is descriptive statistics?
- why you need descriptive statistics?

## Libraries

<hr>

- [descriptr](https://descriptr.rsquaredacademy.com)

```{r install, echo=FALSE, message=FALSE}
library(dplyr)
library(descriptr)
```

## Data

<hr> 

```{r subset}
mtcars
```

## Measures of Location

<hr> 

- mean
- trimmed mean
- median
- mode

## Mean

<hr>

```{r mean}
mean(mtcars$mpg)
```

## Trimmed Mean

<hr>

```{r trim_mean}
mean(mtcars$mpg, trim = 0.05)
```

## Median

<hr>

```{r median}
median(mtcars$mpg)
```

## Mode

<hr>

```{r mode}
ds_mode(mtcars$mpg)
```

## Measures of Variation

<hr> 

- range
- inter-quartile range
- variance
- standard deviation
- coefficient of variation
- standard error

## Range

<hr>

```{r range}
range(mtcars$mpg)
```

## Inter Quartile Range

<hr>

```{r iqr}
IQR(mtcars$mpg)
```

## Variance

<hr>

```{r var}
var(mtcars$mpg)
```

## Standard Deviation

<hr>

```{r sd}
sd(mtcars$mpg)
```

## Coefficient of Variation

<hr>

```{r cvar}
ds_cvar(mtcars$mpg)
```

## Measures of Symmetry

<hr> 

- skewness
- kurtosis

## Skewness

<hr>

```{r skewness}
ds_skewness(mtcars$mpg)
```

## Kurtosis

<hr>

```{r kurtosis}
ds_kurtosis(mtcars$mpg)
```

## Percentiles/Quantiles

<hr> 

```{r percentiles}
ds_percentiles(mtcars, mpg)
```

## Summary

<hr>

```{r summary}
summary(mtcars)
```

## Screen Data

<hr>

```{r screen}
ds_screener(mtcarz)
```

## Summary Statistics

<hr>

```{r summary_stats}
ds_summary_stats(mtcarz, mpg)
```

## Tidy Statistics

<hr>

```{r tidy_stats}
ds_tidy_stats(mtcarz)
```

## Frequency Table

<hr>

```{r freq_table}
ds_freq_table(mtcarz, mpg)
```

## Frequency Table

<hr>

```{r freq_table_plot, fig.align='center'}
mtcarz %>%
  ds_freq_table(mpg) %>%
  plot()
```

## Measures of Location

<hr>

```{r location}
ds_measures_location(mtcarz)
```

## Measures of Variation

<hr>

```{r variation}
ds_measures_variation(mtcarz)
```

## Measures of Symmetry

<hr>

```{r symmetry}
ds_measures_symmetry(mtcarz)
```

## {data-background="thankyou.png"}

