---
output: _output.yaml
---

## {data-background="pipes.png"}

## Agenda

<hr>

- measures of location
- measures of variation
- measures of symmetry
- tabulation
- grouped summaries
- data visualization

## Introduction

<hr>

- what is descriptive statistics?
- why you need descriptive statistics?

## Libraries

<hr>

- [descriptr](https://descriptr.rsquaredacademy.com)
- [skimr](https://github.com/ropensci/skimr)
- [Hmisc](https://cran.r-project.org/package=Hmisc)
- [pastecs](https://cran.r-project.org/package=pastecs)
- [psych](https://cran.r-project.org/package=psych)
- [summarytools](https://cran.r-project.org/package=summarytools)
- [desctable](https://cran.r-project.org/package=desctable)
- [GGally](http://ggobi.github.io/ggally/)

```{r install, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(descriptr)
library(skimr)
library(Hmisc)
library(pastecs)
library(psych)
library(skimr)
library(summarytools)
library(desctable)
library(GGally)
```

## Data

<hr> 

```{r subset}
data <- dplyr::select(mtcarz, mpg, disp, gear, cyl)
data
```

## Hmisc

<hr>

```{r descibe}
Hmisc::describe(data)
```

## pastecs

<hr>

```{r pastecs}
pastecs::stat.desc(data)
```

## psych

<hr>

```{r psych}
psych::describe(data)
```

## summarytools

<hr>

```{r summarytools}
summarytools::descr(data)
```

## summarytools

<hr>

```{r dfsummary}
summarytools::dfSummary(data)
```

## desctable

<hr>

```{r desctable}
desctable::desctable(data)
```

## ggpairs

<hr>

```{r ggally, fig.align='center', warning=FALSE, message=FALSE}
GGally::ggpairs(data)
```

## skimr

<hr>

```{r skimr}
skimr::skim(data)
```

## {data-background="thankyou.png"}

