---
title: "Exporting Data from R"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: moon
    highlight: monochrome
    center: true
    fig_width: 6
    fig_height: 5
    reveal_options:
      slideNumber: true
      previewLinks: true
---

## Introdction

## Libraries

```{r importlib1,eval=FALSE}
library(readr)
library(haven)
library(jsonlite)
library(xml2)
```

```{r importlib2, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(readr)
library(haven)
library(jsonlite)
library(xml2)
```

## Define Arguments

- x/data: name of the data.frame/tibble
- path: path where file must be written
- delim: delimiter to separate values
- append: whether append data or overwrite file (if there is no file, a new
 one will be created)
- col_names: write column names in the first row/top of file
- compress: compression method
- na: string used for representing missing values in data


## CSV

Format

```{r expo1, eval=FALSE}
write_csv(x, path, na = "NA", append = FALSE, col_names = !append)
```

Example

```{r expo2}
write_csv(x = mtcars, path = "mtcars.csv")
```

## TSV

Format

```{r expo3, eval=FALSE}
write_tsv(x, path, na = "NA", append = FALSE, col_names = !append)
```

Example

```{r expo4}
write_tsv(x = mtcars, path = "mtcars.tsv")
```

## Delimited File

Format

```{r expo5, eval=FALSE}
write_delim(x, path, na = "NA", append = FALSE, col_names = !append)
```

Example

```{r expo6}
write_delim(x = mtcars, path = "mtcars.txt")
```

## Write Lines

Format

```{r expo9, eval=FALSE}
write_lines(x, path, na = "NA", append = FALSE)
```

Example

```{r expo10}
write_lines(x = mtcars, path = "mtcars2.txt")
```

## SAS

Format

```{r expo11, eval=FALSE}
write_sas(data, path)
```

Example

```{r expo12}
write_sas(data = mtcars, path = "mtcars.sas7bdat")
```

## SPSS

Format

```{r expo13, eval=FALSE}
write_sav(data, path)
```

Example

```{r expo14}
write_sav(data = mtcars, path = "mtcars.sav")
```

## STATA

Format

```{r expo15, eval=FALSE}
write_dta(data, path, version = 14)
```

Example

```{r expo16}
write_dta(data = mtcars, path = "mtcars.dta", version = 14)
```

## JSON

Format

```{r expo17, eval=FALSE}
write_json(x, path, ...)
```

Example

```{r expo18}
write_json(x = mtcars, path = "mtcars")
```

## RDS

Format

```{r expo23, eval=FALSE}
write_rds(x, path, compress = c("none", "gz", "bz2", "xz"), ...)
```

Example

```{r expo24}
write_rds(x = mtcars, path = "mtcars", compress = "gz")
```

## Summary

## Next Steps..