---
title: "Exporting Data from R"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
---

<style type="text/css">
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: capitalize;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal .sourceCode pre code {
    padding: 0px;
    border: 0px;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>


## Agenda

<hr>

Export data from R to

- csv
- tsv
- text file
- JSON
- SAS
- STATA
- SPSS
- RDS

## Introdction

<hr>

## Libraries

<hr>

```{r importlib1,eval=FALSE}
library(readr)
library(haven)
library(jsonlite)
library(xml2)
```

```{r importlib2, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(readr)
library(haven)
library(jsonlite)
library(xml2)
```

## Commom Arguments

<hr>

- x/data: name of the data.frame/tibble
- path: path where file must be written
- delim: delimiter to separate values
- append: whether append data or overwrite file (if there is no file, a new
 one will be created)
- col_names: write column names in the first row/top of file
- compress: compression method
- na: string used for representing missing values in data


## CSV

<hr>

Format

```{r expo1, eval=FALSE}
write_csv(x, path, na = "NA", append = FALSE, col_names = !append)
```

Example

```{r expo2}
write_csv(x = mtcars, path = "mtcars.csv")
```

## TSV

<hr>

Format

```{r expo3, eval=FALSE}
write_tsv(x, path, na = "NA", append = FALSE, col_names = !append)
```

Example

```{r expo4}
write_tsv(x = mtcars, path = "mtcars.tsv")
```

## Delimited File

<hr>

Format

```{r expo5, eval=FALSE}
write_delim(x, path, na = "NA", append = FALSE, col_names = !append)
```

Example

```{r expo6}
write_delim(x = mtcars, path = "mtcars.txt")
```

## Write Lines

<hr>

Format

```{r expo9, eval=FALSE}
write_lines(x, path, na = "NA", append = FALSE)
```

Example

```{r expo10}
write_lines(x = mtcars, path = "mtcars2.txt")
```

## SAS

<hr>

Format

```{r expo11, eval=FALSE}
write_sas(data, path)
```

Example

```{r expo12}
write_sas(data = mtcars, path = "mtcars.sas7bdat")
```

## SPSS

<hr>

Format

```{r expo13, eval=FALSE}
write_sav(data, path)
```

Example

```{r expo14}
write_sav(data = mtcars, path = "mtcars.sav")
```

## STATA

<hr>

Format

```{r expo15, eval=FALSE}
write_dta(data, path, version = 14)
```

Example

```{r expo16}
write_dta(data = mtcars, path = "mtcars.dta", version = 14)
```

## JSON

<hr>

Format

```{r expo17, eval=FALSE}
write_json(x, path, ...)
```

Example

```{r expo18}
write_json(x = mtcars, path = "mtcars")
```

## RDS

<hr>

Format

```{r expo23, eval=FALSE}
write_rds(x, path, compress = c("none", "gz", "bz2", "xz"), ...)
```

Example

```{r expo24}
write_rds(x = mtcars, path = "mtcars", compress = "gz")
```

## Questions

<hr>

## Thank You

<hr>