---
output: _output.yaml
---

## {data-background="import.png"}

## Agenda

<hr>

- read data from flat or delimited files
- specify column/variable types
- read specific columns/variables 

## Libraries

<hr>

```{r importlib1,eval=FALSE}
library(readr)
```

```{r importlib2, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(readr)
```

## {data-background="csv_file.png"}

## {data-background="semi_colon_file.png"}

## {data-background="space_file.png"}

## {data-background="tab_file.png"}

## CSV 

<hr>

```{r import2, echo=TRUE, results='markup', message=FALSE}
read_csv('mtcars.csv')
```

## CSV 

<hr>

```{r import2a, echo=TRUE, results='markup', message=FALSE}
read_delim('mtcars.csv', delim = ",")
```

## {data-background="columns_names.png"}

## Column Names 

<hr>

```{r import4a, echo=TRUE, results='markup', message=FALSE}
read_csv('mtcars1.csv')
```

## Column Names 

<hr>

```{r import4, echo=TRUE, results='markup', message=FALSE}
read_csv('mtcars1.csv', col_names = FALSE)
```

## {data-background="mtcars_skip_lines_file.png"}

## Skip Lines 

<hr>

```{r import6a, echo=TRUE, results='markup', message=FALSE}
read_csv('mtcars2.csv')
```

## Skip Lines 

<hr>

```{r import6, echo=TRUE, results='markup', message=FALSE}
read_csv('mtcars2.csv', skip = 19)
```

## Maximum Lines 

<hr>

```{r import8, echo=TRUE, results='markup', message=FALSE}
read_csv('mtcars.csv', n_max = 20)
```

## {data-background="column_types_list_canva.png"}

## Column Types

<hr>

```{r import55, echo=TRUE, results='markup', message=FALSE}
spec_csv('mtcars5.csv')
```

## {data-background="column_objective_canva.png"}

## Column Types 

<hr>

```{r import10, echo=TRUE, results='markup', message=FALSE}
read_csv('mtcars5.csv', 
         col_types = list(col_double(), col_factor(levels = c(4, 6, 8)),
                          col_double(), col_integer()))
```

## Skip Columns

<hr>

```{r import10a, echo=TRUE, results='markup', message=FALSE}
read_csv('mtcars5.csv', 
         col_types = list(col_double(), col_factor(levels = c(4, 6, 8)),
                          col_skip(), col_integer()))
```


## Read Specific Columns 

<hr>

```{r import11, echo=TRUE, results='markup', message=FALSE}
read_csv('mtcars5.csv', 
         col_types = cols_only(mpg = col_double(), 
                               cyl = col_factor(levels = c(4, 6, 8))))
```

## {data-background="import_funcs.png"}

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - improve agenda -->
<!--   - outline functions and options in a table -->
