---
output: _output.yaml
---

## {data-background="import.png"}

## Agenda

<hr>

Read data from

- XML
- JSON

## Libraries

<hr>

```{r importlib1,eval=FALSE}
library(jsonlite)
library(xml2)
library(XML)
library(tibble)
library(purrr)
```

```{r importlib2, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(jsonlite)
library(xml2)
library(XML)
library(tibble)
library(purrr)
```

## {data-background="json.png"}

## JSON

<hr>

```{r imp1}
jsonlite::fromJSON('mtcars.json')
```

## {data-background="xml2.png"}

## Read Data

<hr>

```{r imp15}
books <- read_xml('books.xml')
```

## Name

<hr>

```{r imp16}
xml_name(books)
```

## Child Node

<hr>

```{r imp17}
xml_children(books)
```

## Child Node

<hr>

```{r imp18}
xml_child(books)
```

## Contents

<hr>

```{r imp19}
xml2::xml_contents(books)
```

## Structure

<hr>

```{r imp21}
xml2::xml_structure(books)
```

## Length

<hr>

```{r imp22}
xml2::xml_length(books)
```

## Root Node

<hr>

```{r imp23}
xml2::xml_root(books)
```

## Attributes

<hr>

```{r imp24}
xml_find_all(books, './/book') %>% 
  map_chr(xml_attrs)
```

## Book Author

<hr>

```{r imp25}
xml_find_all(books, './/author') %>%
  xml_text()
```

## Book Title

<hr>

```{r imp26}
xml_find_all(books, './/title') %>%
  xml_text()
```

## Book Genre

<hr>

```{r imp27}
xml_find_all(books, './/genre') %>%
  xml_text()
```

## Book Price

<hr>

```{r imp28}
xml_find_all(books, './/price') %>%
  xml_double()
```

## Publishing Date

<hr>

```{r imp29}
xml_find_all(books, './/publish_date') %>%
  xml_text()
```

## Description

<hr>

```{r imp30}
xml_find_all(books, './/description') %>%
  xml_text() %>%
  stringr::str_replace_all("[\n]", "") %>%
  stringr::str_replace_all(pattern = "      ", " ") %>%
  stringr::str_replace_all(pattern = "  ", " ")
```

## tibble

<hr>

```{r imp31}
book <- tibble::tibble(
  author = xml_find_all(books, './/author') %>%
    xml_text(),
  title = xml_find_all(books, './/title') %>%
    xml_text(),
  genre = xml_find_all(books, './/genre') %>%
    xml_text(),
  price = xml_find_all(books, './/price') %>%
    xml_double(),
  published = xml_find_all(books, './/publish_date') %>%
    xml_text(),
  description = xml_find_all(books, './/description') %>%
    xml_text() %>%
    stringr::str_replace_all("[\r\n]", "") %>%
    stringr::str_replace_all(pattern = "      ", " ") %>%
    stringr::str_replace_all(pattern = "  ", " ")
)
```

## tibble

<hr>

```{r imp32}
book
```

## {data-background="xml.png"}

<!-- ## XML -->

<!-- <hr> -->

<!-- - `xmlParse()`: parse xml data -->
<!-- - `class()`: class  -->
<!-- - `xmlRoot()`: root node -->
<!-- - `xmlChildren()`: child nodes -->
<!-- - get child node names -->
<!-- - get child node XML values -->
<!-- - `getNodeSet()`: get particular node value -->
<!-- - `xmlToDataFrame()`: coerce to data.frame  -->


## Read Data

<hr>

```{r imp2}
xm <- xmlParse("sample_xml")
```

## Class

<hr>

```{r imp3}
class(xm)
```

## Root Node

<hr>

```{r imp4}
xmlRoot(xm)
```

## Child Nodes

<hr>

```{r imp5}
c_child <- xmlChildren(xm)
```

## Child Node

<hr>

```{r imp6}
c_child[[1]]
```

## Child Node

<hr>

```{r imp7}
c_child[[1]][1]
```

## Node Name

<hr>

```{r imp9}
sapply(c_child[[1]][1], names)
```

## Node Value

<hr>

```{r imp10}
sapply(c_child[[1]][1], xmlValue)
```

## Book

<hr>

```{r imp11}
getNodeSet(xm, "//book")
```

## Author

<hr>

```{r imp12}
getNodeSet(xm, "//author")
```

## Dataframe

<hr>

```{r imp13}
j <- xmlToDataFrame(
  getNodeSet(xm, "//book")
)
j
```

## XML (Web)

<hr>

```{r xmlread, eval=TRUE}
# xml2::read_xml("http://www.xmlfiles.com/examples/cd_catalog.xml")
```

## HTML

<hr>

```{r imp14}
xml2::read_html("http://r-project.org/")
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add json data structure -->
<!--   - add XML data structure -->
<!--   - change html example -->
<!--   - add xml workflow diagram -->
<!--   - use this example: https://www.w3schools.com/xml/dom_nodes.asp -->
<!--   - try xml2 as well -->