---
title: "Importing Data into R: Part - 3"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
---

<style type="text/css">
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: capitalize;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal .sourceCode pre code {
    padding: 0px;
    border: 0px;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

read data from 

- XML
- JSON

## Introdction

<hr>

Importing data is the first step in any data analysis workflow. If you cannot import data from differernt sources or file formats, you cannot proceed with data analysis. R allows you to import data from several sources and a variety of file formats. In this module, we will focus on googlesheets, JSON and XML file formats.

## Libraries

<hr>

```{r importlib1,eval=FALSE}
library(googlesheets)
library(jsonlite)
library(xml2)
library(XML)
```

```{r importlib2, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(googlesheets)
library(jsonlite)
library(xml2)
library(XML)
```


## JSON

<hr>

read data from a json file

```{r imp1}
jsonlite::fromJSON('mtcars.json')
```

## XML

<hr>

read xml data 

```{r imp2}
xm <- xmlParse("sample_xml")
```

## XML

<hr>

```{r imp3}
class(xm)
```

## XML

<hr>

```{r imp4}
xmlRoot(xm)
```

## XML

<hr>

```{r imp5}
c_child <- xmlChildren(xm)
```

## XML

<hr>

```{r imp6}
c_child[[1]]
```

## XML

<hr>

```{r imp7}
c_child[[1]][1]
```

## XML

<hr>

```{r imp9}
sapply(c_child[[1]][1], names)
```

## XML

<hr>

```{r imp10}
sapply(c_child[[1]][1], xmlValue)
```

## XML

<hr>

```{r imp11}
getNodeSet(xm, "//book")
```

## XML

<hr>

```{r imp12}
getNodeSet(xm, "//author")
```

## XML

<hr>

```{r imp13}
j <- xmlToDataFrame(
  getNodeSet(xm, "//book")
)
j
```

## XML

<hr>

read xml data from the web

```{r xmlread, eval=FALSE}
cd <- read_xml("http://www.xmlfiles.com/examples/cd_catalog.xml")
```

## HTML

<hr>
 
read html data

```{r imp14}
xml2::read_html("http://had.co.nz")
```

## Questions

<hr>

## Thank You

<hr>