---
title: "Importing Data into R: Part - 4"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
---

<style type="text/css">
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: capitalize;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal .sourceCode pre code {
    padding: 0px;
    border: 0px;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

- read data from files
- read data from files line by line
- prompt users for input
- read data from a file in table format and create data.frame
- load R data sets and workspaces
- source R scripts

## Introdction

<hr>

Importing data is the first step in any data analysis workflow. If you cannot import data from differernt sources or file formats, you cannot proceed with data analysis. R allows you to import data from several sources and a variety of file formats. In this module, we will focus on JSON and XML file formats.

## Libraries

<hr>

```{r importlib1,eval=FALSE}
library(readr)
```

```{r importlib2, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(readr)
```

## Read Files

<hr>

```{r imp1, echo=TRUE, results='markup', message=FALSE}
readr::read_file('hsb.txt')
```

## Read Lines

<hr>

skip first 5 lines and read max 10 lines

```{r imp2, echo=TRUE, results='markup', message=FALSE}
readr::read_lines('hsb.txt', skip = 5, n_max = 10)
```

## scan 

<hr>

allow user to input data from console or read data from a file and store the input as a vector or list.

```{r imp3}
x <- scan()
x
```

## scan 

<hr>

```{r imp4}
x <- scan("", what = integer())
x
class(x)
```

## scan 

<hr>

```{r imp5}
x <- scan("", what = list(name = "", age = 0))
x
```

## readline

<hr>

prompt user for an input and store the same as a character vector

```{r imp6}
name <- readline(prompt = "What is your name?: ")
name
```

## readLines

<hr>

read data from a file

```{r imp7}
readLines("hsb.txt")
```

## readLines

<hr>

```{r imp8}
readLines("hsb.txt", n = 10)
```

## read.table

<hr>

read data from a file in table format and create data.frame

```{r imp9}
read.table('hsb.txt', header = TRUE, sep = ' ')
```

## load

<hr>

load: reloads saved data sets and workspaces

```{r imp10}
load('employees.RData')
```

## source

<hr>

source: reads R code from a script and makes those codes available in the current session

```{r imp11}
source('addition.R')
```

## Questions

<hr>

## Thank You

<hr>