---
output: _output.yaml
---

## {data-background="stringr.png"}

## Agenda

<hr>

- how to create strings in R
- basic string operations
	- nchar
	- toupper/tolower
	- is.character
	- toString/as.character
- format strings using:
	- cat
	- paste
	- paste0
	- format
	- glue
 
## Libraries

<hr>

```{r install, eval=FALSE}
library(glue)
```

```{r str1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(glue)
```

## intro

<hr>

```{r intro}
my_name <- "jovial"
my_name <- 'jovial'
```

## Data Type

<hr>

```{r }
my_name <- "Jovial"
my_name
class(my_name)
is.character(my_name)
```

## Number of Characters

<hr>

```{r nchar}
nchar(my_name)
```

## Case

<hr>

```{r change_case}
toupper(my_name)
tolower(my_name)
```

## Conversion

<hr>

```{r conversion}
my_age <- 30
toString(my_age)
as.character(my_age)
```

## cat

<hr>

```{r cat_1}
cat(my_name)
```

## cat

```{r cat_2}
cat("my name is ", my_name)
cat("my name is", my_name)
```

## cat

<hr>

```{r cat_3}
cat("my name is", my_name, sep = ",")
cat("my name is", my_name, sep = ";")
cat("my name is", my_name, sep = ".")
cat("my name is", my_name, sep = "\n")
```

## paste

<hr>

```{r paste1}
paste("my name is", my_name)
paste("my name is", my_name, sep = ",")
```

## paste

<hr>

```{r paste2}
full_name <- c("jovial", "mann")
paste(full_name, collapse = ", ")
```

## paste0

<hr>

```{r paste0}
paste0("my name is", my_name)
paste0("my name is ", my_name)
paste("my name is", my_name) == paste0("my name is ", my_name)
```

## format

<hr>

```{r format1}
format(12.5, digits = 2)
```

## format

<hr>

```{r format2}
format(12.5, nsmall = 3)
```

## format

<hr>

```{r format3}
format(12.5, width = 5)
```

## format

<hr>

```{r format4}
format(c(12.4, 4.82), justify = "left")
format(c(2.4, 4.82), justify = "right")
format(c(2.4, 4.8256), justify = "centre")
```

## glue

<hr>

```{r glue_1}
glue("my name is {my_name}")
```

## glue

<hr>

```{r glue_2}
glue("my first name is {first_name}",
     first_name = "jovial")
```

## glue

<hr>

```{r glue_3}
glue_collapse(full_name)
glue_collapse(full_name, sep = " ")
```

## glue

<hr>

```{r glue_4}
glue_collapse(full_name, sep = " ", width = 10)
```

## glue

<hr>

```{r glue_5}
glue_collapse(c("Jack", "Jill", "John", "Jovial."), 
              sep = ", ",
              last = " and ")
```


## {data-background="thankyou.png"}
