---
output: _output.yaml
---

## {data-background="pipes.png"}

## Agenda

<hr>

- what?
- why?
- how?
- use cases
- HTML basics
- case studies

## What?

<hr>

Extract data from websites

## Why?

<hr>

- data is available but not in a file or a database
- you cannot right click to save/download
- you cannot copy/paste contents into a file
- automation

## How?

<hr>

- fetch the page (HTML source code)
- parse/extract
- convert to tabular data structure
- process/explore

## Use Cases

<hr>

- contact scraping
- online price change monitoring
- price comparison
- review scraping
- real estate listing
- used car prices

## Case Studies

<hr>

- list of [RBI](https://rbi.org.in/) Governors
- list of top 50 websites
- top selling mobile phones on [Amazon India](https://www.amazon.in/mobile-phones/b?ie=UTF8&node=1389401031&ref_=nav_shopall_sbc_mobcomp_all_mobiles)
- IMDB top 50 movies

## Libraries

<hr>

```{r libs}
library(robotstxt)
library(rvest)
library(selectr)
library(xml2)
library(dplyr)
library(stringr)
library(forcats)
library(magrittr)
library(tidyr)
library(ggplot2)
```

## Things to keep in mind..

<hr>

- best for static & well structured HTML pages
- HTML code can change any time
- use API, if available
- do not overwhelm websites with requests
- review **robots.txt** file

## HTML Basics

<hr>

how to view the source code of a web page

## HTML Tags

<hr>

## Nested Tags

<hr>

## Tag Attributes

<hr>

## robotstxt

<hr>

## RBI Governors

<hr>

## robotstxt

<hr>

```{r rbi1}
paths_allowed(
  paths = c("https://en.wikipedia.org/wiki/List_of_Governors_of_Reserve_Bank_of_India")
)
```

## Read HTML

<hr>

```{r rbi2}
rbi_guv <- xml2::read_html("https://en.wikipedia.org/wiki/List_of_Governors_of_Reserve_Bank_of_India")
rbi_guv
```

## List of Governors

<hr>

```{r rbi3}
rbi_guv %>%
  html_nodes("table") %>%
  html_table() %>%
  extract2(2) -> profile
```

## List of Governors

<hr>

```{r rbi4}
profile
```

## Sort

<hr>

```{r rbi5}
profile %>%
  separate(`Term in office`, into = c("term", "days")) %>%
  select(Officeholder, term) %>%
  arrange(desc(as.numeric(term)))
```

## Backgrounds

<hr>

```{r rbi6}
profile %>%
  count(Background) 
```

## Backgrounds

<hr>

```{r rbi7}
profile %>%
  pull(Background) %>%
  fct_collapse(
    Bureaucrats = c("IAS officer", "ICS officer",
    "Indian Administrative Service (IAS) officer",
    "Indian Audit and Accounts Service officer",
    "Indian Civil Service (ICS) officer"),
    `No Info` = c(""),
    `RBI Officer` = c("Career Reserve Bank of India officer")
  ) %>%
  fct_count() %>%
  rename(background = f, count = n) -> backgrounds
```

## Backgrounds

<hr>

```{r rbi8}
backgrounds
```

## Backgrounds

<hr>

```{r rbi9, fig.align='center', fig.height=4}
backgrounds %>%
  ggplot() +
  geom_col(aes(background, count), fill = "blue") +
  xlab("Background") + ylab("Count") +
  ggtitle("Background of RBI Governors")
```

## Top 50 Websites

<hr>

## robotstxt

<hr>

```{r web1}
paths_allowed(
  paths = c("https://www.similarweb.com/top-websites")
)
```

## Read Web Page

<hr>

```{r web2}
top_sites <- read_html("https://www.similarweb.com/top-websites")
top_sites
```

## Website Ranking

<hr>

```{r web3}
top_sites %>%
  html_node("table") %>%
  html_table() -> ranking
```

## Website Ranking

```{r web4}
ranking
```

## Rename Columns

<hr>

```{r web5}
col_names <- c("ranking", "domain", "category", "change", "avg_time_on_site",
               "avg_pages_per_visit", "bounce_rate")
colnames(ranking) <- col_names
head(ranking)
```

## Top Categories

<hr>

```{r web6}
ranking %>%
  count(category)
```

## Best Selling Phones

<hr>

## robotstxt

<hr>

```{r phones1}
paths_allowed(
  paths = c("https://www.amazon.in/mobile-phones/b?ie=UTF8&node=1389401031&ref_=nav_shopall_sbc_mobcomp_all_mobiles")
)
```

## Read Web Page

<hr>

```{r phones2}
top_phones <- read_html("https://www.amazon.in/mobile-phones/b?ie=UTF8&node=1389401031&ref_=nav_shopall_sbc_mobcomp_all_mobiles")
top_phones
```

## Brand Name

<hr>

```{r phones3}
top_phones %>%
  html_nodes(".crwTitle a") %>%
  html_text()
```

## Rating

<hr>

```{r phones4}
top_phones %>%
  html_nodes(".crwProductDetail span .a-icon-alt") %>%
  html_text()
```

## Number of Reviews

<hr>

```{r phones5}
top_phones %>%
  html_nodes(".crwProductDetail span.a-size-small .a-link-normal") %>%
  html_text()
```

## Actual Price

<hr>

```{r phones6}
top_phones %>%
  html_nodes(".crwProductDetail .crwPrice .a-text-strike") %>%
  html_text()
```

## Discounted Price

<hr>

```{r phones7}
top_phones %>%
  html_nodes(".crwProductDetail .crwActualPrice") %>%
  html_text()
```

## IMDB Top 50

<hr>

## robotstxt

<hr>

```{r imdb1}
paths_allowed(
  paths = c("https://www.imdb.com/search/title?groups=top_250&sort=user_rating")
)
```

## Read Web Page

<hr>

```{r imdb2}
imdb <- read_html("https://www.imdb.com/search/title?groups=top_250&sort=user_rating")
imdb
```

## Title

<hr>

```{r imdb3}
imdb %>%
  html_nodes(".lister-item-content h3 a") %>%
  html_text() -> title

title
```

## Year of Release

<hr>

```{r imdb4}
imdb %>%
  html_nodes(".lister-item-content h3 .lister-item-year") %>%
  html_text() -> year

year
```

## Certificate

<hr>

```{r imdb5}
imdb %>%
  html_nodes(".lister-item-content p .certificate") %>%
  html_text() -> certificate

certificate
```

## Runtime

<hr>

```{r imdb6}
imdb %>%
  html_nodes(".lister-item-content p .runtime") %>%
  html_text() -> runtime

runtime
```

## Genre

<hr>

```{r imdb7}
imdb %>%
  html_nodes(".lister-item-content p .genre") %>%
  html_text() -> genre

genre
```

## Rating

<hr>

```{r imdb8}
imdb %>%
  html_nodes(".ratings-bar .ratings-imdb-rating") %>%
  html_attr("data-value") -> rating

rating
```

## Votes

<hr>

```{r imdb9}
imdb %>%
  html_nodes(xpath = '//meta[@itemprop="ratingCount"]') %>% 
  html_attr('content') -> votes

votes
```

## Revenue

<hr>

**Revenue not available for 31 and 47 movie number**

```{r imdb10}
imdb %>%
  html_nodes(xpath = '//span[@name="nv"]') %>%
  html_text() %>%
  str_extract(pattern = "^\\$.*") %>%
  na.omit() %>%
  as.character() %>%
  append(values = NA, after = 30) %>%
  append(values = NA, after = 46) -> revenue

revenue
```  

## {data-background="thankyou.png"}

