---
output: _output.yaml
---

## {data-background="viz_bar.png"}

## Agenda

<hr>

- create 
    - simple bar plot
    - stacked bar plot
    - grouped bar plot
- modify bar
    - direction
    - color
    - line color
    - width
    - labels
- modify axis range
- remove axes from the plot
- specify the line type of the X axes
- offset the Y axes
- modify legend

## Introduction

<hr>

- a bar plot represents data in rectangular bars 
- the length of the bars are proportional to the values they represent 
- bar plots can be either horizontal or vertical 
- the X axis of the plot represents the levels or the categories
- and the Y axis represents the frequency/count of the variable 

<!-- ## Univariate Bar Plot -->

<!-- <hr> -->

<!-- A univariate bar plot represents a single categorical variable. The categories in the variable are represented on the X axis and their frequencies on the Y axis. In the below example, the `cyl` variable from the `mtcars` data set is visualized using a bar plot. The categories or levels are `4`, `6` and `8` which represent the number of cylinders in the automobile and are represented on the X axis. The frequency for each type of cylinder is represented by the Y axis. -->

<!-- In R, bar plots can be created using either the `plot()` or `barplot()` function. The input to both the functions are different. In case of the `plot()` function, we can specify the variable but it must be converted to a factor variable. In case of the `barplot()` function, the input must be the count or frequency of the variable. The `table()` function can be used to generate the counts/frequency for a variable. Let us use both the functions to create the bar plot: -->

## Libraries

<hr>

```{r dplyr1, eval=FALSE}
library(readr)
library(dplyr)
library(magrittr)
```

```{r libs, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(readr)
library(dplyr)
library(magrittr)
```

## Data

<hr>

```{r import, eval=FALSE}
ecom <- read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/ecom.csv',
  col_types = list(col_factor(levels = c('Desktop', 'Mobile', 'Tablet')), 
  col_factor(levels = c(TRUE, FALSE)), col_factor(levels = c(TRUE, FALSE)), 
  col_factor(levels = c('Affiliates', 'Direct', 'Display', 'Organic', 'Paid', 'Referral', 'Social'))))
```

```{r show, echo=FALSE, eval=TRUE, message=FALSE}
ecom <- read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/ecom.csv',
  col_types = list(col_factor(levels = c('Desktop', 'Mobile', 'Tablet')), 
  col_factor(levels = c(TRUE, FALSE)), col_factor(levels = c(TRUE, FALSE)), 
  col_factor(levels = c('Affiliates', 'Direct', 'Display', 'Organic', 'Paid', 'Referral', 'Social'))))
ecom
```

## Data Dictionary

<hr>

Below is the description of the data set:

- device: device used to visit the website
- bouncers: whether visit was a bouncer (exit website from landing page)
- purchase: whether visitor purchased
- referrer: referrer website/search engine

## Using plot function

<hr>

```{r bar1, fig.align='center', fig.height=5, fig.width=5}
plot(ecom$device)
```

## Using barplot function

<hr>

```{r bar2, fig.align='center', fig.height=5, fig.width=5}
barplot(table(ecom$device))
```

## Data

<hr>

```{r bardata}
device_freq <- table(ecom$device)
device_freq
```

## Horizontal Bar Plot

<hr>

```{r bar3, fig.align='center', fig.height=5, fig.width=5}
barplot(device_freq, horiz = TRUE)
```

## Labels

<hr>

```{r bar8, fig.align='center', fig.height=5, fig.width=5}
barplot(device_freq, names.arg = c('Desktop', 'Mobile', 'Tablet'))
```

## Color

<hr>

```{r bar9, fig.align='center', fig.height=5, fig.width=5}
barplot(device_freq, col = 'blue')
```

## {data-background="stacked.png"}

## Data 

<hr>

```{r}
device_referrer <- table(ecom$device, ecom$referrer)
device_referrer
```

## Stacked Bar Plot

<hr>

```{r bar21, fig.align='center', fig.height=5, fig.width=5}
barplot(device_referrer)
```

## Color

<hr>

```{r bar22, fig.align='center', fig.height=5, fig.width=5}
barplot(device_referrer, col = c('blue', 'red', 'green'))
```

## Legend

<hr>

```{r bar23, fig.align='center', fig.height=5, fig.width=5}
barplot(device_referrer, col = c('blue', 'red', 'green'),
        main = 'Gears vs Cylinders', legend.text = TRUE,
        xlab = 'Accquisition Channel', ylab = 'Visitors')
```

## Grouped Bar Plot

<hr>

```{r bar24, fig.align='center', fig.height=5, fig.width=5}
barplot(device_referrer, col = c('blue', 'red', 'green'), beside = TRUE, 
        legend.text = TRUE, main = 'Device Distribution by Referrer Type',
        xlab = 'Accquisition Channel', ylab = 'Visitors')
```

## {data-background="aesthetics.png"}

## Bar Width (Equal Width)

<hr>

```{r bar4, fig.align='center', fig.height=5, fig.width=5}
barplot(device_freq, width = 2)
```

## Space Between Bars

<hr>

```{r bar7, fig.align='center', fig.height=5, fig.width=5}
barplot(device_freq, space = c(1, 1, 2))
```

## Border Color

<hr>

```{r bar12, fig.align='center', fig.height=5, fig.width=5}
barplot(device_freq, border = 'blue')
```

## Remove axes

<hr>

```{r bar15, fig.align='center', fig.height=5, fig.width=5}
barplot(device_freq, axes = FALSE)
```

## Axis Line Type

<hr>

```{r bar16, fig.align='center', fig.height=5, fig.width=5}
barplot(device_freq, axis.lty = 3)
```

## Offset Y Axes

<hr>

```{r bar17, fig.align='center', fig.height=5, fig.width=5}
barplot(device_freq, offset = 10)
```

## Axis Range

<hr>

```{r bar18, fig.align='center', fig.height=5, fig.width=5}
barplot(device_freq, ylim = c(0, 4000))
```

## Putting it all together...

<hr>

```{r bar19, fig.align='center', fig.height=5, fig.width=5}
barplot(device_freq, col = c('blue', 'red', 'green'),
        horiz = TRUE, width = c(1, 1, 2),
        names.arg = c('Desktop', 'Mobile', 'Tablet'),
        axis.lty = 2, offset = 10)
```

## Title & Axis Labels

<hr>

```{r bar20, fig.align='center', fig.height=5, fig.width=5}
barplot(device_freq, col = c('blue', 'red', 'green'), axis.lty = 2,
        width = c(2, 1, 0.5), names.arg = c('Desktop', 'Mobile', 'Tablet'), offset = 2)
title(main = 'Distribution of Devices',
      ylab = 'Visitors', xlab = 'Device')
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - combine introduction and univariate bar plot -->