---
output: _output.yaml
---

## {data-background="viz_bar.png"}

## Agenda

<hr>

- create 
    - simple bar plot
    - stacked bar plot
    - grouped bar plot
- modify bar
    - direction
    - color
    - line color
    - width
    - labels
- modify axis range
- remove axes from the plot
- specify the line type of the X axes
- offset the Y axes
- modify legend

## Introduction

<hr>

A bar plot represents data in rectangular bars. The length of the bars are proportional to the values they represent. Bar plots can be either horizontal or vertical. The X axis of the plot represents the levels or the categories and the Y axis represents the frequency/count of the variable. 

## Univariate Bar Plot

<hr>

A univariate bar plot represents a single categorical variable. The categories in the variable are represented on the X axis and their frequencies on the Y axis. In the below example, the `cyl` variable from the `mtcars` data set is visualized using a bar plot. The categories or levels are `4`, `6` and `8` which represent the number of cylinders in the automobile and are represented on the X axis. The frequency for each type of cylinder is represented by the Y axis.

In R, bar plots can be created using either the `plot()` or `barplot()` function. The input to both the functions are different. In case of the `plot()` function, we can specify the variable but it must be converted to a factor variable. In case of the `barplot()` function, the input must be the count or frequency of the variable. The `table()` function can be used to generate the counts/frequency for a variable. Let us use both the functions to create the bar plot:

## Using plot function

<hr>

```{r bar1, fig.align='center', fig.height=5, fig.width=5}
plot(as.factor(mtcars$cyl))
```

## Using barplot function

<hr>

```{r bar2, fig.align='center', fig.height=5, fig.width=5}
barplot(table(mtcars$cyl))
```

## Data

<hr>

```{r bardata}
cyl_freq <- table(mtcars$cyl)
cyl_freq
```

## Horizontal Bar Plot

<hr>

```{r bar3, fig.align='center', fig.height=5, fig.width=5}
barplot(cyl_freq, horiz = TRUE)
```

## Labels

<hr>

```{r bar8, fig.align='center', fig.height=5, fig.width=5}
barplot(cyl_freq, names.arg = c('Four', 'Six', 'Eight'))
```

## Color

<hr>

```{r bar9, fig.align='center', fig.height=5, fig.width=5}
barplot(cyl_freq, col = 'blue')
```

## {data-background="stacked.png"}

## Data 

<hr>

```{r}
table(mtcars$gear)
cyl_gear <- table(mtcars$cyl, mtcars$gear)
cyl_gear
```

## Stacked Bar Plot

<hr>

```{r bar21, fig.align='center', fig.height=5, fig.width=5}
barplot(cyl_gear)
```

## Color

<hr>

```{r bar22, fig.align='center', fig.height=5, fig.width=5}
barplot(cyl_gear, col = c('blue', 'red', 'green'))
```

## Legend

<hr>

```{r bar23, fig.align='center', fig.height=5, fig.width=5}
barplot(cyl_gear, col = c('blue', 'red', 'green'),
        main = 'Gears vs Cylinders', legend.text = TRUE,
        xlab = 'Number of Gears', ylab = 'Frequency')
```

## Grouped Bar Plot

<hr>

```{r bar24, fig.align='center', fig.height=5, fig.width=5}
barplot(cyl_gear, col = c('blue', 'red', 'green'), beside = TRUE, 
        legend.text = TRUE, main = 'Gears vs Cylinders',
        xlab = 'Number of Gears', ylab = 'Frequency')
```

## {data-background="aesthetics.png"}

## Bar Width (Equal Width)

<hr>

```{r bar4, fig.align='center', fig.height=5, fig.width=5}
barplot(cyl_freq, width = 2)
```

## Space Between Bars

<hr>

```{r bar7, fig.align='center', fig.height=5, fig.width=5}
barplot(cyl_freq, space = c(1, 1, 2))
```

## Border Color

<hr>

```{r bar12, fig.align='center', fig.height=5, fig.width=5}
barplot(cyl_freq, border = 'blue')
```

## Remove axes

<hr>

```{r bar15, fig.align='center', fig.height=5, fig.width=5}
barplot(cyl_freq, axes = FALSE)
```

## Axis Line Type

<hr>

```{r bar16, fig.align='center', fig.height=5, fig.width=5}
barplot(cyl_freq, axis.lty = 3)
```

## Offset Y Axes

<hr>

```{r bar17, fig.align='center', fig.height=5, fig.width=5}
barplot(cyl_freq, offset = 5)
```

## Axis Range

<hr>

```{r bar18, fig.align='center', fig.height=5, fig.width=5}
barplot(cyl_freq, ylim = c(0, 30))
```

## Putting it all together...

<hr>

```{r bar19, fig.align='center', fig.height=5, fig.width=5}
barplot(cyl_freq, col = c('blue', 'red', 'green'),
        horiz = TRUE, width = c(1, 1, 2),
        names.arg = c('Four', 'Six', 'Eight'),
        axis.lty = 2, offset = 2)
```

## Title & Axis Labels

<hr>

```{r bar20, fig.align='center', fig.height=5, fig.width=5}
barplot(cyl_freq, col = c('blue', 'red', 'green'), horiz = TRUE, axis.lty = 2,
        width = c(1, 1, 2), names.arg = c('Four', 'Six', 'Eight'), offset = 2)
title(main = 'Distribution of Cylinders',
      xlab = 'Frequency', ylab = 'Number of Cylinders')
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - combine introduction and univariate bar plot -->