---
output: _output.yaml
---

## {data-background="viz_box.png"}

## Agenda

<hr>

- create univariate/multivariate box plots
- interpret box plots
- create horizontal box plots
- detect outliers
- modify box color
- use formula to compare distributions of different variables
- use notches to compare medians

## Introduction

<hr>

- the box plot is a standardized way of displaying the distribution of data  
- box plots are useful for detecting outliers and for comparing distributions 
- it shows the shape, central tendancy and variability of the data

## Structure

<hr>

- the body of the boxplot consists of a "box" (hence, the name), which goes from the first quartile (Q1) to the third quartile (Q3)
- within the box, a vertical line is drawn at the Q2, the median of the data set
- two horizontal lines, called whiskers, extend from the front and back of the box 
- the front whisker goes from Q1 to the smallest non-outlier in the data set, and the back whisker goes from Q3 to the largest non-outlier
- if the data set includes one or more outliers, they are plotted separately as points on the chart

## {data-background="uboxplot.png"}

## Data

<hr>

```{r import, eval=FALSE}
daily_returns <- read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/tickers.csv')
```

```{r show, echo=FALSE, eval=TRUE, message=FALSE}
daily_returns <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/tickers.csv')
daily_returns
```

## Univariate Box Plot

<hr>

```{r box1, fig.align='center', fig.width=5, fig.height=5}
boxplot(daily_returns$AAPL)
```

## Horizontal Box Plot

<hr>

```{r box2, fig.align='center', fig.width=5, fig.height=5}
boxplot(daily_returns$AAPL, horizontal = TRUE)
```

## Color

<hr>

```{r box3, fig.align='center', fig.width=5, fig.height=5}
boxplot(daily_returns$AAPL, col = 'blue')
```

## Outliers

<hr>

```{r box8, fig.align='center', fig.width=5, fig.height=5}
boxplot(daily_returns$AAPL, range = 1, outline = TRUE)
```

## Outliers

<hr>

```{r box9, fig.align='center', fig.width=10, fig.height=7, echo=FALSE}
init <- par(no.readonly = TRUE)
par(mfrow = c(1, 2))
values <- c(TRUE, FALSE)
for (i in values) {
  boxplot(daily_returns$AAPL, outline = i, range = 1)
  mtext(paste("outline = ", i), side = 3, col = "blue")
}
```

## {data-background="bboxplot.png"}

## Data

<hr>

```{r import2, eval=FALSE}
tidy_returns <- read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/tidy_tickers.csv')
```

```{r show2, echo=FALSE, eval=TRUE, message=FALSE}
tidy_returns <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/tidy_tickers.csv')
tidy_returns
```

## Box Plot

<hr>

```{r box11, fig.align='center', fig.width=5, fig.height=5}
boxplot(tidy_returns$returns ~ tidy_returns$stock)
```

## Color

<hr>

```{r box14, fig.align='center', fig.width=5, fig.height=5}
boxplot(tidy_returns$returns ~ tidy_returns$stock, col = 'blue')
```

## Different Colors

<hr>

```{r box15, fig.align='center', fig.width=5, fig.height=5}
boxplot(tidy_returns$returns ~ tidy_returns$stock, 
        col = c('red', 'blue', 'yellow'))
```

## Compare Medians

<hr>

```{r box16a, fig.align='center', fig.width=5, fig.height=5}
boxplot(tidy_returns$returns ~ tidy_returns$stock, notch = TRUE,
        col = c('red', 'blue', 'yellow'))
```

## Compare Medians

<hr>

```{r box16, fig.align='center', fig.width=5, fig.height=5}
hsb <- read.table('https://stats.idre.ucla.edu/wp-content/uploads/2016/02/hsb2-2.csv', header=T, sep=",")
boxplot(hsb$read ~ hsb$female, notch = TRUE, 
        col = c('red', 'blue'))
```

## Border Color
 
<hr>

```{r box4, fig.align='center', fig.width=5, fig.height=5}
boxplot(daily_returns$AAPL, border = 'red')
```

## Range

<hr>

```{r box5, fig.align='center', fig.width=5, fig.height=5}
boxplot(daily_returns$AAPL, range = 0)
```

## Range

<hr>

```{r box6, fig.align='center', fig.width=5, fig.height=5}
boxplot(daily_returns$AAPL, range = 1)
```

## Range

<hr>

```{r box7, fig.align='center', fig.width=10, fig.height=7, echo=FALSE}
init <- par(no.readonly = TRUE)
par(mfrow = c(2, 2))
values <- c(0, 0.25, 0.63, 1)
for (i in values) {
  boxplot(daily_returns$AAPL, range = i)
  mtext(paste("range = ", i), side = 3, col = "blue")
}
```


## Varwidth

<hr>

```{r box10, fig.align='center', fig.width=10, fig.height=7, echo=FALSE}
init <- par(no.readonly = TRUE)
par(mfrow = c(1, 2))
values <- c(TRUE, FALSE)
for (i in values) {
  boxplot(daily_returns$AAPL, varwidth = i, range = 1)
  mtext(paste("varwidth = ", i), side = 3, col = "blue")
}
```


## Putting it all together..

<hr>

```{r box17, fig.align='center', fig.width=5, fig.height=5}
boxplot(tidy_returns$returns ~ tidy_returns$stock, range = 1, outline = TRUE, 
        col = c('red', 'blue', 'yellow'), main = 'Daily Returns', 
        ylab = 'Stock', xlab = 'Daily Returns',
        names = c('AAPL', 'AMZN', 'FB', 'GOOG', 'MSFT'))
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - improve introduction -->
<!--   - use image to explain structure of box plot -->