---
output: _output.yaml
---

## {data-background="viz_hist.png"}

## Agenda

<hr>

Learn to construct and use histograms to examine the underlying distribution of a continuous variable. Specifically

- create a bare bones histogram
- specify the number of bins/intervals
- represent frequency density on the Y axis
- add colors to the bars and the border
- add labels to the bars


## Introduction

<hr>

A histogram is a plot that can be used to examine the shape and spread of  continuous data. It looks very similar to a bar graph and can be used to detect outliers and skewness in data. The histogram graphically shows the following:

- center (location) of the data
- spread (dispersion) of the data
- skewness
- outliers
- presence of multiple modes

## Histograms

<hr>

To construct a histogram, the data is split into intervals called bins. The intervals may or may not be equal sized. For each bin, the number of data points that fall into it are counted (frequency). The Y axis of the histogram represents the frequency and the X axis represents the variable. 

## Normal Distribution

<hr>

```{r hist1, fig.align='center', fig.height=5, fig.width=5, echo=FALSE}
hist(rbeta(10000, 5, 5), ann = FALSE, col = 'blue')
```

## Skewed Distribution

<hr>

```{r hist2, fig.align='center', fig.width=10, fig.height=5, echo=FALSE}
init <- par(no.readonly = TRUE)
par(mfrow = c(1, 2))
hist(rbeta(10000, 2, 5), ann = FALSE, col = 'blue')
hist(rbeta(10000, 5, 2), ann = FALSE, col = 'blue')
```

## Histogram

<hr>

```{r hist3, fig.align='center', fig.height=5, fig.width=5, echo=FALSE}
hist(mtcars$mpg)
```

## Histogram

<hr>

```{r hist4, fig.align='center'}
h <- hist(mtcars$mpg)
```

## Histogram

<hr>

```{r hist30}
# display number of breaks
h$breaks

# frequency of the intervals
h$counts

# frequency density
h$density

# mid points of the intervals
h$mids

# varible name
h$xname

# whether intervals are of equal size
h$equidist
```

## Bins

<hr>

```{r hist5, fig.align='center', fig.height=5, fig.width=5}
hist(mtcars$mpg, breaks = 10)
```

## Bins 

<hr>

```{r hist6, fig.align='center', fig.width=10, fig.height=7, echo=FALSE}
init <- par(no.readonly = TRUE)
par(mfrow = c(2, 2))
values <- c(5, 10, 15, 20)
for (i in values) {
    hist(mtcars$mpg, breaks = i)
    mtext(paste("breaks = ", i), side = 3, col = "blue")
}
```

## Intervals

<hr>

```{r hist7, fig.align='center', fig.height=5, fig.width=5}
hist(mtcars$mpg, breaks = c(10, 18, 24, 30, 35))
```

## Frequency Density

<hr>

Frequency Density =  Relative Frequency / Class Width

Relative Frequency = Frequency / Total Observations

## Breaks

<hr>

```{r hist8, fig.align='center'}
h <- hist(mtcars$mpg, breaks = c(10, 18, 24, 30, 35))
```

## Frequency Density

<hr>

````{r hist31}
frequency <- h$counts
class_width <- c(8, 6, 6, 5)
rel_freq <- frequency / length(mtcars$mpg)
freq_density <- rel_freq / class_width
d <- data.frame(frequency = frequency, class_width = class_width, relative_frequency = rel_freq, frequency_density = freq_density)
d
```

## Frequency Density

<hr>

When multiplied by the class width, the product will always sum upto 1.

```{r hist9}
sum(d$frequency_density * d$class_width)
```

## Intervals

<hr>

We will learn more about frequency density in a bit. Before we end this section, we need to learn about one more way to specify the intervals of the histogram, algorithms. The `hist()` function allows us to specify the following algorithms:

- Sturges (default)
- Scott
- Freedman-Diaconis (FD)

## Intervals

<hr>

```{r hist10, fig.align='center', fig.width=10, fig.height=7, echo=FALSE}
init <- par(no.readonly = TRUE)
par(mfrow = c(1, 3))
values <- c("Sturges", "Scott", "FD")
for (i in values) {
    hist(mtcars$mpg, breaks = i)
    mtext(paste("algo = ", i), side = 3, col = "blue")
}
```

## Frequency Distribution II

<hr>

```{r hist11, fig.align='center', fig.width=10, fig.height=7, echo=FALSE}
init <- par(no.readonly = TRUE)
par(mfrow = c(1, 2))
values <- c(TRUE, FALSE)
for (i in values) {
    hist(mtcars$mpg, freq = i)
    mtext(paste("freq = ", i), side = 3, col = "blue")
}
```

## Probability

<hr>

```{r hist12, fig.align='center', fig.height=5, fig.width=5}
hist(mtcars$mpg, probability = TRUE)
```

## Color

<hr>

```{r hist13, fig.align='center', fig.height=5, fig.width=5}
hist(mtcars$mpg, col = 'blue')
```

## Border Color

<hr>

```{r hist16, fig.align='center', fig.height=5, fig.width=5}
hist(mtcars$mpg, border = c('red', 'blue', 'green', 'yellow', 'brown'))
```

## Labels

<hr>

```{r hist17, fig.align='center', fig.height=5, fig.width=5}
hist(mtcars$mpg, labels = TRUE)
```

## Labels

<hr>

```{r hist18, fig.align='center', fig.height=5, fig.width=5}
hist(mtcars$mpg, labels = c("6", "12", "8", "2", "4"))
```

## Title & Axis Labels

<hr>

```{r hist19, fig.align='center'}
hist(mtcars$mpg, labels = TRUE, prob = TRUE,
     ylim = c(0, 0.1), xlab = 'Miles Per Gallon',
     main = 'Distribution of Miles Per Gallon',
     col = rainbow(5))
```

## {data-background="thankyou.png"}

<!-- To do: -->
  <!-- - add intro slide -->
  <!-- - section for bins -->
  <!-- - image for frequency density -->
  <!-- - section for intervals -->
  <!-- - section for color -->