---
output: _output.yaml 
---

## {data-background="viz_legend.png"}

## Agenda

<hr>

Learn to add legend to a plot using the `legend()` function. Specifically

- position the legend within the plot
- modify the layout using `ncol` and `horiz` arguments
- add title using the `title.` set of arguments
- modify the appearance and position of the legend box
- modify the appearance of the text in the legend box
  

## Introduction

<hr>

```{r show, echo=FALSE, eval=TRUE, message=FALSE, results='hide'}
library(readr)
users <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/users_device.csv',
  col_types = list(col_date(format = "%m/%d/%y"), col_integer(),
                   col_integer(), col_integer()))
users
```

```{r compare, fig.align='center', fig.width= 7, fig.height=6, echo=FALSE}
init <- par(no.readonly = TRUE)
{plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
lines(users$Date, users$Mobile, type = "l", col = "red")
lines(users$Date, users$Tablet, type = "l", col = "green")}
{plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
lines(users$Date, users$Mobile, type = "l", col = "red")
lines(users$Date, users$Tablet, type = "l", col = "green")  
legend(x = as.Date('2017-10-13'), y = 3900, legend = c('Desktop', 'Mobile', 'Tablet'), lty = 1, 
       col = c('blue', 'red', 'green'), horiz = TRUE)}
```


## Data

<hr>

```{r import, eval=FALSE}
users <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/users_device.csv',
  col_types = list(col_date(format = "%m/%d/%y"), col_integer(),
                   col_integer(), col_integer()))
```

```{r show2, echo=FALSE, eval=TRUE, message=FALSE}
users
```


## Data Dictionary

<hr>

Below is the description of the data set:

- Date: date 
- Desktop: visitors from a desktop device
- Mobile: visitors from a mobile device
- Tablet: visitors from a tablet device

## Line Graph

<hr>

```{r line1, fig.align='center', fig.height=5, fig.width=7}
plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 3000), xlab = 'Date', ylab = 'Visitors')
lines(users$Date, users$Mobile, type = "l", col = "red")
lines(users$Date, users$Tablet, type = "l", col = "green")
```


## Legend Location

<hr>

- use x and y axis coordinates
- use keywords
    - top
    - bottom
    - left
    - right
    - center
    - bottomright
    - bottomleft
    - topright
    - topleft

## Legend Location

<hr>

```{r leg1, fig.align='center', fig.height=5, fig.width=5}
{plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
lines(users$Date, users$Mobile, type = "l", col = "red")
lines(users$Date, users$Tablet, type = "l", col = "green")  
legend(x = as.Date('2017-10-13'), y = 3900, legend = c('Desktop', 'Mobile', 'Tablet'))}
```

## Legend Location

<hr>

```{r leg2, fig.align='center', fig.width=10, fig.height=7, echo=FALSE}
init <- par(no.readonly = TRUE)
par(mfrow = c(3, 3))
values <- c('top', 'bottom', 'left', 'right', 'center',
            'bottomright', 'bottomleft', 'topright', 'topleft')
for (i in values) {
  plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
  lines(users$Date, users$Mobile, type = "l", col = "red")
  lines(users$Date, users$Tablet, type = "l", col = "green")  
  legend(x = i, legend = c('Desktop', 'Mobile', 'Tablet'))
  mtext(paste("x = ", i), side = 3, col = "blue")
}
```

## Lines

<hr>

```{r leg3, fig.align='center', fig.height=5, fig.width=5}
{plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
lines(users$Date, users$Mobile, type = "l", col = "red")
lines(users$Date, users$Tablet, type = "l", col = "green")  
legend(x = as.Date('2017-10-13'), y = 3900, legend = c('Desktop', 'Mobile', 'Tablet'), lty = 1, 
       col = c('blue', 'red', 'green'))}
```

## Points

<hr>

```{r leg4, fig.align='center', fig.height=5, fig.width=5}
{plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
points(users$Date, users$Desktop, pch = 2, col = 'blue')
lines(users$Date, users$Mobile, type = "l", col = "red")
points(users$Date, users$Mobile, pch = 2, col = 'red')
lines(users$Date, users$Tablet, type = "l", col = "green")
points(users$Date, users$Tablet, pch = 2, col = 'green')  
legend(x = as.Date('2017-10-13'), y = 3900, legend = c('Desktop', 'Mobile', 'Tablet'), lty = 1, 
       col = c('blue', 'red', 'green'), pch = 2)}
```

## Text Placement

<hr>

```{r leg5, fig.align='center', fig.height=4, fig.width=6}
{plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
lines(users$Date, users$Mobile, type = "l", col = "red")
lines(users$Date, users$Tablet, type = "l", col = "green")  
legend(x = as.Date('2017-10-13'), y = 3900, legend = c('Desktop', 'Mobile', 'Tablet'), lty = 1, 
       col = c('blue', 'red', 'green'), horiz = TRUE)}
```

## Text Placement

<hr>

```{r leg6, fig.align='center', fig.width= 6, fig.height=4}
{plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
lines(users$Date, users$Mobile, type = "l", col = "red")
lines(users$Date, users$Tablet, type = "l", col = "green")  
legend(x = as.Date('2017-10-13'), y = 3900, legend = c('Desktop', 'Mobile', 'Tablet'), lty = 1, 
       col = c('blue', 'red', 'green'), ncol = 2)}
```

## Text Placement

<hr>

```{r leg7, fig.align='center', fig.width=15, fig.height=10, echo=FALSE}
init <- par(no.readonly = TRUE)
par(mfrow = c(2, 2))
values <- c(1, 2, 3, 4)
for (i in values) {
  plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
lines(users$Date, users$Mobile, type = "l", col = "red")
lines(users$Date, users$Tablet, type = "l", col = "green")  
legend(x = as.Date('2017-10-13'), y = 3900, legend = c('Desktop', 'Mobile', 'Tablet'), lty = 1, 
       col = c('blue', 'red', 'green'), ncol = i)
  mtext(paste("ncol = ", i), side = 3, col = "blue")
}
```

## Title

<hr>

```{r leg8, fig.align='center', fig.height=4, fig.width=6}
{plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
lines(users$Date, users$Mobile, type = "l", col = "red")
lines(users$Date, users$Tablet, type = "l", col = "green")  
legend(x = as.Date('2017-10-13'), y = 3900, legend = c('Desktop', 'Mobile', 'Tablet'), lty = 1, 
       col = c('blue', 'red', 'green'), title = 'Devices', horiz = TRUE)}
```

## Title Color

<hr>

```{r leg9, fig.align='center', fig.height=4, fig.width=6}
{plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
lines(users$Date, users$Mobile, type = "l", col = "red")
lines(users$Date, users$Tablet, type = "l", col = "green")  
legend(x = as.Date('2017-10-13'), y = 3900, legend = c('Desktop', 'Mobile', 'Tablet'), lty = 1, 
       col = c('blue', 'red', 'green'), title = 'Devices', title.col = 'red', horiz = TRUE)}
```

## Title Position

<hr>

```{r leg10, fig.align='center', fig.height=4, fig.width=6}
{plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
lines(users$Date, users$Mobile, type = "l", col = "red")
lines(users$Date, users$Tablet, type = "l", col = "green")  
legend(x = as.Date('2017-10-13'), y = 3900, legend = c('Desktop', 'Mobile', 'Tablet'), lty = 1, 
       col = c('blue', 'red', 'green'), title = 'Devices', title.adj = 0.1, horiz = TRUE)}
```

## Title Position

<hr>

```{r leg11, fig.align='center', fig.width=15, fig.height=10, echo=FALSE}
init <- par(no.readonly = TRUE)
par(mfrow = c(2, 2))
values <- c(0.2, 0.4, 0.6, 0.8)
for (i in values) {
  plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
  lines(users$Date, users$Mobile, type = "l", col = "red")
  lines(users$Date, users$Tablet, type = "l", col = "green")  
  legend(x = as.Date('2017-10-13'), y = 3900, legend = c('Desktop', 'Mobile', 'Tablet'), lty = 1, 
       col = c('blue', 'red', 'green'), title = 'Devices', title.adj = i, horiz = TRUE)
  mtext(paste("title.adj = ", i), side = 3, col = "blue")
}
```

## Box Appearance

<hr>

```{r kable, echo=FALSE}
option <- c('Box Type', 'Background Color', 'Border Line Type',
            'Border Line Width', 'Border Line Color', 
            'Horizontal Justification', 'Vertical Justification',
            'Text Color', 'Text Font')
argument <- c('bty', 'bg', 'box.lty', 'box.lwd', 'box.col',
              'xjust', 'yjust', 'text.col', 'text.font')
values <- c('o, n', 'blue, #0000ff', '1:5', '0.5, 1, 1.5',
            'blue, #0000ff', '0:1', '0:1', 'blue, #0000ff',
            '1:5')
legend <- data.frame(option, argument, values)
knitr::kable(legend, format = "markdown", caption = "Legend Box Options")
```

## Box Type

<hr>

```{r leg12, fig.align='center', fig.width=10, fig.height=5, echo=FALSE}
init <- par(no.readonly = TRUE)
par(mfrow = c(1, 2))
values <- c('o', 'n')
for (i in values) {
  plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
  lines(users$Date, users$Mobile, type = "l", col = "red")
  lines(users$Date, users$Tablet, type = "l", col = "green")  
  legend(x = as.Date('2017-10-13'), y = 3900, legend = c('Desktop', 'Mobile', 'Tablet'), lty = 1, 
       col = c('blue', 'red', 'green'), title = 'Devices', bty = i, horiz = TRUE)
  mtext(paste("bty = ", i), side = 3, col = "blue")
}
```

## Background Color

<hr>

```{r leg13, fig.align='center', fig.height=4, fig.width=6}
{plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
lines(users$Date, users$Mobile, type = "l", col = "red")
lines(users$Date, users$Tablet, type = "l", col = "green")  
legend(x = as.Date('2017-10-13'), y = 3900, legend = c('Desktop', 'Mobile', 'Tablet'), lty = 1, 
       col = c('blue', 'red', 'green'), title = 'Devices', bg = '#ffff66', horiz = TRUE)}
```

## Border

<hr>

The following arguments can be used to modify the border of the legend box:

- `box.lty`: line type
- `box.lwd`: line width
- `box.col`: color

## Border

<hr>

```{r leg14, fig.align='center', fig.height=4, fig.width=6}
{plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
lines(users$Date, users$Mobile, type = "l", col = "red")
lines(users$Date, users$Tablet, type = "l", col = "green")  
legend(x = as.Date('2017-10-13'), y = 3900, legend = c('Desktop', 'Mobile', 'Tablet'), lty = 1, 
       col = c('blue', 'red', 'green'), title = 'Devices', horiz = TRUE,
       box.lty = 3, box.lwd = 1.5, box.col = 'green')}
```

## Justification

<hr>

The `xjust` and `yjust` arguments can be used to position the legend relative to the **X** and **Y** axis respectively. Listed below is the value and the respective justification:

- `0`: left justified
- `0.5`: centered
- `1`: right justified

## Horizontal Justification

<hr>
 
```{r leg15, fig.align='center', fig.width=10, fig.height=7, echo=FALSE}
init <- par(no.readonly = TRUE)
par(mfrow = c(1, 3))
values <- c(0, 0.5, 1)
for (i in values) {
  plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
  lines(users$Date, users$Mobile, type = "l", col = "red")
  lines(users$Date, users$Tablet, type = "l", col = "green")  
  legend(x = as.Date('2017-10-13'), y = 3900, legend = c('Desktop', 'Mobile', 'Tablet'), lty = 1, 
       col = c('blue', 'red', 'green'), title = 'Devices', xjust = i, horiz = TRUE)
  mtext(paste("xjust = ", i), side = 3, col = "blue")
}
```

## Vertical Justification

<hr>

```{r leg16, fig.align='center', fig.width=10, fig.height=7, echo=FALSE}
init <- par(no.readonly = TRUE)
par(mfrow = c(1, 3))
values <- c(0, 0.5, 1)
for (i in values) {
  plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
  lines(users$Date, users$Mobile, type = "l", col = "red")
  lines(users$Date, users$Tablet, type = "l", col = "green")  
  legend(x = as.Date('2017-10-13'), y = 3900, legend = c('Desktop', 'Mobile', 'Tablet'), lty = 1, 
       col = c('blue', 'red', 'green'), title = 'Devices', yjust = i, horiz = TRUE)
  mtext(paste("xjust = ", i), side = 3, col = "blue")
}
```


## Text Appearance

<hr>

```{r leg17, fig.align='center', fig.height=4, fig.width=6}
{plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 4000), xlab = 'Date', ylab = 'Visitors')
lines(users$Date, users$Mobile, type = "l", col = "red")
lines(users$Date, users$Tablet, type = "l", col = "green")  
legend(x = as.Date('2017-10-13'), y = 3900, legend = c('Desktop', 'Mobile', 'Tablet'), lty = 1, 
       col = c('blue', 'red', 'green'), title = 'Devices', 
       text.col = 'green', text.font = 3, horiz = TRUE)}
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - change line graph data -->
<!--   - image for box appearance -->