---
output: _output.yaml
---

## {data-background="thankyou.png"}

## Agenda

<hr>

- create line plots
- add color to lines
- modify line type/style
- modify line width
- add points to the lines
- modify axis range
- add additional lines to the plot

## Data

<hr>

```{r airpas}
head(AirPassengers)
```

## Line Chart

<hr>

```{r line1, fig.align='center', fig.height=5, fig.width=5}
data <- head(AirPassengers)
plot(data, type = 'l')
```

## Lines & Points

<hr>

```{r line2, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'b')
```

## Overplotted

<hr>

```{r line3, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'o')
```

## Lines with Breaks

<hr>

```{r line4, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'c')
```

## Line Color

<hr>

```{r line5, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'l', col = 'blue')
```

## Color

<hr>

```{r line6, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'b', col = 'blue')
```

## Line Type

<hr>

```{r line9, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'l', lty = 3)
```

## Line Types

<hr>

- 1: solid
- 2: dashed
- 3: dotted
- 4: dotdash
- 5: longdash
- 6: twodash

## Line Types

<hr>

```{r lin7, echo=FALSE, fig.align='center', fig.height=7, fig.width=10}
init <- par(no.readonly = TRUE)
par(mfrow = c(3, 2))
values <- 1:6
for (i in values) {
    plot(data, type = "l", lty = i, col ="red")
    mtext(paste("lty = ", i), side = 3, col = "blue", line = 1)
}
```

## Line Types

<hr>

```{r lin8, echo=FALSE, fig.align='center', fig.height=7, fig.width=10}
init <- par(no.readonly = TRUE)
par(mfrow = c(3, 2))
values <- c("solid", "dashed", "dotted", "dotdash", "longdash", "twodash")
for (i in values) {
    plot(data, type = "l", lty = i, col ="red")
    mtext(paste("lty = ", i), side = 3, col = "blue", line = 1)
}
```

## Line Width

<hr>

```{r line10, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'l', lwd = 2.5)
```

## Line Width

<hr>

```{r lin11, echo=FALSE, fig.align='center', fig.height=7, fig.width=10}
init <- par(no.readonly = TRUE)
par(mfrow = c(2, 2))
values <- c(0.5, 1, 1.5, 2)
for (i in values) {
    plot(data, type = "l", lwd = i, col ="red")
    mtext(paste("lwd = ", i), side = 3, col = "blue", line = 1)
}
```

## Enhance Points

<hr>

```{r line12, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'b', pch = 23, col = 'red', cex = 1.5)
```

## Add Points

<hr>

```{r line13, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'l', col = 'red')
points(data, pch = 23, col = 'blue', bg = 'green', cex = 1.5)
```

## Additional Lines

<hr>

```{r line14, fig.align='center', fig.height=5, fig.width=5}
data1 <- c(7.2, 7.6, 6.8, 6.5, 7)
data2 <- c(6.8, 7.2, 7.8, 7, 6.2)
plot(data1, type = "b", col = "blue")
lines(data2, type = "b", col = "red")
```

## Additional Lines - Modify Axis Range

<hr>

```{r line15, fig.align='center', fig.height=5, fig.width=5}
plot(data1, type = "b", col = "blue", ylim = c(5, 9))
lines(data2, type = "b", col = "red")
```

## Line Plot

<hr>

```{r line16, fig.align='center', fig.height=5, fig.width=5}
plot(data1, type = "b", col = "blue", 
     ylim = c(5, 9), main = 'Line Graph',
     xlab = 'Index', ylab = 'Data')
lines(data2, type = "b", col = "red")
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - change data -->