---
title: "Line Charts"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

<style type="text/css">
    .reveal p img {
    align: middle;
  }
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: none;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
    font-size: 25px;
  }
  .reveal h1 {
    font-size: 24px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal p img {
    align: middle;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

- create line plots
- add color to lines
- modify line type/style
- modify line width
- add points to the lines
- modify axis range
- add additional lines to the plot

## Data

<hr>

```{r airpas}
head(AirPassengers)
```

## Line Chart

<hr>

```{r line1, fig.align='center', fig.height=5, fig.width=5}
data <- head(AirPassengers)
plot(data, type = 'l')
```

## Lines & Points

<hr>

```{r line2, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'b')
```

## Overplotted

<hr>

```{r line3, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'o')
```

## Lines with Breaks

<hr>

```{r line4, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'c')
```

## Line Color

<hr>

```{r line5, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'l', col = 'blue')
```

## Color

<hr>

```{r line6, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'b', col = 'blue')
```

## Line Type

<hr>

```{r line9, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'l', lty = 3)
```

## Line Types

<hr>

- 1: solid
- 2: dashed
- 3: dotted
- 4: dotdash
- 5: longdash
- 6: twodash

## Line Types

<hr>

```{r lin7, echo=FALSE, fig.align='center', fig.height=7, fig.width=10}
init <- par(no.readonly = TRUE)
par(mfrow = c(3, 2))
values <- 1:6
for (i in values) {
    plot(data, type = "l", lty = i, col ="red")
    mtext(paste("lty = ", i), side = 3, col = "blue", line = 1)
}
```

## Line Types

<hr>

```{r lin8, echo=FALSE, fig.align='center', fig.height=7, fig.width=10}
init <- par(no.readonly = TRUE)
par(mfrow = c(3, 2))
values <- c("solid", "dashed", "dotted", "dotdash", "longdash", "twodash")
for (i in values) {
    plot(data, type = "l", lty = i, col ="red")
    mtext(paste("lty = ", i), side = 3, col = "blue", line = 1)
}
```

## Line Width

<hr>

```{r line10, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'l', lwd = 2.5)
```

## Line Width

<hr>

```{r lin11, echo=FALSE, fig.align='center', fig.height=7, fig.width=10}
init <- par(no.readonly = TRUE)
par(mfrow = c(2, 2))
values <- c(0.5, 1, 1.5, 2)
for (i in values) {
    plot(data, type = "l", lwd = i, col ="red")
    mtext(paste("lwd = ", i), side = 3, col = "blue", line = 1)
}
```

## Enhance Points

<hr>

```{r line12, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'b', pch = 23, col = 'red', cex = 1.5)
```

## Add Points

<hr>

```{r line13, fig.align='center', fig.height=5, fig.width=5}
plot(data, type = 'l', col = 'red')
points(data, pch = 23, col = 'blue', bg = 'green', cex = 1.5)
```

## Additional Lines

<hr>

```{r line14, fig.align='center', fig.height=5, fig.width=5}
data1 <- c(7.2, 7.6, 6.8, 6.5, 7)
data2 <- c(6.8, 7.2, 7.8, 7, 6.2)
plot(data1, type = "b", col = "blue")
lines(data2, type = "b", col = "red")
```

## Additional Lines - Modify Axis Range

<hr>

```{r line15, fig.align='center', fig.height=5, fig.width=5}
plot(data1, type = "b", col = "blue", ylim = c(5, 9))
lines(data2, type = "b", col = "red")
```

## Line Plot

<hr>

```{r line16, fig.align='center', fig.height=5, fig.width=5}
plot(data1, type = "b", col = "blue", 
     ylim = c(5, 9), main = 'Line Graph',
     xlab = 'Index', ylab = 'Data')
lines(data2, type = "b", col = "red")
```

## Questions

<hr>

## Thank You

<hr>