---
output: _output.yaml
---

## {data-background="viz_line.png"}

## Agenda

<hr>

- create line plots
- add color to lines
- modify line type/style
- modify line width
- add points to the lines
- modify axis range
- add additional lines to the plot

## Data

<hr>

```{r import, eval=FALSE}
users <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/users_device.csv',
  col_types = list(col_date(format = "%m/%d/%y"), col_integer(),
                   col_integer(), col_integer()))
```

```{r show, echo=FALSE, eval=TRUE, message=FALSE}
library(readr)
users <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/users_device.csv',
  col_types = list(col_date(format = "%m/%d/%y"), col_integer(),
                   col_integer(), col_integer()))
users
```


## Data Dictionary

<hr>

Below is the description of the data set:

- Date: date 
- Desktop: visitors from a desktop device
- Mobile: visitors from a mobile device
- Tablet: visitors from a tablet device

## Line Chart

<hr>

```{r line1, fig.align='center', fig.height=5, fig.width=7}
plot(users$Date, users$Desktop, type = 'l', main = 'Daily Visitors',
     xlab = 'Date', ylab = 'Visitors')
```

## Lines & Points

<hr>

```{r line2, fig.align='center', fig.height=5, fig.width=7}
plot(users$Date, users$Desktop, type = 'b', main = 'Daily Visitors',
     xlab = 'Date', ylab = 'Visitors')
```

## Overplotted

<hr>

```{r line3, fig.align='center', fig.height=5, fig.width=7}
plot(users$Date, users$Desktop, type = 'o', main = 'Daily Visitors',
     xlab = 'Date', ylab = 'Visitors')
```

## Lines with Breaks

<hr>

```{r line4, fig.align='center', fig.height=5, fig.width=7}
plot(users$Date, users$Desktop, type = 'c', main = 'Daily Visitors',
     xlab = 'Date', ylab = 'Visitors')
```

## Line Color

<hr>

```{r line5, fig.align='center', fig.height=5, fig.width=7}
plot(users$Date, users$Desktop, type = 'l', col = 'blue', main = 'Daily Visitors',
     xlab = 'Date', ylab = 'Visitors')
```

## Color

<hr>

```{r line6, fig.align='center', fig.height=5, fig.width=7}
plot(users$Date, users$Desktop, type = 'b', col = 'blue', main = 'Daily Visitors',
     xlab = 'Date', ylab = 'Visitors')
```

## Line Type

<hr>

```{r line9, fig.align='center', fig.height=5, fig.width=7}
plot(users$Date, users$Desktop, type = 'l', lty = 3, main = 'Daily Visitors',
     xlab = 'Date', ylab = 'Visitors')
```

## Line Types

<hr>

- 1: solid
- 2: dashed
- 3: dotted
- 4: dotdash
- 5: longdash
- 6: twodash

## Line Types

<hr>

```{r lin7, echo=FALSE, fig.align='center', fig.height=7, fig.width=10}
init <- par(no.readonly = TRUE)
par(mfrow = c(3, 2))
values <- 1:6
for (i in values) {
    plot(users$Date, users$Desktop, type = "l", lty = i, col ="red",
     xlab = 'Date', ylab = 'Visitors')
    mtext(paste("lty = ", i), side = 3, col = "blue", line = 1)
}
```

## Line Types

<hr>

```{r lin8, echo=FALSE, fig.align='center', fig.height=7, fig.width=10}
init <- par(no.readonly = TRUE)
par(mfrow = c(3, 2))
values <- c("solid", "dashed", "dotted", "dotdash", "longdash", "twodash")
for (i in values) {
    plot(users$Date, users$Desktop, type = "l", lty = i, col ="red",
     xlab = 'Date', ylab = 'Visitors')
    mtext(paste("lty = ", i), side = 3, col = "blue", line = 1)
}
```

## Line Width

<hr>

```{r line10, fig.align='center', fig.height=5, fig.width=7}
plot(users$Date, users$Desktop, type = 'l', lwd = 2.5,
     xlab = 'Date', ylab = 'Visitors')
```

## Line Width

<hr>

```{r lin11, echo=FALSE, fig.align='center', fig.height=7, fig.width=10}
init <- par(no.readonly = TRUE)
par(mfrow = c(2, 2))
values <- c(0.5, 1, 1.5, 2)
for (i in values) {
    plot(users$Date, users$Desktop, type = "l", lwd = i, col ="red", main = 'Daily Visitors',
     xlab = 'Date', ylab = 'Visitors')
    mtext(paste("lwd = ", i), side = 3, col = "blue", line = 1)
}
```

## Enhance Points

<hr>

```{r line12, fig.align='center', fig.height=5, fig.width=7}
plot(users$Date, users$Desktop, type = 'b', pch = 23, col = 'red', cex = 1.5, 
	main = 'Daily Visitors', xlab = 'Date', ylab = 'Visitors')
```

## Add Points

<hr>

```{r line13, fig.align='center', fig.height=5, fig.width=7}
plot(users$Date, users$Desktop, type = 'l', col = 'red', 
	main = 'Daily Visitors', xlab = 'Date', ylab = 'Visitors')
points(users$Date, users$Desktop, pch = 23, col = 'blue', bg = 'green', cex = 1.5)
```

## Additional Lines

<hr>

```{r line14, fig.align='center', fig.height=5, fig.width=7}
plot(users$Date, users$Desktop, type = "l", col = "blue", 
	main = 'Daily Visitors', xlab = 'Date', ylab = 'Visitors')
lines(users$Date, users$Mobile, type = "l", col = "red")
lines(users$Date, users$Tablet, type = "l", col = "green")
```

## Additional Lines - Modify Axis Range

<hr>

```{r line15, fig.align='center', fig.height=5, fig.width=7}
plot(users$Date, users$Desktop, type = "l", col = "blue", ylim = c(0, 3000), 
	main = 'Daily Visitors', xlab = 'Date', ylab = 'Visitors')
lines(users$Date, users$Mobile, type = "l", col = "red")
lines(users$Date, users$Tablet, type = "l", col = "green")
```

## Putting it all together...

<hr>

```{r line16, fig.align='center', fig.height=5, fig.width=7}
plot(users$Date, users$Desktop, type = "l", col = "blue", main = 'Daily Visitors',
     ylim = c(0, 3000), xlab = 'Date', ylab = 'Visitors')
lines(users$Date, users$Mobile, type = "l", col = "red")
lines(users$Date, users$Tablet, type = "l", col = "green")
```

## {data-background="thankyou.png"}

