---
output: _output.yaml
---

## {data-background="viz_scatter.png"}

## Agenda

<hr>

- create scatter plot
- modify symbol
    - color
    - shape 
    - and shape
- add horizontal & vertical lines
- fit regression line

## Intro

<hr>

A scatter plot displays the relationship between two continuous variables. In ggplot2, we can build a scatter plot using `geom_point()`. Scatterplots can show you visually 

- the strength of the relationship between the variables 
- the direction of the relationship between the variables 
- and whether outliers exist


## Scatter Plot

```{r scatter1, fig.align='center', fig.height=5, fig.width=5}
plot(mtcars$disp, mtcars$mpg)
```

## Title & Axis Labels

<hr>

```{r scatter2, fig.align='center', fig.height=5, fig.width=5}
plot(mtcars$disp, mtcars$mpg,
     main = 'Displacement vs Miles Per Gallon',
     xlab = 'Displacement', ylab = 'Miles Per Gallon')
```

## Shape

<hr>

```{r scatter-shape, fig.align='center', fig.width=5, fig.height=5}
plot(mtcars$disp, mtcars$mpg, pch = 6)
```

## Shape 

<hr>

```{r scatter-pch, echo=FALSE, fig.align='center', fig.width=10, fig.height=7}
init <- par(no.readonly = TRUE)
par(mfrow = c(2, 2))
values <- c(4, 10, 16, 22)
for (i in values) {
    plot(mtcars$disp, mtcars$mpg, pch = i, cex = 1.5)
    mtext(paste("pch = ", i), side = 3, col = "blue")
}
```

## Shape

<hr>

```{r scatter-pch2, fig.align='center', fig.width=5, fig.height=5}
plot(mtcars$disp, mtcars$mpg, pch = nlevels(factor(mtcars$cyl)))
```

## Shape

<hr>

```{r scatter-pch3, fig.align='center', fig.width=5, fig.height=5}
plot(mtcars$disp, mtcars$mpg, pch = unclass(mtcars$cyl))
```

## Size

<hr>

```{r scatter-size, fig.align='center', fig.width=5, fig.height= 5}
plot(mtcars$disp, mtcars$mpg, cex = 1.5)
```

## Size

<hr>

```{r scatter-cex, echo=FALSE, fig.align='center', fig.width=10, fig.height=7}
init <- par(no.readonly = TRUE)
par(mfrow = c(2, 2))
values <- c(0.5, 1, 1.5, 2)
for (i in values) {
    plot(mtcars$disp, mtcars$mpg, pch = 6, cex = i)
    mtext(paste("cex = ", i), side = 3, col = "blue")
}
```

## Color

<hr>

```{r col1, fig.align='center', fig.width=5, fig.height=5}
plot(mtcars$disp, mtcars$mpg, pch = 5, col = 'blue', bg = 'red')
```

## Color

<hr>

```{r col2, fig.align='center', fig.width=5, fig.height=5}
plot(mtcars$disp, mtcars$mpg, pch = 24, col = 'red', bg = 'blue')
```

## Color

<hr>

```{r col3, fig.align='center', fig.width=5, fig.height=5}
plot(mtcars$disp, mtcars$mpg, pch = 5, col = factor(mtcars$cyl))
```

## Horizontal Line

<hr>

```{r hline, fig.align='center', fig.width=5, fig.height=5}
plot(mtcars$disp, mtcars$mpg)
abline(h = 20)
```

## Vertical Line

<hr>

```{r vline, fig.align='center', fig.width=5, fig.height=5}
plot(mtcars$disp, mtcars$mpg)
abline(v = 200)
```

## Regression Line (Intercept & Slope)

<hr>

```{r rline1, fig.align='center', fig.width=5, fig.height=5}
plot(mtcars$disp, mtcars$mpg)
abline(a = 29.59, b = -0.0412)
```

## Regression Line (Coefficients)

<hr>

```{r rline2, fig.align='center', fig.width=5, fig.height=5}
plot(mtcars$disp, mtcars$mpg)
abline(c(29.59, -0.0412))
```

## Regression Line (Using Model)

<hr>

```{r rline3, fig.align='center', fig.width=5, fig.height=5}
plot(mtcars$disp, mtcars$mpg)
model <- lm(mpg ~ disp, data = mtcars)
abline(coef(model))
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - show how to fit regression line -->