---
title: "Scatter Plot"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

<style type="text/css">
    .reveal p img {
    align: middle;
  }
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: none;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
    font-size: 25px;
  }
  .reveal h1 {
    font-size: 24px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal p img {
    align: middle;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

- create scatter plot
- modify symbol
    - color
    - shape 
    - and shape


## Scatter Plot

```{r scatter1, fig.align='center', fig.height=5, fig.width=5}
plot(mtcars$disp, mtcars$mpg)
```

## Title & Axis Labels

<hr>

```{r scatter2, fig.align='center', fig.height=5, fig.width=5}
plot(mtcars$disp, mtcars$mpg,
     main = 'Displacement vs Miles Per Gallon',
     xlab = 'Displacement', ylab = 'Miles Per Gallon')
```

## Shape

<hr>

```{r scatter-shape, fig.align='center', fig.width=5, fig.height=5}
plot(mtcars$disp, mtcars$mpg, pch = 6)
```

## Shape 

<hr>

```{r scatter-pch, echo=FALSE, fig.align='center', fig.width=10, fig.height=7}
init <- par(no.readonly = TRUE)
par(mfrow = c(2, 2))
values <- c(4, 10, 16, 22)
for (i in values) {
    plot(mtcars$disp, mtcars$mpg, pch = i, cex = 1.5)
    mtext(paste("pch = ", i), side = 3, col = "blue")
}
```

## Shape

<hr>

```{r scatter-pch2, fig.align='center', fig.width=5, fig.height=5}
plot(mtcars$disp, mtcars$mpg, pch = nlevels(factor(mtcars$cyl)))
```

## Shape

<hr>

```{r scatter-pch3, fig.align='center', fig.width=5, fig.height=5}
plot(mtcars$disp, mtcars$mpg, pch = unclass(mtcars$cyl))
```

## Size

<hr>

```{r scatter-size, fig.align='center', fig.width=5, fig.height= 5}
plot(mtcars$disp, mtcars$mpg, cex = 1.5)
```

## Size

<hr>

```{r scatter-cex, echo=FALSE, fig.align='center', fig.width=10, fig.height=7}
init <- par(no.readonly = TRUE)
par(mfrow = c(2, 2))
values <- c(0.5, 1, 1.5, 2)
for (i in values) {
    plot(mtcars$disp, mtcars$mpg, pch = 6, cex = i)
    mtext(paste("cex = ", i), side = 3, col = "blue")
}
```

## Color

<hr>

```{r col1, fig.align='center', fig.width=5, fig.height=5}
plot(mtcars$disp, mtcars$mpg, pch = 5, col = 'blue', bg = 'red')
```

## Color

<hr>

```{r col2, fig.align='center', fig.width=5, fig.height=5}
plot(mtcars$disp, mtcars$mpg, pch = 24, col = 'red', bg = 'blue')
```

## Color

<hr>

```{r col3, fig.align='center', fig.width=5, fig.height=5}
plot(mtcars$disp, mtcars$mpg, pch = 5, col = factor(mtcars$cyl))
```

## Questions

<hr>

## Thank You

<hr>