---
output: _output.yaml 
---

## {data-background="viz_title.png"}

## Agenda

<hr>

- Learn to add
    - title
    - subtitle
    - axis Labels

- learn to modify
    - axis range

## Syntax

<hr>

|Feature|Argument|Value|Example|
|:-------|:--------|:-----|:-------|
|Title  |main    |string|"Scatter Plot"|
|Subtitle  |sub    |string|"Displacement vs Miles Per Gallon"|
|X Axis Label  |xlab    |string|"Displacement"|
|Y Axis Label  |ylab    |string|"Miles Per Gallon"|


## Title

<hr>

```{r title, fig.align='center',fig.height=5,fig.width=5}
plot(mtcars$disp, mtcars$mpg, 
     main = 'Displacement vs Miles Per Gallon')
```

## Subtitle

<hr>

```{r subtitle, fig.align='center',fig.height=5,fig.width=5}
plot(mtcars$disp, mtcars$mpg, 
     main = 'Displacement vs Miles Per Gallon',
     sub = 'Mileage is inversely related to Displacement')
```

## Axis Labels

<hr>

```{r axis, fig.align='center',fig.height=5,fig.width=5}
plot(mtcars$disp, mtcars$mpg, 
     main = 'Displacement vs Miles Per Gallon',
     sub = 'Mileage is inversely related to Displacement',
     xlab = 'Displacement', ylab = 'Miles Per Gallon')
```

## title()

<hr>

```{r title1, fig.align='center',fig.height=5,fig.width=5}
plot(mtcars$disp, mtcars$mpg) 

title(main = 'Displacement vs Miles Per Gallon',
      sub = 'Mileage is inversely related to Displacement',
      xlab = 'Displacement', ylab = 'Miles Per Gallon')
```

## Annotate

<hr>

```{r title2, fig.align='center',fig.height=5,fig.width=5}
plot(mtcars$disp, mtcars$mpg, ann = FALSE) 

title(main = 'Displacement vs Miles Per Gallon',
      sub = 'Mileage is inversely related to Displacement',
      xlab = 'Displacement', ylab = 'Miles Per Gallon')
```

## Axis Range

<hr>

```{r axisrange, fig.align='center',fig.height=5,fig.width=5}
plot(mtcars$disp, mtcars$mpg, 
     xlim = c(0, 600), ylim = c(0, 50))
```

## {data-background="thankyou.png"}

<!-- To do: -->
<!--   - add intro slide -->
<!--   - change syntax image -->
  