---
title: "Title & Axis Labels"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
    fig_height: 5
    fig_width: 6
---

<style type="text/css">
    .reveal p img {
    align: middle;
  }
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: none;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
    font-size: 25px;
  }
  .reveal h1 {
    font-size: 24px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
  }
  .reveal p img {
    align: middle;
  }
  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

- Learn to add
    - title
    - subtitle
    - axis Labels

- learn to modify
    - axis range

## Syntax

<hr>

|Feature|Argument|Value|Example|
|:-------|:--------|:-----|:-------|
|Title  |main    |string|"Scatter Plot"|
|Subtitle  |sub    |string|"Displacement vs Miles Per Gallon"|
|X Axis Label  |xlab    |string|"Displacement"|
|Y Axis Label  |ylab    |string|"Miles Per Gallon"|


## Title

<hr>

```{r title, fig.align='center',fig.height=5,fig.width=5}
plot(mtcars$disp, mtcars$mpg, 
     main = 'Displacement vs Miles Per Gallon')
```

## Subtitle

<hr>

```{r subtitle, fig.align='center',fig.height=5,fig.width=5}
plot(mtcars$disp, mtcars$mpg, 
     main = 'Displacement vs Miles Per Gallon',
     sub = 'Mileage is inversely related to Displacement')
```

## Axis Labels

<hr>

```{r axis, fig.align='center',fig.height=5,fig.width=5}
plot(mtcars$disp, mtcars$mpg, 
     main = 'Displacement vs Miles Per Gallon',
     sub = 'Mileage is inversely related to Displacement',
     xlab = 'Displacement', ylab = 'Miles Per Gallon')
```

## title()

<hr>

```{r title1, fig.align='center',fig.height=5,fig.width=5}
plot(mtcars$disp, mtcars$mpg) 

title(main = 'Displacement vs Miles Per Gallon',
      sub = 'Mileage is inversely related to Displacement',
      xlab = 'Displacement', ylab = 'Miles Per Gallon')
```

## Annotate

<hr>

```{r title2, fig.align='center',fig.height=5,fig.width=5}
plot(mtcars$disp, mtcars$mpg, ann = FALSE) 

title(main = 'Displacement vs Miles Per Gallon',
      sub = 'Mileage is inversely related to Displacement',
      xlab = 'Displacement', ylab = 'Miles Per Gallon')
```

## Axis Range

<hr>

```{r axisrange, fig.align='center',fig.height=5,fig.width=5}
plot(mtcars$disp, mtcars$mpg, 
     xlim = c(0, 600), ylim = c(0, 50))
```


## Questions

<hr>

## Thank You

<hr>