---
title: "dbplyr: Data Wrangling in Database"
output: 
  revealjs::revealjs_presentation:
    transition: slide
    incremental: true
    theme: default
    highlight: haddock
    center: false
    reveal_options:
      slideNumber: true
      previewLinks: true
---

<style type="text/css">
  .reveal p {
    text-align: left;
    font-size: 20px;
  }
  .reveal body:after {
    content: url(handshake.png);
    position: fixed;
    bottom: 3.5em;
    left: 3.5em;
    box-shadow: 5px 5px 10px #000; 
  }
  .reveal hr {
    border-color: #0077aa;
  }
  .reveal h1, h2, h3, h4, h5, h6 {
    text-transform: capitalize;
    text-align: center;
    color: #0077aa;
    font-family: 'Helvetica Neue';
  }
  .reveal h1 {
    font-size: 24px;
  }
  .reveal ul {
    display: block;
  }
  .reveal ul li {
    color: #3c3c3c; 
    /* set color of list item text */
    list-style: none; 
-  }

  .reveal ul li::before {
      color: #006b6e;  
      /* color of bullet or square */
      content: "\2022"; 
      /* Unicode of character to precede the list item */
      font-size: 1em; 
      /* em or %*/ 
      padding-right: 1.1225em; 
      /* use em*/
      position: relative;
      top: 0em; 
      /* use em or % */
  }
  .reveal ul {
    line-height:150%
  }
  .reveal .fragment {
    font-size: 20px;
  }
  .reveal ol {
    display: block;
  }
</style>

## Agenda

<hr>

## Introduction

<hr>

## Libraries

<hr>

```{r litelib,eval=FALSE}
library(dbplyr)
library(dplyr)
library(DBI)
library(RSQLite)
```


```{r lite1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(dbplyr)
library(dplyr)
library(DBI)
library(RSQLite)
```

## Connect to Database

<hr>

```{r dbply1}
con <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
```

## Copy Data

<hr>

```{r dbply2}
copy_to(con, mtcars)
```

## reference copied data frame

<hr>

```{r dbply3}
mtcars2 <- tbl(con, "mtcars")
mtcars2
```

## query data

<hr>

```{r dbply4}
mtcars2 %>%
  select(mpg, cyl, drat)
```

## query data

<hr>

```{r dbply5}
mtcars2 %>%
  filter(mpg > 25)
```

## query data

<hr>

```{r dbply6}
mtcars2 %>%
  group_by(cyl) %>%
  summarise(mileage = mean(mpg))
```

## show query

<hr>

```{r dbply7}
mileages <- mtcars2 %>%
  group_by(cyl) %>%
  summarise(mileage = mean(mpg))

mileages %>% show_query()
```

## Collect Data

<hr>

```{r dbply9}
data <- mileages %>% collect()
data
```

## Questions

<hr>

## Thank You

<hr>