---
output: _output.yaml
---

## {data-background="database.png"}

## Agenda

<hr>

- select 
    - single column
    - multiple columns
    - distinct values in a column
- limit the number of records returned
- handle NULL values
- and filter columns using the following operators
    - WHERE
    - AND, or & NOT
    - BETWEEN
    - IN
    - LIKE
- aggregate data
- order data
- group data



```{r liteload, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(dbplyr)
library(dplyr)
library(DBI)
library(RSQLite)
```

## Data

<hr>

<br>
<br>

```{r show, echo=FALSE, eval=TRUE, message=FALSE}
trade <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv') %>%
  select(device, referrer, purchase, duration, n_visit, country)
trade
```

## Data Dictionary

<hr>

- id: row id
- referrer: referrer website/search engine
- device: device used to visit the website
- duration: time spent on the website (in seconds)
- country: country of origin
- purchase: whether visitor purchased
- n_visit: number of visits

```{r lite2, echo=FALSE}
con <- DBI::dbConnect(RSQLite::SQLite(), 
                      dbname = "mydatabase.db")
```

## Select Single Column

<hr>

```{sql sql1, connection = con, max.print = 6}
SELECT device FROM trade
```

## Select Multiple Columns

<hr>

Select the following fields from the `trade` table:

- `referrer`
- `device`
- `purchase`

## Select Multiple Columns

<hr>

```{sql sql2, connection = con, max.print = 6}
SELECT referrer, device, purchase  FROM trade
```

## Select All Columns

<hr>

```{sql sql3, connection = con, max.print = 6}
SELECT * FROM trade
```

## Limit

<hr>

```{sql sql4, connection = con, max.print = 6}
SELECT * FROM trade limit 10
```

## Distinct

<hr>

```{sql sql5, connection = con, max.print = 6}
SELECT distinct referrer FROM trade
```

## Filter

<hr>

Now that we know how to select columns, let us focus on filtering data. In SQL, the WHERE keyword is used to extract only those records that fulfill a specified condition. Data filter based on both text and numeric values in a table. Below are a few comparison operators we can use:

- `=` 
- `<>` 
- `<` 
- `>` 
- `<=` 
- `>=` 

## Filter

<hr>

```{sql sql6, connection = con, max.print = 6}
SELECT * FROM trade WHERE duration > 300
```

## Filter

<hr>

```{sql sql7, connection = con, max.print = 6}
SELECT * FROM trade WHERE device == 'mobile'
```

## And, Or & Not

<hr>

The WHERE clause can be combined with other operators to filter data based on more than one condition or to create more complex conditions. Below are the operators:  

- AND - displays a record if all the conditions separated by AND is TRUE
- OR  - displays a record if any of the conditions separated by OR is TRUE
- NOT - displays a record if the condition(s) is NOT TRUE

## And, Or & Not

<hr>

```{sql sql8, connection = con, max.print = 6}
SELECT * FROM trade WHERE n_visit > 3 AND duration > 100
```

## And, Or & Not

<hr>

In the next example, we will use both AND & OR. Our goal is to filter all 
rows from the `trade` table that follow the below conditions:

- `n_visit` (visit count) is either equal to 3 or 5
- `device` used to visit the website is either `mobile` or `tablet`

## And, Or & Not

<hr>

```{sql sql9, connection = con, max.print = 6}
SELECT * FROM trade WHERE (n_visit == 5 OR n_visit == 3)  
  AND (device = 'mobile' OR device = 'tablet')
```

## BETWEEN

<hr>

The BETWEEN operator selects values within a given range and is inclusive: begin and end values are included. The values can be numbers, text, or dates. In the following example, we filter rows from 
the `trade` table where the visit count is between 1 and 3, and the `device` used to visit the website is
`mobile`.

## BETWEEN

<hr>

```{sql sql10, connection = con, max.print = 6}
SELECT * FROM trade WHERE n_visit BETWEEN 1 AND 3 AND device = 'mobile'
```

## IN

<hr>

The IN operator allows us to specify multiple values in a WHERE clause. It is a shorthand 
for multiple OR conditions. In the below example, we filter rows from the `trade` table where
the visit count is either 2 or 4 or 6 or 8 or 10. Instead of using multiple OR conditions, 
we use the IN operator.

## IN

<hr>

```{sql sql11, connection = con, max.print = 6}
SELECT * FROM trade WHERE n_visit IN (2, 4, 6, 8, 10)
```

## IS NULL

<hr>

A field with a NULL value is a field with no value. If a field in a table is optional, it is possible to insert a new record or update a record without adding a value to this field. Then, the field will be saved with a NULL value. In the next example, we filter all rows from the `trade` table where the `device` column
has NULL values.

## IS NULL

<hr>

```{sql sql12, connection = con, max.print = 6}
SELECT * FROM trade WHERE device IS NULL
```

## LIKE

<hr>

The LIKE operator is used to search for a specific pattern in a column. There are two wildcards used in conjunction with the LIKE operator:

- `%` : represents zero, one, or multiple characters
- `_` : represents a single character

## LIKE

<hr>

```{sql sql13, connection = con, max.print = 6}
SELECT * FROM trade WHERE country LIKE 'P%'
```

## LIKE

<hr>

Let us look at another example where we filter all rows from the `trade` table where the name of the
country should follow the below conditions:

- name can start with any character
- the second character must be `o`
- it can have any type or number of characters after the second character

## LIKE

<hr>

```{sql sql14, connection = con, max.print = 6}
SELECT * FROM trade WHERE country LIKE '_o%'
```

## Aggregate

<hr> 

Let us combine the aggregate statements with WHERE statement to filter data. `SUM()` returns the total sum of a numeric column.

```{sql sql15, connection = con, max.print = 6}
SELECT SUM(n_visit) FROM trade
```

<br>

```{sql sql16, connection = con, max.print = 6}
SELECT SUM(n_visit) FROM trade WHERE n_visit > 5
```

## Aggregate

<hr>

`AVG()` returns the average value of a numeric column.

```{sql sql17, connection = con, max.print = 6}
SELECT AVG(n_visit) FROM trade
```

<br>

```{sql sql18, connection = con, max.print = 6}
SELECT AVG(n_visit) FROM trade 
  WHERE country LIKE 'P%'
```

## Aggregate

<hr>

`MAX()` returns the largest value of the selected column.

```{sql sql19, connection = con, max.print = 6}
SELECT MAX(n_visit) FROM trade
```

<br>

```{sql sql20, connection = con, max.print = 6}
SELECT MAX(n_visit) FROM trade 
  WHERE device == 'tablet'
```

## Aggregate

<hr>

`MIN()` returns the smallest value of the selected column.

```{sql sql21, connection = con, max.print = 6}
SELECT MIN(n_visit) FROM trade
```

<br>

```{sql sql22, connection = con, max.print = 6}
SELECT MIN(n_visit) FROM trade 
  WHERE duration BETWEEN 600 AND 900
```

## Alias 

<hr>

SQL aliases are used to give a table, or a column in a table, a temporary name. They are often used to make column names more readable. An alias only exists for the duration of the query. 

## Alias 

<hr>

```{sql sql23, connection = con, max.print = 6}
SELECT AVG(n_visit) AS avg_mobile FROM trade 
  WHERE device == 'mobile'
```

## Alias 

<hr>

```{sql sql24, connection = con, max.print = 6}
SELECT MAX(n_visit) AS max_visit FROM trade
```

## Alias 

<hr>

```{sql sql25, connection = con, max.print = 6}
SELECT MIN(duration) AS min_duration FROM trade
```

## Order By

<hr>

The ORDER BY keyword is used to sort the records in ascending or descending order. By default, the records are sorted in ascending order. Use the DESC keyword if you want to sort the records in descending order, 

```{sql sql26, connection = con, max.print = 6}
SELECT * FROM trade ORDER BY country
```

## Order By

<hr>

```{sql sql27, connection = con, max.print = 6}
SELECT * FROM trade ORDER BY duration
```

## Order By

<hr>

```{sql sql28, connection = con, max.print = 6}
SELECT * FROM trade ORDER BY n_visit DESC
```

## Group By 

<hr>

```{sql sql29, connection = con, max.print = 6}
SELECT device, count(*) AS visits 
  FROM trade 
  GROUP BY device 
  ORDER by visits DESC
```

## Group By 

<hr>

```{sql sql30, connection = con, max.print = 6}
SELECT device, MAX(duration) AS max_duration 
  FROM trade 
  GROUP BY device 
  ORDER by max_duration DESC
```

## Summary

<hr>

In this module we learnt to 

- select 
    - single column
    - multiple columns
    - distinct values in a column
- limit the number of records returned
- handle NULL values
- and filter columns using the following operators
    - WHERE
    - AND, or & NOT
    - BETWEEN
    - IN
    - LIKE
- aggregate data
- order data
- group data


## {data-background="thankyou.png"}

