---
output: _output.yaml
---

## {data-background="pipes.png"}

## Agenda

<hr>

- one sample tests
- two sample tests
- one way ANOVA

## Introduction

<hr>

- what is hypothesis testing?
- type 1 & II error
- critical value
- which test should you use
- parametric vs non-parametric

## Libraries

<hr>

- [inferr](https://inferr.rsquaredacademy.com)

```{r install, echo=FALSE, message=FALSE}
library(dplyr)
library(inferr)
```

## Data

<hr> 

```{r subset}
hsb
```

## One Sample Tests

<hr> 

- t test
- proportion test
- variance test

## One Sample t Test

<hr>

```{r ost}
infer_os_t_test(hsb, write, mu = 50, alternative = 'all')
```

## One Sample Proportion Test

<hr>

```{r osp1}
infer_os_prop_test(hsb, female, prob = 0.5)
```

## One Sample Proportion Test

<hr>

```{r osp2}
infer_os_prop_test(200, prob = 0.5, phat = 0.3)
```

## One Sample Variance Test

<hr>

```{r osv}
infer_os_var_test(mtcars, mpg, 5, alternative = 'all')
```

## Two Sample Tests

<hr> 

- independent t test
- paired t test
- proportion test
- variance test

## Independent Sample t Test

<hr>

```{r ist}
infer_ts_ind_ttest(hsb, female, write, alternative = 'all')
```

## Paired Sample t Test

<hr>

```{r pst}
infer_ts_paired_ttest(hsb, read, write, alternative = 'all')
```

## Two Sample Proportion Test

<hr>

```{r tsp}
infer_ts_prop_test(treatment, treatment1, treatment2, alternative = 'less')
```

## Two Sample Variance Test

<hr>

```{r tsv}
infer_ts_var_test(hsb, read, write, alternative = 'less')
```

## One Way ANOVA

<hr>

```{r anova}
infer_oneway_anova(hsb, write, prog)
```

## Chi Square Test of Association

<hr>

```{r chisq1}
infer_chisq_assoc_test(hsb, female, schtyp)
```

## Chi Square Goodness of Fit Test

<hr>

```{r chisq2}
infer_chisq_gof_test(hsb, race, c(20, 20, 20, 140))
```

## {data-background="thankyou.png"}

